---
title: 'Aim 1: Survival Analysis'
author: "Magali Blanco"
date: ' `r Sys.Date()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---

```{r}
# --> TO DO
# --> road proximity exposure sensitivity analysis
# --> use FSS for IPWs? 
# --> NO2 predictions from ST model
# --> stratify  by all confounders? See other General Exam
# --> add sensitivity analyses: 1) Restrict primary analysis to (say) people with 80% of their time from 1990 forward at an exact location in Puget Sound; fill in up to 20% with zip center and/or national model. Include extending the first address backward 10 years
## 2) Sensitivity analysis 1 (most restricted): The subset of those folks who are never movers living within PS at an exact address
## 3) Sensitivity 2 (least restrictive): Include anyone with zip or better at any location for any amount of time
## 4) Sensitivity 3 (using only predictions from ST model): People with 80% exact coverage and up to 20% zip, but only using ST model
##  5) Maybe sensitivity 4 (address extension): Drop address extension time (problem is that all the events before 1999 will be dropped, may cause immortal time bias, check with Lianne)

```

```{r, echo=F}
#NOTES
##4. merge with github. always do these 4 steps all together. enter the following into the Terminal (does not work any other way for some reason)
###a. git pull origin master #do this before start editing & before you're read to upload changes again
###b. git add A1.1_Survival.Rmd 
###c. git commit -m "my commit message in parentheses" 
###d. git push origin master

##########
# calculating gap lengths:
# x = c(1,2,4,NA,NA,6,NA,19,NA,NA)
# res = rle(is.na(x))
# rep(res$values*res$lengths,res$lengths)

```

```{r}
# ABOUT THIS DATASET
# exact_ variables: NAs indicate either no location, or no pollutant level info over the period of interest. Whenever you see an NA on exact_, there should also be NA in the corresponding exp_avg var

# imputed_coverage10_yr: 0 if did not assumed a constant residential hx or constant AP prediction prior to first available FOR THAT YEAR



```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=T, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE,
                      dpi=300,
                      fig.height = 4
                      )  
set.seed(1)

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
      detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

#Himsc: describe(); EnvStats: summaryFull(); survminer: ggflexsurvplot()  
pacman::p_load(knitr, kableExtra,
               Hmisc, EnvStats, lubridate, 
               survival, survminer, glmnet,
               #attach these last to use these functions as the defaults?
               tidyverse, dplyr
               )  

#other directory:/home/gradstudent/magali/a. TRAP Project/R TRAP Project/data/issue_001.sas7bdat 
# dem0 <- read_sas(file.path("~", "a. TRAP Project", "R TRAP Project", "data", "issue_001.sas7bdat"), NULL)

source("0.Global_Fns.R")
source("A1.0.2_Var&Fns.R")

##########
# dem0.0 <- haven::read_sas(file.path("Data", "Aim 1", "issue_001.sas7bdat"), NULL)

##########

# main dataset
dem0.0 <- haven::read_sas(file.path("Data", "Aim 1", "200228", "issue_001.sas7bdat"), NULL)
# exposure at age 65
dem0.0_65 <-  haven::read_sas(file.path("Data", "Aim 1", "200228", "issue_001_supp.sas7bdat"), NULL)

#remove SAS labels
labelled::var_label(dem0.0) <- NULL
labelled::var_label(dem0.0_65) <- NULL

#data that can be used for all analyses 
dem0.1 <- dem0.0 %>% 
  #drop PM10 predictions
  filter(pollutant %in% c("no2", "pm25")) %>%
  select(
    study_id,
    birthdt,
    intakedt,
    last_visit,
    birth_cohort,
    corrected_onsetdate,
    onsetage,
    #"corrected_" = final decision on outcome
    corrected_anydementia,
    corrected_dsmivdx,
    final_nindx,

    #M1, 6, sensitivity analyses
    exposure_year:exp_wks_coverage20_yr,

    #M2
    apoe,
    male,
    education, degree,
    ## household income at the longest addressed lived at prior to baseline. It may include the date of the baseline
    tr_med_inc_hshld,
    race,

    #M3
    smoke, pack_years,
    exercise, exercise_reg,
    cohort,

    #M4a
    Hypertension, Diabetes,
    CV_DIS, #indicates whether subject has ever reported having stroke, TIA, or CEA.
    Heart_Dis, #indicates whether subject has ever reported having an MI, Angina, CABG, or Angioplasty.
    bmi, bmi4,

    #M4b
    casi_irt, casi_valid,
    #address exactness
    exact_coverage01_yr:exact_coverage20_yr,

    # sensitivity analyses
    ## 1 if participants never moved
    one_location,
    imputed_coverage10_yr, imputed_coverage01_yr,
    exact_coverage10_yr, exact_coverage01_yr
    ) %>%
  #create/clean up variables
  mutate(
    #rename updated variables for correct outcomes
    nindx = final_nindx,
    dsmivdx = corrected_dsmivdx,
    anydementia = corrected_anydementia,
    onsetdate = corrected_onsetdate,
    
    age_intake = as.numeric(intakedt - birthdt)/365,  
    age_last_visit = as.numeric(last_visit - birthdt)/365,
    #traditional definition used by ACT for follow-up time for cases and controls
    age_onset = as.numeric(onsetdate - birthdt)/365, 
    age_act = ifelse(!is.na(age_onset), age_onset, age_last_visit), 
    #collapse earliest & latest cohorts since there are few ppl here 
    birth_cohort = factor(ifelse(birth_cohort <= 1905, 1895,
                           ifelse(birth_cohort >= 1935, 1935, birth_cohort))),
    # update anydementia. Unless someone was suspected of dementia and was evaluated for it, the ANYDEMENTIA (as well as DSMIVDX, ANYAD, NINDX, etc.) field will be blank. Think of ‘0’ as a confirmation of no dementia, while blank as a presumption of no dementia.
    anydementia = ifelse(is.na(anydementia), 0, anydementia),
    dsmivdx = ifelse(is.na(dsmivdx), 0, dsmivdx),
    nindx = ifelse(is.na(nindx), 0, nindx),
     #Time-varying covariates
     ##starting age is age on Jan 1st of a given year if participant was enrolled, or on intake date, whichever came later
    age_start_exposure = ifelse(format(intakedt, "%Y") == exposure_year,
                                as.numeric(intakedt-birthdt)/365,
                                as.numeric(as.Date(paste0(exposure_year, "-01-01")) - birthdt)/365), 
      #3),
    ##ending age is age at end of year or last visit date, whichever came first 
    age_end_exposure = ifelse(format(last_visit, "%Y") == exposure_year,
                                as.numeric(last_visit - birthdt)/365,
                                as.numeric(as.Date(paste0(exposure_year, "-12-31")) - birthdt)/365), 
    #3),
    ### need to change age_last_visit to age_act for sensitivity analyses
    dementia_now = ifelse(age_end_exposure < age_last_visit, 0, anydementia),
    
    ad_nincds = ifelse(nindx %in% c(1,2), 1, 0),
    ad_now = ifelse(age_end_exposure < age_last_visit, 0, ad_nincds),
    
    ## filter out invalid casi scores 
    casi_irt = ifelse(casi_valid ==1, casi_irt, NA),
    
    # Roy - accidentally made this a character?
    tr_med_inc_hshld = as.numeric(tr_med_inc_hshld),
    
    income_cat = ifelse(tr_med_inc_hshld < 35000, 1,  
                                      ifelse(tr_med_inc_hshld >= 35000 & tr_med_inc_hshld < 50000, 2,
                                             ifelse(tr_med_inc_hshld >= 50000 & tr_med_inc_hshld < 75000, 3, 4)
                                             )),  
    race_white = ifelse(race == 1, 1, 0),
    # make unknown degree=9 an 6 ("other") to avoid creating more NAs; 
    # Rachel/Lianne - change 9 to none? 
    degree = ifelse(degree != 9, degree, 0),
    # combine GED and HS. Note, there is now no degree==2, but shoudn't matter since this is treated as a categorical variable
    degree = ifelse(degree %in% c(1:2), 1, degree)
    ) %>%
  select(
  study_id,
  birthdt, birth_cohort, cohort, intakedt, onsetdate, last_visit,
  age_intake, age_act, age_last_visit,

  anydementia, dsmivdx, nindx, ad_nincds,
  dementia_now, ad_now,

  age_start_exposure, age_end_exposure,
  exposure_year, pollutant, exp_avg01_yr:exp_wks_coverage20_yr,

  apoe:degree, tr_med_inc_hshld, income_cat,race, race_white, smoke:bmi4, casi_irt,

  one_location,
  imputed_coverage10_yr, imputed_coverage01_yr,
  exact_coverage10_yr, exact_coverage01_yr
  )
 
#how long have participants been followed? 
yr <- dem0.1 %>%
  group_by(study_id) %>%
  mutate(
    intake_yr = as.numeric(min(format(intakedt, "%Y"))),
    last_visit_yr = as.numeric(min(format(last_visit, "%Y"))),
    fu_yrs = last_visit_yr - intake_yr +1
    ) %>%
  select(study_id, intake_yr, last_visit_yr, fu_yrs) %>%
  unique()

# 1 row per follow-up/exposure year
enrollment_yrs <-rep(yr$study_id, yr$fu_yrs) %>%
  as.data.frame() %>%
  rename(study_id = ".") %>%
  left_join(yr, by = "study_id") %>%
  group_by(study_id) %>%
  mutate(
    #use min() b/c need 1 value and all vector values are same
    enrollment_yr = seq(1:min(fu_yrs)),
    exposure_year = seq(from=min(intake_yr), to=min(last_visit_yr))) %>%
  select(study_id, fu_yrs, enrollment_yr, exposure_year) %>%
  as.data.frame()

dem1 <-  dem0.1 %>% 
  #keep all enrollment years
  right_join(enrollment_yrs, by = c("study_id", "exposure_year")) %>%
  select(
  study_id:age_end_exposure,
  fu_yrs,
  enrollment_yr,

  exposure_year:casi_irt,

  one_location,
  imputed_coverage10_yr, # "doesn't exist"?? #imputed_coverage01_yr,
  exact_coverage10_yr, exact_coverage01_yr
  ) %>%
  filter(
    #drop ppl with only a baseline observation
    intakedt != last_visit,
    #drop exposure predictions after the last visit when individuals are no longer "at risk"  
    ## --> note: when using act_age, have to change last_visit to onsetdate
    exposure_year <= as.numeric(format(last_visit, "%Y")),  
    ) %>%
  mutate(
    # Delete _original?
    #copy all original predictions (for sensitivity analysis using min.pct = 0.75 vs 0.95)
    exp_avg01_yr_original = exp_avg01_yr,
    exp_avg05_yr_original = exp_avg05_yr,
    exp_avg10_yr_original = exp_avg10_yr,
    exp_avg20_yr_original = exp_avg20_yr,
    exp_avg10_yr10yrlag_original = exp_avg10_yr10yrlag,
    exp_avg10_yr05yrlag_original = exp_avg10_yr05yrlag,
    
    # [don't do this for now b/c coverage variable is impacted by interpolation assumption when participants first enroll]
    #only keep predictions w/ at least x% coverage (for primary analysis)
    # exp_avg01_yr = ifelse(exp_wks_coverage01_yr >= min.pct, exp_avg01_yr, NA),
    # exp_avg05_yr = ifelse(exp_wks_coverage05_yr >= min.pct, exp_avg05_yr, NA),
    # exp_avg10_yr = ifelse(exp_wks_coverage10_yr >= min.pct, exp_avg10_yr, NA),
    # exp_avg20_yr = ifelse(exp_wks_coverage20_yr >= min.pct, exp_avg10_yr, NA),
    # exp_avg10_yr10yrlag = ifelse(exp_wks_coverage10_yr10yrlag >= min.pct, exp_avg10_yr10yrlag, NA),
    # exp_avg10_yr05yrlag = ifelse(exp_wks_coverage10_yr05yrlag >= min.pct, exp_avg10_yr05yrlag, NA),
    
    )   

#make wide format
no2 <- dem1 %>% 
  filter(pollutant == "no2") %>%
  select(-pollutant) %>%
  rename(
    #rename values to be used in main analysis
    no2_1yr = exp_avg01_yr,
    no2_5yr = exp_avg05_yr,
    no2_10yr = exp_avg10_yr,
    no2_20yr = exp_avg20_yr,
    no2_10yr10yrlag = exp_avg10_yr10yrlag,
    no2_10yr05yrlag = exp_avg10_yr05yrlag,
    
    # delete _original? don't need?
    #rename original columns
    no2_1yr_original = exp_avg01_yr_original,
    no2_5yr_original = exp_avg05_yr_original,
    no2_10yr_original = exp_avg10_yr_original,
    no2_20yr_original = exp_avg20_yr_original,
    no2_10yr10yrlag_original = exp_avg10_yr10yrlag_original,
    no2_10yr05yrlag_original = exp_avg10_yr05yrlag_original,
    
    no2_coverage_1yr = exp_wks_coverage01_yr,
    no2_coverage_5yr = exp_wks_coverage05_yr,
    no2_coverage_10yr = exp_wks_coverage10_yr,
    no2_coverage_20yr = exp_wks_coverage20_yr,
    no2_coverage_10yr10yrlag = exp_wks_coverage10_yr10yrlag,
    no2_coverage_10yr05yrlag = exp_wks_coverage10_yr05yrlag,
    
    no2_imputed_coverage10_yr = imputed_coverage10_yr,
    no2_exact_coverage10_yr = exact_coverage10_yr, no2_exact_coverage01_yr = exact_coverage01_yr
    
    ) 

pm25 <- dem1 %>% 
  filter(pollutant == "pm25") %>%
  select(-pollutant) %>%
  rename(
    pm25_1yr = exp_avg01_yr,
    pm25_5yr = exp_avg05_yr,
    pm25_10yr = exp_avg10_yr,
    pm25_20yr = exp_avg20_yr,
    pm25_10yr10yrlag = exp_avg10_yr10yrlag,
    pm25_10yr05yrlag = exp_avg10_yr05yrlag,
    
    # delete _original? don't need?
    #rename original columns
    pm25_1yr_original = exp_avg01_yr_original,
    pm25_5yr_original = exp_avg05_yr_original,
    pm25_10yr_original = exp_avg10_yr_original,
    pm25_20yr_original = exp_avg20_yr_original,
    pm25_10yr10yrlag_original = exp_avg10_yr10yrlag_original,
    pm25_10yr05yrlag_original = exp_avg10_yr05yrlag_original,
    
    pm25_coverage_1yr = exp_wks_coverage01_yr,
    pm25_coverage_5yr = exp_wks_coverage05_yr,
    pm25_coverage_10yr = exp_wks_coverage10_yr,
    pm25_coverage_20yr = exp_wks_coverage20_yr,
    pm25_coverage_10yr10yrlag = exp_wks_coverage10_yr10yrlag,
    pm25_coverage_10yr05yrlag = exp_wks_coverage10_yr05yrlag,
    
     pm25_imputed_coverage10_yr = imputed_coverage10_yr,
    pm25_exact_coverage10_yr = exact_coverage10_yr, pm25_exact_coverage01_yr = exact_coverage01_yr
    )   
dem.w <- full_join(no2, pm25)  

# add exposure at age 65
dem0.1_65 <- dem0.0_65 %>%
  select(
    study_id,
    exposure_year_65 = exposure_year,
    pollutant, exp_avg10_age_65,
    exp_wks_coverage10_age_65, exact_coverage10_age_65, imputed_coverage10_age_65
  )
 
dem_65_no2 <- dem0.1_65 %>% 
   filter(pollutant== "no2") %>%
   select(-pollutant) %>%
   rename(
    no2_exp_avg10_age_65 = exp_avg10_age_65,
    no2_exp_wks_coverage10_age_65 = exp_wks_coverage10_age_65,
    no2_exact_coverage10_age_65 = exact_coverage10_age_65,
    no2_imputed_coverage10_age_65 = imputed_coverage10_age_65
   )
    
dem_65_pm25 <- dem0.1_65 %>% 
   filter(pollutant== "pm25") %>%
   select(-pollutant) %>%
   rename(
    pm25_exp_avg10_age_65 = exp_avg10_age_65,
    pm25_exp_wks_coverage10_age_65 = exp_wks_coverage10_age_65,
    pm25_exact_coverage10_age_65 = exact_coverage10_age_65,
    pm25_imputed_coverage10_age_65 = imputed_coverage10_age_65
   )
 
dem_65 <- full_join(dem_65_no2, dem_65_pm25)

dem.w <- left_join(dem.w, dem_65)
 
#add variables for alternative analysis using age at onset (cases) or age at last visit (noncases)
dem.w2 <- dem.w %>%
  mutate(
    last_act_fu_yr = ifelse(!is.na(onsetdate), year(onsetdate),
                             year(last_visit)),
    age_end_exposure2 = ifelse(last_act_fu_yr == exposure_year,
                                age_act,
                                as.numeric(as.Date(paste0(exposure_year, "-12-31")) - birthdt)/365),  
    dementia_now2 = ifelse(age_end_exposure2 < age_act, 0, anydementia),
    ad_now2 = ifelse(age_end_exposure2 < age_act, 0, ad_nincds),
  ) %>%
  #drop exposure predictions after individuals are no longer "at risk"  
  filter(exposure_year <= last_act_fu_yr)


# create a fixed covariates dataset 
fixed.covariates <- dem.w %>%
  select(
    study_id:ad_nincds,
    apoe:casi_irt
    ) %>% 
  #make all variables numeric; code has values starting at 0, whereas as.numeric() starts at 1, subtract 1 to fix
  mutate(
    birth_cohort = as.numeric(birth_cohort),
    birth_yr = as.numeric(format(birthdt, "%Y")),
    intake_yr = as.numeric(format(intakedt, "%Y")),
    onset_yr = as.numeric(format(onsetdate, "%Y")),
    last_visit_yr = as.numeric(format(last_visit, "%Y")),
    
    # # Roy - accidentally made this a character?
    # tr_med_inc_hshld = as.numeric(tr_med_inc_hshld)
    ) %>%
  #drop dates, which are not "numeric" 
  select(
    -birthdt, -intakedt, -onsetdate, -last_visit,
      -onset_yr
  ) %>%
  #one row per individual
  unique() 

# str(fixed.covariates)

```

# Check variables  

## NO2, 10 yr

Person-years with missing NO2 predictions 

```{r}
table(no2_missing = is.na(dem.w$no2_10yr))

```

Distribution of NO2 predictions 

### --> why are there negative predictions? 

```{r}
dem.w %>%
  ggplot(aes(x=no2_10yr)) + 
  geom_histogram() + 
  geom_vline(aes(xintercept=0, col = "Prediction < 0")) +
  scale_color_manual(name = "", values = c('Prediction < 0' = "red"))  

```

## PM2.5, 10 yr

Person-years with missing PM2.5 predictions 

```{r}
table(pm2.5_missing = is.na(dem.w$pm25_10yr))

```

Distribution of PM2.5 predictions 

```{r}
dem.w %>%
  ggplot(aes(x=pm25_10yr)) + 
  geom_histogram()

```

## APOE 

Plots of APOE availability by enrollment year. See slightly proportinally more missing APOE readings later in time. This may be because they have not yet had their blood work processed.
 
```{r}
apoe_avail_ct_p <- dem.w %>%
  filter(enrollment_yr==1) %>%
  mutate(exposure_year = factor(exposure_year)) %>%
  ggplot(aes(x=exposure_year, fill = !is.na(apoe))) + 
  geom_bar() + 
  theme(axis.text.x = element_text( angle = 90)) + 
  labs(x = "Exposure year",
       y = "Count",
      fill = "APOE available"
       )

apoe_avail_prop_p <- dem.w %>%
  filter(enrollment_yr==1,
         ) %>%
  mutate(exposure_year = factor(exposure_year)) %>%
  ggplot(aes(x=exposure_year, fill = !is.na(apoe))) + 
  geom_bar(position = "fill") + 
  theme(axis.text.x = element_text( angle = 90)) + 
  labs(x = "Exposure year",
       y = "proportion",
       fill = "APOE available"
       )

ggarrange(apoe_avail_ct_p,
          apoe_avail_prop_p, 
          ncol=1,   
          common.legend = T, legend = "bottom",
          heights = c(1,1), 
          labels = "Enrolless by APOE availability"
          )

```

## Age variables 

Make sure age variables make sense.

```{r, include=F}
#check that ages make sense ages
table(round(dem.w$age_intake, 1) <= round(dem.w$age_act, 1))  
#View(dem.w[which(dem.w$age_intake > dem.w$age_act),]) 
table(dem.w$age_intake <= dem.w$age_last_visit)
#View(dem.w[which(dem.w$age_intake > dem.w$age_last_visit),]) #study_id: 4647 has wrong last_visit/intakedt? last_visit is correct. "Will be updated in next freeze"
table(dem.w$age_act <= dem.w$age_last_visit)

#dementia 
dementia_counts <- dem.w %>% 
  group_by(study_id) %>%
  dplyr::summarize(
    #study_id = mean(study_id),
    #number of times each patient has dementia_now==1 (should be max 1)
    anydementia = mean(anydementia),
    dem_now_n = sum(dementia_now==1)
  ) 

##check that individuals who have anydementia=1 also have dementia_now=1 & those who don't don't  
dementia_counts %>%
  with(., table(anydementia, dem_now_n))

##check that individuals only have 1 dementia_now
dementia_counts %>%
  with(., table(dem_now_n <= 1))
#View(dem.w[which(dem.w$age_intake > dem.w$age_last_visit),]) 
#study_id 1027 has 2 dementia_now's b/c last visit age vs end of year age, which rounded to same thing
#View(dem.w[dem.w$study_id==1027,]) 

#AD 
ad_counts <- dem.w %>% 
  group_by(study_id) %>%
  dplyr::summarize(
    ad_nincds = mean(ad_nincds),
    ad_now_n = sum(ad_now==1)
  ) 
##check that individuals who have ad_nincds=1 also have ad_now=1 & those who don't don't #looks good
ad_counts %>%
  with(., table(ad_nincds, ad_now_n))

##check that individuals only have 1 ad_now
ad_counts %>%
  with(., table(ad_now_n <= 1))

```

## median income    

There are only a few people with missing values.

```{r}

table(missing = is.na(fixed.covariates$tr_med_inc_hshld))

```

## Predictions at age 65 

```{r}
first_st_no2_pred <- 1979

```

Number of predictions prior to first available prediction from the spatiotemporal model (`r first_st_no2_pred`)

```{r}
dem.w %>% 
  filter(enrollment_yr ==1) %>%
  dplyr::summarize(no = sum(exposure_year_65 < first_st_no2_pred, na.rm = T)) %>%
  as.numeric()

```

```{r}
dem.w %>%
  filter(enrollment_yr==1) %>%
  ggplot(aes(x=exposure_year_65)) + 
  geom_histogram() + 
  geom_vline(aes(xintercept = first_st_no2_pred, col = "first available ST prediction")) + 
  scale_color_manual(name = "", values = c('first available ST prediction' = "red")) +
  labs(title = "Distribution of years used to assess exposure at age 65") + 
  theme(legend.position = "bottom")

```

## Coverage Proportion

Plot Coverage variable against participant enrollment year. Important to look at this variable because we were originally excluding person-years based on this variable

```{r}
# See that coverage is more likely to be 1 at the begining of participant enrollment. This is because the coverage variables takes the imputed variables prior to enrollment into account. 
```

### --> ? no one has full coverage for a 10 yr exposure period? 

```{r}
t <- dem.w %>% 
  select(study_id, enrollment_yr, exposure_year, no2_10yr, no2_10yr_original, no2_coverage_10yr, no2_imputed_coverage10_yr, one_location) 

t %>%
  ggplot(aes(x=enrollment_yr, y = no2_coverage_10yr)) + 
  geom_point(alpha = 0.05) + 
  geom_hline(yintercept = 0.95, col = "red") +
  geom_smooth() + 
  labs(
    x = "Participant Enrollment Year in Study",
    y = "NO2 Coverage past 10yr",
    title = "Coverage during Past 10 Years for person-years",
    caption = "coverage calculated from exp_wks_coverage10_yr for NO2"
  )

```

summary of 10 yr no2 coverage

```{r}
summary(t$no2_coverage_10yr)
#distribution.table(t, "no2_coverage_10yr", round.int = 2)

```


Plot Coverage variable against calendar year.
 
```{r}
t %>%
  ggplot(aes(x=exposure_year, y = no2_coverage_10yr)) + 
  geom_point(alpha = 0.05) + 
  geom_hline(yintercept = 0.95, col = "red") +
  geom_smooth() + 
  labs(
    x = "Calendar Year",
    y = "NO2 Coverage past 10yr"
  )

```


## Imputed coverage 

Imputation occurs at enrollment if we do not have sufficient residential history or AP predictions going back far enough. Imputed coverage should drop to 0 by enrollment year ~ 10.

```{r}
#  
dem.w %>%
  ggplot(aes(x=enrollment_yr, y = no2_imputed_coverage10_yr)) + 
  geom_point(alpha = 0.05) + 
  geom_smooth() + 
  labs(
    x = "Participant Enrollment Year in Study",
    y = "NO2 Imputed Coverage past 10 yrs"
  )

```

Summary of imputed coverage variable ("no2_imputed_coverage10_yr")

```{r}
summary(dem.w$no2_imputed_coverage10_yr)

```

We are more likey to impute their coverage if they enrolled earlier in the study.  

### --> are imputed coverage values decreasing over time b/c we have residential hx for these people through the Group Health/KP system even if they were not in the study?

```{r}
dem.w %>%
  ggplot(aes(x=exposure_year, y = no2_imputed_coverage10_yr)) + 
  geom_point(alpha = 0.05) + 
  geom_smooth() + 
  labs(
    x = "Year",
    y = "NO2 Imputed Coverage past 10 yrs"
  )

```


## Exact Address Coverage  

```{r}
#### by calendar year 

dem.w %>%
  ggplot(aes(x=exposure_year, y = no2_exact_coverage01_yr)) + 
  geom_point(alpha = 0.05) + 
  geom_smooth() + 
  labs(
    x = "Year",
    y = "Exact Address (1 yr)"
  )

``` 

```{r}
dem.w %>%
  ggplot(aes(x=exposure_year, y = no2_exact_coverage10_yr)) + 
  geom_point(alpha = 0.05) + 
  geom_smooth() + 
  labs(
    x = "Year",
    y = "Exact Address (10 yr)"
  )





#### by enrollment year

dem.w %>%
  ggplot(aes(x=enrollment_yr, y = no2_exact_coverage01_yr)) + 
  geom_point(alpha = 0.05) + 
  geom_smooth() + 
  labs(
    x = "Participant Enrollment Year in Study",
    y = "Exact Address (1 yr)"
  )

```
 
Summary of exact coverage (no2_exact_coverage01_yr)

```{r}
summary(dem.w$no2_exact_coverage01_yr)

```

Summary of exact coverage (no2_exact_coverage10_yr)

```{r}
summary(dem.w$no2_exact_coverage10_yr)

```

### --> use exact address for 1 yr or 10 yrs?

Proportion of the time participants have an exact address during their entire follow-up time (mean(no2_exact_coverage_01_yr)). Analysis drops people without an address for a particular year. 

For addresses we do know, only ~ 5% of people have an address history that is exact less than 86% of the time. Or, participant address history is exact at least 86% of the time for ~95% of participants. 

```{r}
# [?? Not using annual estimates of exact address because some subjects have no addresses for particular years and have "NAs" even though they do have 10 year estimates that are used to calculate avg 10-year exposures.]

```

```{r}
dem.w %>%
  group_by(study_id) %>%
  dplyr::summarize(exact_address_prop = mean(no2_exact_coverage01_yr, na.rm=T)) %>%
  select(exact_address_prop) %>%
  describe()
  
 

```

### Amanda's suggested primary analysis:

Only including people with an exact address at least 80% of their follow-up time. 

```{r}
#  (don't have people living at an exact address 100% of the time for a 10 yr period)
```

```{r}
exact_prop_threshold <- 0.8

```

Number of individuals dropped using an exact address threshold of `r exact_prop_threshold` based on mean(no2_exact_coverage01_yr).

### --> since this is a time-varying analysis, should person-years (rather than entire subjects) be dropped if the exact variables falls below a threshold of ~ 80% for the previous 10 yrs (??or each year)? i.e., we can't be certain that our exposure estimate is correct for most of the period of interest.

### --> ? update exact coverage variable to 10 yr??

```{r}
dem.w %>%
    group_by(study_id) %>%
  dplyr::summarize(
    exact_address_prop = mean(no2_exact_coverage01_yr, na.rm=T)
  ) %>%
  filter(exact_address_prop < exact_prop_threshold) %>%
  nrow()

```


Similar to Amanda's sensitivity analysis (1):

Only including people who never moved and living at an exact address ____ at least 80% of the time ___ (don't have ppl living at an exact address 100% of the time or info for seeing who is in ST area). 

Proportion of the time participants have an exact address during their entire follow-up time (mean(no2_exact_coverage10_yr)).

### --> ? update exact coverage variable to 1/10 yr??

```{r}

dem.w %>%
  filter(one_location==1) %>%
    group_by(study_id) %>%
  dplyr::summarize(exact_address_prop = mean(no2_exact_coverage10_yr, na.rm=T)) %>%
  filter(exact_address_prop >= exact_prop_threshold) %>%
  select(exact_address_prop) %>%
  describe()

```

Proportion of the time participants have an exact address during their entire follow-up time (mean(no2_exact_coverage01_yr)).

```{r}

dem.w %>%
  filter(one_location==1) %>%
    group_by(study_id) %>%
  dplyr::summarize(exact_address_prop = mean(no2_exact_coverage01_yr, na.rm=T)) %>%
  filter(exact_address_prop >= exact_prop_threshold) %>%
  select(exact_address_prop) %>%
  describe()

```

Similar to Amanda's sensitivity analysis (3):

Using only predictions from ST model, people with 80% exact coverage and up to 20% zip.

Average proportion of the time for which we have exact addresses for each subject, during the previous 10 years.  

```{r}

```

### Address gaps. How many people have these? How long are they?

An address gap is defined as having an NA in the 1 yr exact coverage variable (no address of any kind).


```{r}
# # Distribution of no2_exact_coverage01_yr.
# 
# dem.w$no2_exact_coverage01_yr %>%
#   describe()

```

Note that some people have address gaps 100% of the time. Their exposures must be based on a previous address since there are estimate for longer exposure periods (e.g., ID: 1004 has no 1 yr estimates, but does have 5 yr estimates with coverage that starts ~ 0.73 and decreases over time).

Using exact_coverage01_yr for NO2 and PM2.5.

```{r}

dem.w %>%
  mutate(address_unknown = is.na(no2_exact_coverage01_yr)) %>%
  group_by(study_id) %>%
  dplyr::summarize(
    Years_of_unknown_address =  sum(address_unknown),
    Proportion_of_time =  mean(address_unknown)
  ) %>%
  #View()
  gather(-study_id, key = var, value = value) %>%
  ggplot(aes(x=value)) + 
  geom_histogram() +
  facet_wrap(~var, scales = "free") + 
  labs(title = "Address gaps per participant",
       subtitle = "An address gap is defined as having an NA in the 1 yr NO2 exact coverage variable"
       )

 
dem.w %>%
  mutate(address_unknown = is.na(pm25_exact_coverage01_yr)) %>%
  group_by(study_id) %>%
  dplyr::summarize(
    Years_of_unknown_address =  sum(address_unknown),
    Proportion_of_time =  mean(address_unknown)
  ) %>%
  #View()
  gather(-study_id, key = var, value = value) %>%
  ggplot(aes(x=value)) + 
  geom_histogram() +
  facet_wrap(~var, scales = "free") + 
  labs(title = "Address gaps per participant",
       subtitle = "An address gap is defined as having an NA in the 1 yr PM25 exact coverage variable"
       )

```

 
 
## Compare numbers vs Rachel's

```{r, include=F}
summary(dem.w$age_end_exposure)

summary(dem.w2$age_end_exposure2 )

summary(dem.w$age_act) 
#summary(dem.w$age_act[dem.w$anydementia==1]) 

summary(dem.w2$last_act_fu_yr)
 

summary(fixed.covariates$age_last_visit)

table(dem.w$dementia_now)
table(dem.w2$dementia_now2)

summary(fixed.covariates$age_last_visit)
summary(fixed.covariates$age_act) 

summary(fixed.covariates$age_last_visit[fixed.covariates$anydementia==1]) 
summary(fixed.covariates$age_act[fixed.covariates$anydementia==1]) 

```

# IPW since there is a lot of missing APOE values 

Proportion of individuals with missing APOE status 

```{r}
sum(is.na(fixed.covariates$apoe))/nrow(fixed.covariates)

```

### --> check that "with.nas" (variables w/ missingness) is correct in final dataset. for ease, could make mynumeric all ==0 so mode is calculated for any variable w/ missingness

### --> run primary analysis w/ filled in dataset or use a full-case analysis for IPW? 

```{r}
# see B540 L7-8
# Approach:
#1. fill in missing values with their most likely option (mean for continuous, mode for categorical variables)
#2. use selection model (e.g. lasso for glm) to select predictors of whether or not we have APOE values for people: apoe_available (in apoe.w)
# 3. fit logistic model to determine probability of observing an observation given predictors: glm(apoe_available ~ [predictors]). use model w/ dataset to get predicted values (probability of being observed), 
# 4. calculate stablized weights 

apoe0 <- fixed.covariates %>%
  mutate(apoe_available = !is.na(apoe)) %>% 
  #don't inclue race b/c causes issues (creates a very large wt for race==5)
 dplyr::select(-apoe, -race) 

#1. replace missing values w/ most likely outcome  
##variables w/ missing values  
with.nas <- names(which(colSums(is.na(apoe0)) > 0))
# "education"    "tr_med_inc_hshld" "income_cat" "race_white"   "smoke"        "pack_years"   "exercise"     "exercise_reg" "Hypertension" "Diabetes" "CV_DIS"       "Heart_Dis"    "bmi"          "bmi4"         "casi_irt"      

#1 = TRUE, 0 = FALSE
mynumeric = c(0, 1, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 0, 1)  #c(0, 1, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 0, 1, 0)

apoe <- apoe0  
#replace  missing values w/ mean (if numeric) or mode (if factor)
for (i in seq_along(with.nas)) {
  #vector w/ missing values 
  myvar <- with.nas[i]  
  apoe[[myvar]] <- replace.nas.fn(apoe[[myvar]], 
                                  numeric = mynumeric[i])
  }

#check that there are no more NAs #looks good
#describe(apoe) 
#table(!is.na(apoe))
 
#2. Use Lasso to select variables most predictive of APOE being present
#apoe2 <-apoe 

apoe <- apoe %>%
  #convert character variables to numberic
  mutate_if(is.character, as.numeric) %>%
  #convert these to factors
  mutate_at(c("birth_cohort", 
              "nindx",
              "education",
              "degree",
              "income_cat",
              #"race_white",
              "smoke",
              "bmi4"
              ), 
            as.factor)

lasso_results <- lasso_fn(dt = apoe, x_names = names(apoe)[2:(ncol(apoe)-1)], y_name = "apoe_available", family. = "binomial", 
                          lambda. = ""
                          )

lasso.vars <- lasso_results$results$cov

#select selected variables if they are in main dataset as is or if the name is prior to a number. prevents similarly-named variables (e.g. exercise vs exercise_reg) from being selected
#save names & relabel categorized bmi4 variable for now (has issues in loop below b/c of #4 - selects both "bmi" & "bmi4") 
lasso_names <- lasso.vars %>%
  str_replace(pattern = "bmi4", replacement = "bmi_cat" )
apoe_names <- names(apoe) %>%
  str_replace(pattern = "bmi4", replacement = "bmi_cat" )

selected_cov <- vector()

for (i in seq_len(length(apoe_names))) {
  #i=19
  find_pattern <- paste0("^", apoe_names[i], "$", "|", "^", apoe_names[i], "\\d")
  if(any(grepl(find_pattern, lasso_names)  )) {
    
    selected_cov <- append(selected_cov, apoe_names[i])
    }
}

#change BMI back to orinal name
selected_cov <- selected_cov %>%
  str_replace(pattern = "bmi_cat", replacement = "bmi4" )

apoe <- apoe %>%
  select(study_id,
         #y
        apoe_available,
        #categories selected via Lasso 
        selected_cov)  

#3. model to predict APOE variable being present
apoe.m <- apoe %>%
  select(-study_id) %>%
  glm(apoe_available ~ .,
  family = "binomial",
  data = .) 
#apoe.m %>% summary()
##denominator: probability of APOE being present
apoe_available_prob <- predict(apoe.m, type = "response") 

#min(apoe_available_prob) #check that probability isn't Tiny (like when used "race")
 
#numerator: stabilizer 
apoe.male.m <- apoe %>%
  glm(apoe_available ~ male,
  family = "binomial",
  data = .) 
 
apoe_male_prob <- predict(apoe.male.m, type = "response") 

apoe.wts <- apoe %>% 
  select(study_id) %>%
  mutate(model_wt = apoe_male_prob/apoe_available_prob)
#summary(apoe.wts$model_wt)

# 4. take the inverse of probabilities to calculate "weights", use these weights in coxph(, weights = c()) 
dem.w <- dem.w %>% 
  #add stabilized weights  
  left_join(apoe.wts, by = "study_id")

#for sensitivity analyses 
dem.w2 <- dem.w2 %>% 
  #add stabilized weights  
  left_join(apoe.wts, by = "study_id")

```

Lasso-selected covariates

```{r}
selected_cov

```

Stabilizing weights for IPW based on logistic regression models. Using lasso-selected variable predictors of APOE availability for the denominator and a gender predictor for the numerator.

```{r}

summary(dem.w$model_wt)

```


```{r}
#Rachel uses FSS & variables: cohort, race, diabetes, gender, degree, CV_Dis, birth cohort, intake age
# see 556 wk 4 lab; machine learning hw 5 

#library(leaps)
# fss = regsubsets(Y~., data=df, method = "forward", )
# reg.fwd.summary= summary(fss)

#library(MASS)




# par(mfrow=c(2,2))
# which.max(reg.fwd.summary$adjr2) #max r2 @ p=3
# plot(reg.fwd.summary$adjr2, xlab="Number of Variables", ylab="Adjusted Rsq", type = "l")
# points(3,reg.fwd.summary$adjr2[3], col="red", cex=2, pch=20)
# 
# p = which.min(reg.fwd.summary$cp) #find min cp  
# plot(reg.fwd.summary$cp, xlab="Number of Variables", ylab="Cp", type = "l")
# points(p,reg.fwd.summary$cp[p], col="red", cex=2, pch=20)
# 
# p = which.min(reg.fwd.summary$bic) #find min bic  
# plot(reg.fwd.summary$bic, xlab="Number of Variables", ylab="BIC", type = "l")
# points(p,reg.fwd.summary$bic[p], col="red", cex=2, pch=20)




```
 

# Descriptive Statistics   
## Sample size in entire dataset 

```{r, echo=F}
#sample size 
##no. in cohort 
cohort_ids <- dem.w %>%
  select(study_id) %>%
  unique() 

n_cohort <- length(cohort_ids$study_id)

##no. of cases
cases_ids <- dem.w %>%
  filter(anydementia==1) %>%
  select(study_id) %>%
  unique() 

n_cases <- length(cases_ids$study_id)

## no. noncases
non_cases_ids <-dem.w %>%
  filter(anydementia==0) %>%
  select(study_id) %>%
  unique()  

n_noncases <- length(non_cases_ids$study_id)

case_proportion <- round(n_cases/n_cohort, 2)
  
n_personyears <- nrow(dem.w)

data.frame(cohort = n_cohort, 
          cases = n_cases, 
          non_cases =n_noncases,
          case_proportion = case_proportion,
                total_person_years = n_personyears
           ) %>%
  kable(caption = "Sample size") %>%
  kable_styling()
 
```

## Dementia by calendar time and enrollment age

there are cases every year. 

```{r dementia}
### --> why does 1st plot show no cases in 1995 but second does? 

#enrolled participants & dementia incidence over time
dem.w %>%
  mutate(dementia_now = factor(dementia_now, levels = c(1,0))) %>%
  ggplot(aes(x= exposure_year, 
             fill= dementia_now)) + 
  geom_bar() + 
  labs(title = "enrolled participants",
       fill = "dementia incidence"
       ) #+ scale_y_log10()


#dementia incidence over time
dem.w %>%
  filter(dementia_now == 1) %>%
  ggplot(aes(x= exposure_year)) + 
  geom_bar() + 
  labs(title= "number of dementia incident cases over time")

```

Dementia Incidence by participant age at enrollment. Shows that those who enrill later on in life are more likely to develop dementia. What we would expect (but different than what Gail has seen?).

```{r}

fixed.covariates %>%
  mutate(age_intake_decade = cut(age_intake, breaks = 5)) %>%
  ggplot(aes(x=age_intake_decade)) + 
  geom_bar(position = "fill",
           aes(fill = anydementia ==1)) + 
  geom_text(stat='count',
            aes(label= paste0("N = ", ..count..),
                y = ..prop..),
            vjust=-1) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Proportion of participants who develop dementia by age of enrollment",
       x = "Age at Enrollment", 
       y = "Proportion who develop dementia",
       fill = "Develop dementia"
       )

```


## Distribution of key covariates.    

Check that categories have a sufficient number of individuals. Do any have less than ~50 individuals?

```{r, fig.height=10}

#histograms
fixed.covariates %>% 
  mutate(birth_cohort = as.numeric(birth_cohort)) %>%
  gather(key = "variable",  value = "value", -study_id) %>%   
  ggplot(aes(x=value)) + 
  geom_histogram() +  
    facet_wrap(~variable, scales = "free")

```

## Prediction source  

### ---> Do predictions come from the Spatiotemporal or National model? 

```{r}


```

  
## Table 1
Participant baseline characteristics. Percents are calculated based on the number of individuals with available data (missing values removed). Some participants did not have baseline exposure estimates and are thus excluded from above/below median columns. 
 
How others have stratified:   
-most proivde a table for the entire cohort (? & don't state which went into models?)   
- entire cohort vs dementia cases (chen 2017)   
- by exposure quartiles at baseline (baseline was during same time period, e.g., 1993-1995) (Oudin 2016; Chang 2014) ["baseline" was same time period for cohort]   
- mean personal AP  

Dementia

### --> OK to calculate %s across columns (No. participants) & within a column (e.g., edu)?

```{r}
#table1
m2_variables <- c("apoe", "male", "race_white", "income_cat", "degree", "birth_cohort") 

t1_tot <- dem.w %>%
  # only keep individuals included in Model 2
  drop_na(m2_variables) %>%
  #1st year of enrollment (baseline)
  filter(enrollment_yr == 1) %>%
  t1.fn() %>%
  mutate_all(~str_replace(., "0 \\(NaN%\\)", "NA")) %>%
  rename(Total = V1)

```

```{r}

t1_dem <- dem.w %>%
  # only keep individuals included in Model 2
  drop_na(m2_variables) %>%
  #1st year of enrollment (baseline)
  filter(enrollment_yr == 1) %>%
  group_by(anydementia) %>%
  t1.fn() %>%
  #replace "NaN" estimates for APOE w/ "NA"
  mutate_all(~str_replace(., "0 \\(NaN%\\)", "NA")) %>%
  rename("No Dementia" = V1,
         "Dementia" = V2) %>%
#remove grouping rows
  slice(-c(1)) %>%
  # add total column
  right_join(t1_tot, .)

t1_dem %>%
  column_to_rownames() %>%
  kable(caption = "Cohort characteristics by dementia incidence", 
        align = "c") %>%
  add_indent(c(5:11,15:20,22:25, 27:29, 32:35, 37:39, 44:46)) %>%
  add_footnote(c(
    "Table includes participants in primary analysis (those with complete observations for: APOE status, gender, race, Census Tract income, education and birth cohort)",
    "Baseline cognition is based on CASI scores that have been adjusted for accuracy using Item Response Theory (IRT). Higher numbers indicate greater cognitive function."
    ))

## save T1 as csv
# write.csv(t1_dem, file.path("Output", "Aim 1", "Tables", "Table 1.csv"), row.names = F)

```

```{r}
## saves as html
# as_image(t1_dem, file = file.path(images_path0, "t1_dem.png"), width=9)

# saves as png but image is poor quality
#save_kable(t1_dem, file = file.path(images_path0, "t1_dem.png"))

## issue - prints black image?
# kable_as_image(kable_input = t1_dem, filename = file.path(images_path0, "test") )


```

 
AD

```{r}

t1_ad <- dem.w %>%
  drop_na(m2_variables) %>%
  filter(enrollment_yr == 1) %>%
  group_by(ad_nincds) %>%
  t1.fn() %>%
  mutate_all(~str_replace(., "0 \\(NaN%\\)", "NA")) %>%
  rename("No AD" = V1,
         "AD" = V2) %>%
#remove grouping rows
  slice(-c(1)) %>%
  right_join(t1_tot, .)  

t1_ad %>%
    column_to_rownames() %>%
  kable(caption = "Cohort characteristics by Alzheimer's Disease incidence", 
        align = "c") %>%
  add_indent(c(5:11,15:20,22:25, 27:29, 32:35, 37:39, 44:46)) %>%
  add_footnote(c(
    "Table includes participants in primary analysis (those with complete observations for: APOE status, gender, race, Census Tract income, education and birth cohort)",
    "Baseline cognition is based on CASI scores that have been adjusted for accuracy using Item Response Theory (IRT). Higher numbers indicate greater cognitive function."
    ))

```

## Exposure Distribution

Plot of NO2 exposure over time

```{r}
# ? delete? 

# long_fu <- 15
# 
# one_loc_ids <- dem.w %>%
#   filter(one_location==1,
#          fu_yrs >= long_fu) 
# 
# one_loc_ids <- unique(one_loc_ids$study_id )

``` 

### --> ?? NO2 prediction error? plots all look weird

No2 exposure by participant age and birth cohort 

```{r}

dem.w %>%
  ggplot(aes(x=age_start_exposure, y = no2_1yr,
             col = birth_cohort
             )) +
  geom_point(alpha=0.05) +
  geom_smooth(col="black") + 
  geom_smooth(aes(col = birth_cohort)) + 
    labs(x = "Age (years)",
         y = "NO2 Exposure (ppb)",
         col = "Birth Cohort",
         title = "NO2 exposure by participant age",
         caption = "Black loess line is for all person-years"
           )  

```

PM2.5 exposure by participant age and birth cohort 

```{r}
dem.w %>%
  ggplot(aes(x=age_start_exposure, y = pm25_1yr,
             col = birth_cohort
             )) +
  geom_point(alpha=0.05) +
  geom_smooth(col="black") + 
  geom_smooth(aes(col = birth_cohort)) + 
    labs(x = "Age (years)",
         y = "PM2.5 Exposure (ug/m3)",
         col = "Birth Cohort",
         title = "PM2.5 exposure by participant age",
         caption = "Black loess line is for all person-years"
           ) 

```


Table 2. No2 & PM2.5 over time.

```{r}
# exposure over time for everyone in cohort

# NO2 
all_yrs <- dem.w %>%
  mutate(Years = "All") %>%
  group_by(Years,  
           ) %>%
  drop_na(no2_10yr) %>%
  distribution.table(var.string = "no2_10yr")

by_yr <- dem.w %>%
      drop_na(no2_10yr) %>%
  mutate(Years = cut_interval(x = exposure_year, n = 4, 
                                     dig.lab=4)) %>%
  group_by(Years) %>%
    distribution.table(var.string = "no2_10yr")

no2_all <- rbind(all_yrs, by_yr) 

no2_all %>%
  kable(caption = "10-yr average NO2 (ppb) exposure predictions over time for everyone in Cohort (N = person-years)") %>%
  kable_styling()

# PM2.5
all_yrs <- dem.w %>%
  mutate(Years = "All") %>%
  group_by(Years,  
           ) %>%
  drop_na(pm25_10yr) %>%
  distribution.table(var.string = "pm25_10yr")

by_yr <- dem.w %>%
      drop_na(pm25_10yr) %>%
  mutate(Years = cut_interval(x = exposure_year, n = 4, 
                                     dig.lab=4)) %>%
  group_by(Years) %>%
    distribution.table(var.string = "pm25_10yr")

pm25_all <- rbind(all_yrs, by_yr) 

pm25_all %>%
  kable(caption = "10-yr average PM2.5 (ug/m3) exposure predictions over time for everyone in Cohort (N = person-years)") %>%
  kable_styling()
 
```

## Correlations of variables in models     

See high correlatin (R^2) between NO2 and PM2.5. Results may be confounded by PM2.5.

```{r}
#dataset with model variables of interest for descriptive purposes (similar dataset generated in models.fn())
model.vars <- dem.w %>%
  select(
    #m1 & sensitivity
    no2_1yr:no2_20yr ,
    #m2
    apoe,
    male,
    race_white,
    income = income_cat,
    edu = degree,
    birth_cohort,
    #m3
    smoke,
    exercise_reg,
    #m4
    Hypertension,
    Diabetes, 
    CV_DIS,
    Heart_Dis,
    bmi = bmi4,
    #m6
    pm25_1yr:pm25_20yr,
    #pm25_10yr
    ) %>%
  mutate(
  #make factors
    income = factor(income),
    edu = factor(edu),
    birth_cohort = factor(birth_cohort),
    smoke = factor(smoke),
    
    #make 10 ppb units
    no2_10ppb_1yr = no2_1yr/my.no2.units,
    no2_10ppb_5yr = no2_5yr/my.no2.units,
    no2_10ppb_10yr = no2_10yr/my.no2.units,
    no2_10ppb_20yr = no2_20yr/my.no2.units,
    no2_10ppb_10yr10yrlag = no2_10yr10yrlag/my.no2.units,
    no2_10ppb_10yr05yrlag = no2_10yr05yrlag/my.no2.units,
    
    #make 5 mg/m3
    pm25_1ugm3_1yr = pm25_1yr/my.pm25.units,
    pm25_1ugm3_5yr = pm25_5yr/my.pm25.units,
    pm25_1ugm3_10yr = pm25_10yr/my.pm25.units,
    pm25_1ugm3_20yr = pm25_20yr/my.pm25.units,
    pm25_1ugm3_10yr10yrlag = pm25_10yr10yrlag/my.pm25.units,
    pm25_1ugm3_10yr05yrlag = pm25_10yr05yrlag/my.pm25.units,
   ) %>%
  #drop these columns to avoid confusion, since using 10 ppb for NO2 and 5 ug/m3 for pm2.5
  select(-c(no2_1yr:no2_20yr, 
            pm25_1yr:pm25_20yr)) 

# correlation plot
model.vars %>% 
  mutate(
    #make numeric for cor() fn 
    income = as.numeric(income),
    edu = as.numeric(edu),
    birth_cohort = as.numeric(birth_cohort),
    smoke = as.numeric(smoke)
    ) %>%
  select(apoe:bmi, no2_10ppb_10yr, pm25_1ugm3_10yr) %>%
  rename(no2_10yr = no2_10ppb_10yr,
         pm25_10yr = pm25_1ugm3_10yr) %>%
  cor(method = "pearson",
      use ="complete.obs") %>%
  corrplot::corrplot(method="color") 

```

R2 for 10 yr NO2 and PM2.5 exposure

```{r}
with(dem.w, cor(no2_10yr, pm25_10yr,
                method = "pearson",
                use = "complete.obs")^2) %>% round(2)

```

# Inferential Statistics

## Primary Analysis, Redued & Extended Models

Dementia

### --> note, usring clusetr = study_id in order for coxph() to calculate robust SEs

```{r, fig.height=8}

dem_10yr <- models.fn(mydata = dem.w,
                models = paste0("m", c(1, 2, "3a", "3b",  "4a", "4b", 5, 6)),  
                surv.time2 = "age_end_exposure", 
                surv.event = "dementia_now", 
                no2.var = "no2_10yr",
                pm25.var = "pm25_10yr")

#dem_10yr$raw_model_output[7]

dem_10yr_hrs <- dem_10yr$hrs %>%
  mutate(
    Model = c("1",
           "2", 
           "3a", 
           "3b", 
           "4a", 
           "4b", 
           "5", 
           "5", 
           "6"),
    Description = c("Reduced",
           "Primary", 
           "Extended", 
           "Extended, ACT Cohort", 
           "Extended & Mediation", 
           "Extended & Mediation, CASI Score", 
           "Interaction, no APOE", 
           "Interaction, APOE", 
           "PM2.5 Adjusted")
    )

 
# dem_10yr_hrs

#m2 for other exposure years
dem_other_yrs <- data.frame()
expo.prds <- paste0("no2_", c("1yr", "5yr", "20yr", "10yr10yrlag", "10yr05yrlag"))

for (i in seq_along(expo.prds)) {
  df <- models.fn(mydata = dem.w,
                models = "m2",
                surv.time2 = "age_end_exposure", 
                surv.event = "dementia_now", 
                no2.var = expo.prds[i])
  
  dem_other_yrs <- rbind(dem_other_yrs, df$hrs)
  }

dem_other_yrs <- dem_other_yrs %>%
  mutate(
    Model = "2",
    #Description = "Different exposure period"
    Description = c("1 yr Exposure", 
                    "5 yr Exposure", 
                    "20 yr Exposure", 
                    "10 yr 5 yr Lag Exposure", 
                    "10 yr 10 yr Lag Exposure")
    )

#HR table 
dem_hrs <- rbind(dem_10yr_hrs, dem_other_yrs) %>%
  mutate(Exposure = factor(Exposure, levels = c("1yr", "5yr", "10yr", "20yr", "10yr10yrlag", "10yr05yrlag")))

# delete? show this later w/ all other estimates
# dem_hrs %>% kable(caption = "Dementia HRs") %>%
#   kable_styling()

# #HR plot 
# dem_hrs %>% hr.plot(., outcome.var = "dementia")  

```

AD 

```{r, fig.height=8}
ad_10yr <- models.fn(mydata = dem.w,
                models = paste0("m", c(1, 2, "3a", "3b",  "4a", "4b", 5, 6)),  
                surv.time2 = "age_end_exposure", 
                surv.event = "ad_now", 
                no2.var = "no2_10yr",
                pm25.var = "pm25_10yr")

ad_10yr_hrs <- ad_10yr$hrs %>%
  mutate(
    Model = c("1",
           "2", 
           "3a", 
           "3b", 
           "4a", 
           "4b", 
           "5", 
           "5", 
           "6"),
    Description = c("Reduced",
           "Primary", 
           "Extended", 
           "Extended, ACT Cohort", 
           "Extended & Mediation", 
           "Extended & Mediation, CASI Score", 
           "Interaction, no APOE", 
           "Interaction, APOE", 
           "PM2.5 Adjusted")
    )

#m2 for other exposure years
ad_other_yrs <- data.frame()

for (i in seq_along(expo.prds)) {
  df <- models.fn(mydata = dem.w,
                models = "m2",
                surv.time2 = "age_end_exposure", 
                surv.event = "ad_now", 
                no2.var = expo.prds[i])
  
  ad_other_yrs <- rbind(ad_other_yrs, df$hrs)
  }

ad_other_yrs <- ad_other_yrs %>% 
  mutate(Model = "2",
         #Description = "Different exposure period"
         Description = c("1 yr Exposure", 
                    "5 yr Exposure", 
                    "20 yr Exposure", 
                    "10 yr 5 yr Lag Exposure", 
                    "10 yr 10 yr Lag Exposure")
  )

#HR table 
ad_hrs <- rbind(ad_10yr_hrs, ad_other_yrs) %>%
  mutate(Exposure = factor(Exposure, levels = c("1yr", "5yr", "10yr", "20yr", "10yr10yrlag", "10yr05yrlag")))  

# delete? show this later w/ all other estimates
# ad_hrs %>% kable(caption = "AD HRs") %>%
#   kable_styling()
# 
# #HR plot 
# ad_hrs %>% hr.plot(., outcome.var = "AD")   

```

## Other Sensitivity Analyses    

### M2 using age at ACT onset diagnosis (cases) or last visit (noncases) 

Dementia 

```{r}

#check that new coding is correct #looks good
# t <- dem.w2 %>% select(study_id, onsetdate, last_visit, last_act_fu_yr, 
#                        exposure_year,
#                        age_start_exposure, age_end_exposure, age_end_exposure2,
#                        dementia_now, dementia_now2,
#                        ad_now, ad_now2) 
# View(t)

# dem.w2 %>%
#   filter(exposure_year == last_act_fu_yr) %>%
#   ggplot(aes(x=age_end_exposure, y= age_end_exposure2)) + 
#   geom_point(alpha=0.15) +
#   geom_abline(intercept = 0, slope = 1, col = "red") + 
#   labs(x = "age at last visit (everyone)", 
#        y = "age at onset (cases) or last visit (non-cases)")

dem_10yr_act <- dem.w2 %>%
  models.fn(.,
            models = paste0("m2"),
            surv.time2 = "age_end_exposure2", 
            surv.event = "dementia_now2", 
            no2.var = "no2_10yr",
            pm25.var = "pm25_10yr") 

dem_10yr_act <- dem_10yr_act$hrs %>%
  mutate(Model = "2",
         Description = "Onset Age"
         )

# dem_10yr_act %>% kable(caption = "Dementia HR using age at onset (cases) or last visit (non-cases)") %>%
#   kable_styling()

```

AD

```{r}
ad_10yr_act <- dem.w2 %>%
  models.fn(.,
            models = paste0("m2"),
            surv.time2 = "age_end_exposure2", 
            surv.event = "ad_now2", 
            no2.var = "no2_10yr",
            pm25.var = "pm25_10yr")  

ad_10yr_act <- ad_10yr_act$hrs %>%
  mutate(Model = "2",
         Description = "Onset age"
         )

# ad_10yr_act %>% kable(caption = "AD HR using age at onset (cases) or last visit (non-cases)") %>%
#   kable_styling()

```

### M2 without IPW for missing APOE status

Dementia

```{r}
dem_10yr_m2_wt1 <- dem.w %>%
  mutate(model_wt = 1) %>%
  models.fn(mydata = .,
            models = "m2",
            surv.time2 = "age_end_exposure",
            surv.event = "dementia_now",
            no2.var = "no2_10yr",
            pm25.var = "pm25_10yr")
 
dem_10yr_m2_wt1_hr <- dem_10yr_m2_wt1$hrs %>%
  mutate(Model = "2",
         Description = "No IPW")

# dem_10yr_m2_wt1_hr %>% kable(caption = "unweighted HR for dementia incidence based on 10 yr avg exposure") %>%
#   kable_styling()

```

AD

```{r}
ad_10yr_m2_wt1 <- dem.w %>%
  mutate(model_wt = 1) %>%
  models.fn(mydata = .,
            models = "m2",
            surv.time2 = "age_end_exposure",
            surv.event = "ad_now",
            no2.var = "no2_10yr",
            pm25.var = "pm25_10yr")
 
ad_10yr_m2_wt1_hr <- ad_10yr_m2_wt1$hrs %>%
  mutate(Model = "2",
         Description = "No IPW")

# ad_10yr_m2_wt1_hr %>% 
#   kable(caption = "unweighted HR for AD incidence based on 10 yr avg exposure") %>%
#   kable_styling()


```

### [? DELETE?] keep exposure observations if address available at least 75% of time 

Dementia

```{r}

# dem_address_0.75 <- dem.w %>%
#   filter(no2_coverage_10yr >= 0.75) %>%
#   models.fn(mydata = ., models = "m2",
#                         surv.time2 = "age_end_exposure",
#                         surv.event = "dementia_now",
#                         no2.var = "no2_10yr_original")
# 
# dem_address_0.75_hrs <- dem_address_0.75$hrs %>%
#   mutate(Model = "2",
#          Description = "Address available at least 75% of time") 

```

AD

```{r}

# ad_address_0.75 <- dem.w %>%
#   filter(no2_coverage_10yr >= 0.75) %>%
#   models.fn(mydata = ., models = "m2",
#                         surv.time2 = "age_end_exposure",
#                         surv.event = "ad_now",
#                         no2.var = "no2_10yr_original")
# 
# ad_address_0.75_hrs <- ad_address_0.75$hrs %>%
#   mutate(Model = "2",
#          Description = "Address available at least 75% of time") 

 
```


### Only include individuals for whom we have complete air pollution predictions and address history during the exposure period of interest

note: for a 10 yr exposure period, we should have complete AP predictions for everyone who has an address starting ~ 1999 (first adddresses available 1989 + 10 yrs)

Dementia

```{r}
dem_no_imputation <- dem.w %>%
  # --> ? or only keep individuals who have Never had imputed coverage?
  filter(no2_imputed_coverage10_yr == 0) %>%
  models.fn(mydata = .,
                models = "m2",
                surv.time2 = "age_end_exposure",
                surv.event = "dementia_now",
                no2.var = "no2_10yr",
                pm25.var = "pm25_10yr")

#dem_no_imputation$raw_model_output
dem_no_imputation_hrs <- dem_no_imputation$hrs %>%
  mutate(Model = "2",
         Description = "No Address Hx assumptions")  #or early AP

```

AD

```{r}
ad_no_imputation <- dem.w %>%
  # --> ? or only keep individuals who have Never had imputed coverage?
  filter(no2_imputed_coverage10_yr == 0) %>%
  models.fn(mydata = .,
                models = "m2",
                surv.time2 = "age_end_exposure",
                surv.event = "ad_now",
                no2.var = "no2_10yr",
                pm25.var = "pm25_10yr")

ad_no_imputation_hrs <- ad_no_imputation$hrs %>%
  mutate(Model = "2",
         Description = "No Address Hx Assumptions")  #or early AP

```

### Only include individuals who have never moved

one_location variable is fixed for any one individual (not time-varying).

Dementia


```{r}
# ? used to have issue w/ birth cohort b/c few incident cases per strata? but OK now.

# unlikely to develop dementia in later cohorts
# with(dem.w, table(bc=birth_cohort, dem = anydementia)) %>% prop.table(margin = 1)

dem_one_location <- dem.w %>%
  filter(one_location == 1) %>%
#mutate(birth_cohort = as.character(birth_cohort),
  # get error & large estiamtes for last 2 birth_cohort categories otherwise if don't do this
  #birth_cohort = ifelse(birth_cohort %in% c(1930, 1935), 1925, birth_cohort)
 # ) %>%  
  models.fn(mydata = .,
                models = "m2",
                surv.time2 = "age_end_exposure",
                surv.event = "dementia_now",
                no2.var = "no2_10yr",
                pm25.var = "pm25_10yr")

#dem_one_location$raw_model_output

dem_one_location_hrs <- dem_one_location$hrs %>%
  mutate(Model = "2",
         Description = "Never moved") 

```

AD

```{r}
# # unlikely to develop dementia in later cohorts
# with(dem.w, table(bc=birth_cohort, dem = ad_nincds)) %>% prop.table(margin = 1)

ad_one_location <- dem.w %>%
  filter(one_location == 1) %>%
  #mutate(birth_cohort = as.character(birth_cohort),
       # get error & large estiamtes for last 2 birth_cohort categories otherwise if don't do this
       #birth_cohort = ifelse(birth_cohort %in% c(1930, 1935), 1925, birth_cohort)
   #      ) %>%  
  models.fn(mydata = .,
                models = "m2",
                surv.time2 = "age_end_exposure",
                surv.event = "ad_now",
                no2.var = "no2_10yr",
                pm25.var = "pm25_10yr")


# ad_one_location$raw_model_output
ad_one_location_hrs <- ad_one_location$hrs %>%
  mutate(Model = "2",
         Description = "Never moved") 

```

### Fixed AP exposure

--> note: ? may assume constant air pollution predictions if estimate is prior to first available prediction estimate (~ 1979) 

#### AP at age 65

Dementia

```{r}

dem_age_65 <- dem.w %>%
  # fixed exposure 
  filter(enrollment_yr == 1) %>%
  models.fn(mydata = .,
            models = "m2", 
            surv.time1 = "age_intake",
            surv.time2 = "age_last_visit",
            surv.event = "anydementia",
            no2.var = "no2_exp_avg10_age_65")

dem_age_65_hrs <- dem_age_65$hrs %>%
  mutate(
    Exposure = "10yr",
    Model = "2",
    Description = "Fixed exposure at age 65") 

```

AD

```{r}
ad_age_65 <- dem.w %>%
  # fixed exposure 
  filter(enrollment_yr == 1) %>%
  models.fn(mydata = .,
            models = "m2", 
            surv.time1 = "age_intake",
            surv.time2 = "age_last_visit",
            surv.event = "ad_nincds",
            no2.var = "no2_exp_avg10_age_65")

ad_age_65_hrs <- ad_age_65$hrs %>%
  mutate(
    Exposure = "10yr",
    Model = "2",
    Description = "Fixed exposure at age 65") 

```


#### AP at study entry

Dementia 

```{r}
dem_age_entry <- dem.w %>%
  # fixed exposure 
  filter(enrollment_yr == 1) %>%
  models.fn(mydata = .,
            models = "m2", 
            surv.time1 = "age_intake",
            surv.time2 = "age_last_visit",
            surv.event = "anydementia",
            no2.var = "no2_10yr")

dem_age_entry_hrs <- dem_age_entry$hrs %>%
  mutate(
    Exposure = "10yr",
    Model = "2",
    Description = "Fixed exposure at entry") 

```

AD

```{r}
ad_age_entry <- dem.w %>%
  # fixed exposure 
  filter(enrollment_yr == 1) %>%
  models.fn(mydata = .,
            models = "m2", 
            surv.time1 = "age_intake",
            surv.time2 = "age_last_visit",
            surv.event = "ad_nincds",
            no2.var = "no2_10yr")

ad_age_entry_hrs <- ad_age_entry$hrs %>%
  mutate(
    Exposure = "10yr",
    Model = "2",
    Description = "Fixed exposure at entry") 

```

### Road Proximity exposure

Dementia

```{r}

```

AD

```{r}

```



### Optional/time permitting 

1. finer adjustments for confounders
2. calculate IPW weights based on the probability that All of the M2 predictors (not just APOE) are observed (e.g., all_covs_available ~ [predictors])
3. one-location. very small sample size 


```{r}


```


## Compare all HRs

### Dementia

### --> note that fixed covariates are for persons, not person-years 

```{r, fig.height=8}
#combine HRs from all analyses 

dem_all_hrs <- rbind(
  #reduced, primary & extended models, using 10 yr exposure
  dem_hrs,
  #M2 using age at ACT onset diagnosis (cases) or last visit (noncases)  
  dem_10yr_act,
  #no IPW
  dem_10yr_m2_wt1_hr,
  #address min 75%
  #dem_address_0.75_hrs,
  # no assumptions about initial address hx or early AP
  dem_no_imputation_hrs,
  # using fixed exposure at age 65
  dem_age_65_hrs,
  # using fixed exposure at entry
  dem_age_entry_hrs
  
  # never moved
  #dem_one_location_hrs
  ) %>%
  mutate(
    model_description = paste0(Model, ". ", Description),
    hr_ci = paste0(hr, " (", lower_limit, ", ", upper_limit, ")"),
    Exposure = recode_factor(Exposure,
                              "1yr" = "1 yr",
                              "5yr" = "5 yr",
                              "10yr" = "10 yr",
                              "20yr" = "20 yr",
                              "10yr10yrlag" = "10 yr 10 yr lag",
                              "10yr05yrlag" = "10 yr 5 yr lag"
                              ),
    Description = factor(Description, levels = Description)
    )  


# table
dem_all_hrs %>%
  select(
    Model, Description, Exposure,
    hr_ci,
    person_years, number_events
    ) %>%
  rename(
    "HR (CI)" = hr_ci,
    "Person-years" = person_years,
    "No. events" = number_events,
    ) %>%
  kable() %>%
  kable_styling()

 

#plot HRs
dem_all_hrs %>% hr.plot(., outcome_var = "dementia")

# plots separated
# ## M1-6
# dem_all_hrs %>% hr.plot.m1_m6(dt = ., outcome.var = "dementia")
# ## diff exposure periods
# dem_all_hrs %>% hr.plot.diff.expo.prds(dt = ., outcome.var = "Dementia")
# #other sensitivity analyses
# dem_all_hrs %>% hr.plot.other.models(dt = ., outcome.var = "dementia")

```

 
### AD

```{r, fig.height=8}
#AD
ad_all_hrs <- rbind(
  #reduced, primary & extended models, using 10 yr exposure 
  ad_hrs, 
  #M2 using age at ACT onset diagnosis (cases) or last visit (noncases)  
  ad_10yr_act, 
  #no IPW
  ad_10yr_m2_wt1_hr,
  #address min 75%
  #ad_address_0.75_hrs,
  # no assumptions about initial address hx or early AP
  ad_no_imputation_hrs,
  # using fixed exposure at age 65
  ad_age_65_hrs,
  # using fixed exposure at entry
  ad_age_entry_hrs
  # never moved
  #ad_one_location_hrs
) %>%
  mutate(
    model_description = paste0(Model, ". ", Description),
    hr_ci = paste0(hr, " (", lower_limit, ", ", upper_limit, ")"),
    Exposure = recode_factor(Exposure,
                              "1yr" = "1 yr",
                              "5yr" = "5 yr",
                              "10yr" = "10 yr",
                              "20yr" = "20 yr",
                              "10yr10yrlag" = "10 yr 10 yr lag",
                              "10yr05yrlag" = "10 yr 5 yr lag"
                              ),
    Description = factor(Description, levels = Description)
  )

# table
ad_all_hrs %>%
  select(
    Model, Description, Exposure,
    hr_ci,
    person_years, number_events
    ) %>%
  rename(
    "HR (CI)" = hr_ci,
    "Person-years" = person_years,
    "No. events" = number_events,
    ) %>%
  kable() %>%
  kable_styling()

ad_all_hrs %>% hr.plot(., outcome_var = "AD")

# # Plots seprated
# #M1-6
# ad_all_hrs %>% hr.plot.m1_m6(dt = ., outcome.var = "AD")
# #diff exposure periods
# ad_all_hrs %>% hr.plot.diff.expo.prds(dt = ., outcome.var = "AD")
# #other models
# ad_all_hrs %>% hr.plot.other.models(dt = ., outcome.var = "AD")

```

# Running models for PM2.5 as the primary exposure (Rachel's analysis)

Dementia 

-reduced, primary, exntended models; different exposure windows

### --> why is the NO2 adjusted diff than the PM2.5 adjusted above?? Should be the same model & use the same data
### --> ? change PM2.5 units to 10 µg/m3 to see if HRs get wider?  

```{r}
# NOTE: treating PM2.5 as if it were my primary exposure, thus treating it as "NO2" in the models. 
dem_10yr_pm25 <- models.fn(mydata = dem.w,
                models = paste0("m", c(1, 2, "3a", "3b",  "4a", "4b", 5, 6)),  
                surv.time2 = "age_end_exposure", 
                surv.event = "dementia_now", 
                no2.var = "pm25_10yr",
                pm25.var ="no2_10yr", 
                no2.units = 1, 
                pm25.units = 10)

#dem_10yr$raw_model_output[7]

dem_10yr_pm25_hrs <- dem_10yr_pm25$hrs %>%
  mutate(
    Model = c("1",
           "2", 
           "3a", 
           "3b", 
           "4a", 
           "4b", 
           "5", 
           "5", 
           "6"),
    Description = c("Reduced",
           "Primary", 
           "Extended", 
           "Extended, ACT Cohort", 
           "Extended & Mediation", 
           "Extended & Mediation, CASI Score", 
           "Interaction, no APOE", 
           "Interaction, APOE", 
           "NO2 Adjusted")
    )

 
# dem_10yr_hrs

#m2 for other exposure years
dem_other_yrs_pm25 <- data.frame()
expo.prds <- paste0("pm25_", c("1yr", "5yr", "20yr", "10yr10yrlag", "10yr05yrlag"))

for (i in seq_along(expo.prds)) {
  df <- models.fn(mydata = dem.w,
                models = "m2",
                surv.time2 = "age_end_exposure", 
                surv.event = "dementia_now", 
                no2.var = expo.prds[i])
  
  dem_other_yrs_pm25 <- rbind(dem_other_yrs_pm25, df$hrs)
  }

dem_other_yrs_pm25 <- dem_other_yrs_pm25 %>%
  mutate(
    Model = "2",
    Description = c("1 yr Exposure", 
                    "5 yr Exposure", 
                    "20 yr Exposure", 
                    "10 yr 5 yr Lag Exposure", 
                    "10 yr 10 yr Lag Exposure")
    )

#HR table 
dem_hrs_pm25 <- rbind(dem_10yr_pm25_hrs, dem_other_yrs_pm25) %>%
  mutate(
    Exposure = substr(as.character(Exposure), 2, nchar(as.character(Exposure))),
    Exposure = factor(Exposure, levels = c("1yr", "5yr", "10yr", "20yr", "10yr10yrlag", "10yr05yrlag")))

```

-onset age

```{r}
dem_10yr_act_pm25 <- dem.w2 %>%
  models.fn(.,
            models = paste0("m2"),
            surv.time2 = "age_end_exposure2", 
            surv.event = "dementia_now2", 
            no2.var = "pm25_10yr",
            pm25.var = "no2_10yr") 

dem_10yr_act_pm25 <- dem_10yr_act_pm25$hrs %>%
  mutate(Model = "2",
         Exposure = substr(as.character(Exposure), 2, nchar(as.character(Exposure))),
         Description = "Onset Age"
         )


```

Table & plot of all HRs

```{r}
dem_all_hrs_pm25 <- rbind(
  #reduced, primary & extended models, using 10 yr exposure
  dem_hrs_pm25,
  #M2 using age at ACT onset diagnosis (cases) or last visit (noncases)  
  dem_10yr_act_pm25
  ) %>%
  mutate(
    model_description = paste0(Model, ". ", Description),
    hr_ci = paste0(hr, " (", lower_limit, ", ", upper_limit, ")"),
    Exposure = recode_factor(Exposure,
                              "1yr" = "1 yr",
                              "5yr" = "5 yr",
                              "10yr" = "10 yr",
                              "20yr" = "20 yr",
                              "10yr10yrlag" = "10 yr 10 yr lag",
                              "10yr05yrlag" = "10 yr 5 yr lag"
                              ),
    Description = factor(Description, levels = Description)
    )  

```

```{r, fig.height=8}

# table
dem_all_hrs_pm25 %>%
  select(
    Model, Description, Exposure,
    hr_ci,
    person_years, number_events
    ) %>%
  rename(
    "HR (CI)" = hr_ci,
    "Person-years" = person_years,
    "No. events" = number_events,
    ) %>%
  kable() %>%
  kable_styling()

#plot HRs
dem_all_hrs_pm25 %>% hr.plot(., outcome_var = "dementia", title_pollutant = "PM2.5 (ug/m3)")

```













# Cox model assumptions and inference validity    
1. Proportional hazards Assumption    
a) Schoenfeld residuals. If the proportional hazards model is correctly specified, there should be no trend in the residuals over time (uncorrelated, with mean zero across event times)

```{r, eval=F}
m2 <- dem_10yr$raw_model_output$m2

```

```{r, eval=F}
## ?? interpretation of schenresid matrix?
schoenresid = residuals(m2, type = "scaledsch")

time = as.numeric(rownames(schoenresid))

# link no2_10yr & "times" column
cbind(resid=schoenresid[,1], time) %>%
  as.data.frame() %>%
  ggplot(aes(x=time, y=resid)) + 
  geom_point() + 
  geom_smooth() + 
  labs(y = "scaled Schoenfeld residuals for NO2 coefficient",
       x = "failure time (age at dementia incidence)")

```

b. formally test. using cox.zph(). if global p-val is significant, assumption is violated   
-assumption is not violated 

```{r, eval=F}
cox.zph(m2)

```

2. linear relationship between continuous covariates and log hazard of the outcome    
If a correct model is used, these residuals are uncorrelated and have mean nearly zero.


Martingale residuals are used to check if the functional form of the covariate is okay (eg. linear vs quadratic terms). 

```{r, eval=F}
##??
martingalesid = residuals(m2, type = "martingale", data = dem.w) %>%
  as.data.frame() %>%
  #df rows used in M2
  rownames_to_column(.) 
names(martingalesid)[names(martingalesid) %in% "."] <- "martingaleresid" 
  
#only keep rows used in M2. 
dem.w.m2 <- dem.w[martingalesid$rowname,] 

plot.martingaleresid <- cbind(resid=martingalesid$martingaleresid, no2_10yr=dem.w.m2$no2_10yr) %>%
  as.data.frame() 

plot.martingaleresid %>%
  ggplot(aes(x=no2_10yr, y=resid)) +
  geom_point(alpha=0.3) +
  geom_smooth() +
  labs(y = "Martingale residuals",
       x = "NO2 (ppb), 10 yr avg"
      )

#mean, +-SD  # looks good, mean is close to 0?
print("mean, SD")
mean(plot.martingaleresid$resid)
sd(plot.martingaleresid$resid)

```

3. no outliers or influential observations, $\triangle \beta$
 
```{r, fig.height=8, eval=F}
# B537 L3, slide 125

dfbetas <- residuals(m2, type="dfbeta") 
#rename columns
colnames(dfbetas) <- names(coef(m2))

dfbetas %>%  
  as.data.frame() %>%
  #get rows used in model
  rownames_to_column(var = "ind") %>%
  mutate(ind = as.numeric(ind)) %>%
  #make long format
  gather(covariate, value, -ind) %>%
  as.data.frame() %>%
  #plot
  ggplot(aes(x=ind, y=value)) +
  geom_point() +
  facet_wrap(~covariate, scales="free") + 
  labs(x="observation index",
       y= "delta beta")


```
