---
title: "Aim 1: Survival Analysis"
author: "Magali Blanco"
date: " `r Sys.Date()` "
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
editor_options: 
  chunk_output_type: console
---
```{r, echo=F}
#NOTES
##4. merge with github. always do these 4 steps all together. enter the following into the Terminal (does not work any other way for some reason)
###a. git pull origin master #do this before start editing & before you're read to upload changes again
###b. git add A1.1_Survival.Rmd 
###c. git commit -m "my commit message in parentheses" 
###d. git push origin master

# see BIOST 537 Ch 5 Lecture, slide 33/57 

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache=T, cache.comments = F, message = F, warning = F, tidy.opts=list(width.cutoff=60), tidy=TRUE, fig.height = 8)  

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
      detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(survival, haven, tidyverse, knitr, Hmisc, EnvStats)  #Himsc: describe(); EnvStats: summaryFull() #coxphw: coxphw() for non-prop hazards

# #minimum percent of months required for us to consider a prediction "valid"
# min.pct <- 0.95

#To open doc:/home/gradstudent/magali/a. TRAP Project/R TRAP Project/data/issue_001.sas7bdat 

# dem0 <- read_sas(file.path("~", "a. TRAP Project", "R TRAP Project", "data", "issue_001_3.sas7bdat"), NULL)

source("A1.0.2_Var&Fns.R")

#dem0.0 <- read_sas(file.path("Data", "Aim 1", "issue_001_3.sas7bdat"), NULL)
dem0.0 <- read_sas(file.path("Data", "Aim 1", "issue_001_4.sas7bdat"), NULL)

#remove SAS labels
labelled::var_label(dem0.0) <- NULL

#data that can be used for all analyses 
dem0.1 <- dem0.0 %>% 
  filter(
    #drop PM10 predictions
    pollutant %in% c("no2", "pm25")
    ) %>%
  select(
    study_id,
    birthdt,
    intakedt,
    last_visit,
    birth_cohort,
    corrected_onsetdate,
    onsetage,
    #"corrected_" = final decision on outcome
    corrected_anydementia,  
    corrected_dsmivdx,
    final_nindx,
    
    #M1, 6, sensitivity analyses
    exposure_year:exp_mos_coverage20_yr, 
    
    #M2
    apoe,
    male,
    education, degree,
    tr_med_inc_hshld, 
    race,
    
    #M3
    smoke, pack_years, 
    exercise, exercise_reg,
    
    #M4a
    Hypertension, Diabetes,
    CV_DIS, #indicates whether subject has ever reported having stroke, TIA, or CEA.
    Heart_Dis, #indicates whether subject has ever reported having an MI, Angina, CABG, or Angioplasty.
    bmi, bmi4,
    
    #M4b
    casi_irt, casi_valid
    ) %>%
  
  #create/clean up variables
  mutate(
    #rename updated variables for correct outcomes
    nindx = final_nindx,
    dsmivdx = corrected_dsmivdx,
    anydementia = corrected_anydementia,
    onsetdate = corrected_onsetdate,
    
    age_intake = round(as.numeric(intakedt - birthdt)/365, 3),  
    age_last_visit = round(as.numeric(last_visit - birthdt)/365, 3),  
    #traditional definition used by ACT for follow-up time for cases and controls
    age_onset = round(as.numeric(onsetdate - birthdt)/365, 3),
    age_act = round(ifelse(!is.na(age_onset), age_onset, age_last_visit), 3),
    #calculate birth cohort every 10 yrs since only 1 person born btwn 1890-1894 & this causes a model error
    birth_cohort = round(as.numeric(format(birthdt, "%Y")), -1),
    birth_cohort = ifelse(birth_cohort >= min(birth_cohort) & 
                            birth_cohort < min(birth_cohort) + 20, min(birth_cohort),
                          ifelse(
                            birth_cohort >= min(birth_cohort) + 20 & 
                            birth_cohort < min(birth_cohort) + 20*2, min(birth_cohort) + 20,
                            min(birth_cohort) + 20*2
                            )
                          ),
    # update anydementia. Unless someone was suspected of dementia and was evaluated for it, the ANYDEMENTIA (as well as DSMIVDX, ANYAD, NINDX, etc.) field will be blank. Think of ‘0’ as a confirmation of no dementia, while blank as a presumption of no dementia.
    anydementia = ifelse(is.na(anydementia), 0, anydementia),
    dsmivdx = ifelse(is.na(dsmivdx), 0, dsmivdx),
    #dsmivdx = factor(dsmivdx, levels = c(0:6), labels = c("no dementia", "AD", "Vascular", "Other medical", "Substance", "Multiple etiologies", "Other or unknown cause")),
    nindx = ifelse(is.na(nindx), 0, nindx),
    #nindx = factor(nindx, levels = c(0:3), labels = c("no dementia", "probable AD", "possible AD", "Dementia, type unknown")),
     #Time-varying covariates
     ##starting age is age on Jan 1st of a given year if participant was enrolled, or on intake date, whichever came later
    age_start_exposure = round(ifelse(format(intakedt, "%Y") == exposure_year,
                                as.numeric(intakedt-birthdt)/365,
                                as.numeric(as.Date(paste0(exposure_year, "-01-01")) - birthdt)/365), 3),
    ##ending age is age at end of year or last visit date, whichever came first 
    age_end_exposure = round(ifelse(format(last_visit, "%Y") == exposure_year,
                                as.numeric(last_visit - birthdt)/365,
                                as.numeric(as.Date(paste0(exposure_year, "-12-31")) - birthdt)/365), 3),
    ### need to change age_last_visit to age_act for sensitivity analyses
    dementia_now = ifelse(age_end_exposure < age_last_visit, 0, anydementia),
    ad_nincds = ifelse(nindx %in% c(1,2), 1, 0),
    ad_now = ifelse(age_end_exposure < age_last_visit, 0, ad_nincds),
    
    bmi4 = factor(bmi4, levels = c(0:3), labels = c("underweight", "normal", "overweight", "obese")),
    ## filter out invalid casi scores 
    casi_irt = ifelse(casi_valid ==1, casi_irt, NA),
    income_cat = ifelse(tr_med_inc_hshld < 20000, 1,
                               ifelse(tr_med_inc_hshld >= 20000 & tr_med_inc_hshld < 35000, 2,
                                      ifelse(tr_med_inc_hshld >= 35000 & tr_med_inc_hshld < 50000, 3,
                                             ifelse(tr_med_inc_hshld >= 50000 & tr_med_inc_hshld < 75000, 4, 5)
                                             ))),
    race_white = ifelse(race==1, 1, 0)
    ) %>%
  # group_by(study_id, pollutant) %>%
  # mutate(
  #   #calculate how long someone has been followed for
  #   enrollment_yr = seq(1:n())
  #   ) %>%
  select(
    study_id,
    birthdt, birth_cohort, intakedt, onsetdate, last_visit, 
    age_intake, age_act, age_last_visit, 
    
    anydementia, dsmivdx, nindx, ad_nincds,
    dementia_now, ad_now,
    
    age_start_exposure, age_end_exposure,
    #enrollment_yr,
    exposure_year, pollutant, exp_avg01_yr:exp_mos_coverage20_yr,
    
    apoe:degree, tr_med_inc_hshld, income_cat,race, race_white, smoke:bmi4, casi_irt, 
  ) 
 
#how long have participants been followed? 
yr <- dem0.1 %>%
  group_by(study_id) %>%
  mutate(
    intake_yr = as.numeric(min(format(intakedt, "%Y"))),
    last_visit_yr = as.numeric(min(format(last_visit, "%Y"))),
    fu_yrs = last_visit_yr - intake_yr +1
    ) %>%
  select(
    study_id, intake_yr, last_visit_yr, fu_yrs #yr_n  #pollutant
  ) %>%
  unique()

# 1 row per follow-up/exposure year
enrollment_yrs <-rep(yr$study_id, yr$fu_yrs) %>%
  as.data.frame() %>%
  rename(study_id = ".") %>%
  left_join(yr) %>%
  group_by(study_id) %>%
  mutate(
    enrollment_yr = seq(1:min(fu_yrs)),
    exposure_year = seq(from=min(intake_yr), to=min(last_visit_yr))) %>%
  select(study_id, enrollment_yr, exposure_year) %>%
  as.data.frame()

#data that can be used for all analyses 
dem0.1 <-  dem0.1 %>% 
  #keep all enrollment years
  right_join(enrollment_yrs) %>%
  select(
    study_id:age_end_exposure,
    enrollment_yr,
    exposure_year:casi_irt
  )

#rm(dem0.0)


# #proportion of observations with at least 95% of months in 10-year period; for NO2 and PM2.5
# prop.table(table(dem0.1$exp_mos_coverage10_yr>0.95)) %>% round(2)

#data for primary and some secondary analyses 
dem1 <- dem0.1 %>% 
  filter(
    #drop exposure predictions after the last visit when individuals are no longer "at risk" of the terminating event. 
    ## --> note: when using act_age, have to change last_visit to onsetdate
    exposure_year <= as.numeric(format(last_visit, "%Y")),
  ) %>%
  mutate(
     #only keep predictions w/ at least x% coverage
    exp_avg01_yr = ifelse(exp_mos_coverage01_yr >= min.pct, exp_avg01_yr, NA),
    exp_avg05_yr = ifelse(exp_mos_coverage05_yr >= min.pct, exp_avg05_yr, NA),
    exp_avg10_yr = ifelse(exp_mos_coverage10_yr >= min.pct, exp_avg10_yr, NA),
    exp_avg20_yr = ifelse(exp_mos_coverage20_yr >= min.pct, exp_avg10_yr, NA),
    exp_avg10_yr10yrlag = ifelse(exp_mos_coverage10_yr10yrlag >= min.pct, exp_avg10_yr10yrlag, NA),
    exp_avg10_yr20yrlag = ifelse(exp_mos_coverage10_yr20yrlag >= min.pct, exp_avg10_yr20yrlag, NA),
    
    #data used in models
    in_m2 = (pollutant == "no2" & !is.na(exp_avg10_yr) & !is.na(male) & !is.na(degree) & !is.na(race_white) & !is.na(income_cat) & !is.na(birth_cohort) & !is.na(apoe))
    ) # %>% select(in_m2, pollutant, exp_avg01_yr, male, degree, race_white, income_cat, birth_cohort, apoe) 
 



#make wide format
no2 <- dem1 %>% 
  filter(pollutant == "no2") %>%
  select(-pollutant) %>%
  rename(
    no2_1yr = exp_avg01_yr,
    no2_5yr = exp_avg05_yr,
    no2_10yr = exp_avg10_yr,
    no2_20yr = exp_avg20_yr,
    no2_10yr10yrlag = exp_avg10_yr10yrlag,
    no2_10yr20yrlag = exp_avg10_yr20yrlag,
    no2_coverage_1yr = exp_mos_coverage01_yr,
    no2_coverage_5yr = exp_mos_coverage05_yr,
    no2_coverage_10yr = exp_mos_coverage10_yr,
    no2_coverage_20yr = exp_mos_coverage20_yr,
    no2_coverage_10yr10yrlag = exp_mos_coverage10_yr10yrlag,
    no2_coverage_10yr20yrlag = exp_mos_coverage10_yr20yrlag
    ) 

pm25 <- dem1 %>% 
  filter(pollutant == "pm25") %>%
  select(-pollutant) %>%
  rename(
    pm25_1yr = exp_avg01_yr,
    pm25_5yr = exp_avg05_yr,
    pm25_10yr = exp_avg10_yr,
    pm25_20yr = exp_avg20_yr,
    pm25_10yr10yrlag = exp_avg10_yr10yrlag,
    pm25_10yr20yrlag = exp_avg10_yr20yrlag,
    pm25_coverage_1yr = exp_mos_coverage01_yr,
    pm25_coverage_5yr = exp_mos_coverage05_yr,
    pm25_coverage_10yr = exp_mos_coverage10_yr,
    pm25_coverage_20yr = exp_mos_coverage20_yr,
    pm25_coverage_10yr10yrlag = exp_mos_coverage10_yr10yrlag,
    pm25_coverage_10yr20yrlag = exp_mos_coverage10_yr20yrlag
    )   
dem.w <- full_join(no2, pm25)

```

```{r}
fixed.covariates <- dem.w %>%
  select(
    # study_id, birth_cohort,
    # age_intake, age_act, age_last_visit,
    # anydementia, dsmivdx, nindx, ad_nincds,
    study_id:ad_nincds,
    apoe:casi_irt
    ) %>% 
  #make all variables numeric; code has values starting at 0, whereas as.numeric() starts at 1, subtract 1 to fix
  mutate(bmi4 = as.numeric(bmi4)-1,
         # dsmivdx = as.numeric(dsmivdx)-1,
         # nindx = as.numeric(nindx)-1
         birth_yr = as.numeric(format(birthdt, "%Y")),
         intake_yr = as.numeric(format(intakedt, "%Y")),
         onset_yr = as.numeric(format(onsetdate, "%Y")),
         last_visit_yr = as.numeric(format(last_visit, "%Y"))
         ) %>%
  #drop dates, which are not "numeric" 
  select(
    -birthdt, -intakedt, -onsetdate, -last_visit
  ) %>%
  #one row per individual
  unique() 

```


```{r}
### --> ?? why is there one AD case (NINCDS) w/o dementia (DSM-IV)? different criteria used?
#table(dem=dem$dementia_now, ad=dem$ad_now)

```


# Quality Control

Make sure age variables make sense.
### --> check any_dementia now. some ppl have > 1

```{r, echo=T}

table(round(dem.w$age_intake, 1) <= round(dem.w$age_act, 1))  
#View(dem.w[which(dem.w$age_intake > dem.w$age_act),]) 
table(dem.w$age_intake <= dem.w$age_last_visit)
#View(dem.w[which(dem.w$age_intake > dem.w$age_last_visit),]) #study_id: 4647 has wrong last_visit/intakedt? last_visit is correct. "Will be updated in next freeze"
table(dem.w$age_act <= dem.w$age_last_visit)

#dementia 
dementia_counts <- dem.w %>% 
  group_by(study_id) %>%
  dplyr::summarize(
    #study_id = mean(study_id),
    #number of times each patient has dementia_now==1 (should be max 1)
    anydementia = mean(anydementia),
    dem_now_n = sum(dementia_now==1)
  ) 

##check that individuals who have anydementia=1 also have dementia_now=1 & those who don't don't  
dementia_counts %>%
  with(., table(anydementia, dem_now_n))

##check that individuals only have 1 dementia_now
dementia_counts %>%
  with(., table(dem_now_n <= 1))
#View(dem.w[which(dem.w$age_intake > dem.w$age_last_visit),]) 
#study_id 1027 has 2 dementia_now's b/c last visit age vs end of year age, which rounded to same thing
#View(dem.w[dem.w$study_id==1027,]) 

#AD 
ad_counts <- dem.w %>% 
  group_by(study_id) %>%
  dplyr::summarize(
    ad_nincds = mean(ad_nincds),
    ad_now_n = sum(ad_now==1)
  ) 
##check that individuals who have ad_nincds=1 also have ad_now=1 & those who don't don't #looks good
ad_counts %>%
  with(., table(ad_nincds, ad_now_n))

##check that individuals only have 1 ad_now
ad_counts %>%
  with(., table(ad_now_n <= 1))


```

Dementia   

```{r}
#number of individuals w/ dementia incidence: ~1270 
dem.w %>% group_by(study_id) %>%
  dplyr::summarize(
  dem = mean(anydementia)
) %>% 
  with(., table(dem))

#enrolled participants & dementia incidence over time
dem.w %>%
  mutate(dementia_now = factor(dementia_now, levels = c(1,0))) %>%
  ggplot(aes(x= exposure_year, 
             fill= dementia_now
             #fill=dsmivdx
             )) + 
  geom_bar() + 
  labs(title = "enrolled participants",
       fill = "dementia incidence"
       ) #+ 
  #scale_y_log10()


#dementia incidence over time
dem.w %>%
  filter(dementia_now==1) %>%
  ggplot(aes(x= exposure_year,  fill=dsmivdx)) + 
  geom_bar() + 
  labs(title= "number of dementia incident cases over time")

#follow-up time
##new id by intake date  
new_id <- dem.w %>%
  arrange(intakedt) %>%
  select(study_id) %>%
  unique() %>%
  mutate(id = row_number())

dem.w %>% 
  #filter(study_id < 1100) %>%
  left_join(new_id) %>%
  mutate(dementia_now = factor(dementia_now)) %>%
  ggplot(aes(x=exposure_year, y=id, group=id, colour=dementia_now)) + 
  geom_line() +
  geom_point(aes())
  #geom_point(data=dem.w[dem.w$dementia_now==1,], aes(x=exposure_year, y=id))

```

# Descriptive Statistics 

## Distribution of key covariates.

```{r}

#histograms
fixed.covariates %>% 
  gather(key = "variable",  value = "value", -study_id) %>%  
  ggplot(aes(x=value)) + 
  geom_histogram() +
    facet_wrap(~variable, scales = "free")

```

Individuals in each birth Cohort and birth year. Some 5-year cohorts have few only 1 individual

```{r}
#birth cohort/year
fixed.covariates %>%
  gather(key = "variable",  value = "value", -study_id) %>%  
  filter(variable %in% c("birth_yr", "birth_cohort")) %>%
  #ggplot(aes(x=birth_cohort, group=birth_cohort)) + 
  #geom_histogram()
  ggplot(aes(x=value)) + 
  geom_histogram() +
    facet_wrap(~variable, scales = "free")

#table(fixed.covariates$birth_yr)
table(fixed.covariates$birth_cohort)

```


## Missingness of key covariates. 

Data is for all individuals with NO2 and/or PM2.5 predictions.    
-lot of individuals with missing apoe readings    

```{r}

fixed.covariates %>% summaryFull(digit.type = "round")  
#describe()  #Info: how continuous the variable is, Gmd: Gini's mean difference - measure of dispersion

```


## Baseline Statistics

Statistics calculated for individuals for whom we know at least `r min.pct*100`% of their past residential address (e.g., 114/120 months for 10-year avg).

Median NO2 (ppb) and PM2.5 (µg/m3) concentration at baseline.

```{r baseline characteristics}
#participants in M2 at any time point
m2.participants <- dem1 %>%
  filter(in_m2 == TRUE) %>%
  select(study_id) %>%
  unique()  
m2.participants <- m2.participants$study_id

dem.bsl <- dem1 %>%
  filter(
    #1st year of enrollment (baseline)
    enrollment_yr == 1
    ) %>%
  #drop unnecessary rows
  select(-c(exp_avg01_yr, exp_avg05_yr, exp_avg10_yr10yrlag:exp_avg20_yr,
            exp_mos_coverage01_yr:exp_mos_coverage20_yr)) %>%
  mutate(
    #in_m2 as long as Any rows (years) used (thus this count is greater than in_m2 variable)
    m2_id = study_id %in% m2.participants
  )
#table(in_m2=dem.bsl$in_m2, in_m2.2=dem.bsl$m2_id)  
#View(dem.bsl[c("study_id", "in_m2", "m2_id")])


#calculate enrollment pollutant levels
bsl.expo.distrib <- dem.bsl %>% 
  group_by(pollutant) %>%
  drop_na(exp_avg10_yr) %>%
  dplyr::summarize(
    N = n(),
    median = median(exp_avg10_yr),
    iqr = IQR(exp_avg10_yr),
    mean = mean(exp_avg10_yr),
    sd = sd(exp_avg10_yr),
    min = min(exp_avg10_yr),
    max = max(exp_avg10_yr)
  )

kable(bsl.expo.distrib, digits = 0, caption = "")

median.bsl.no2 <- bsl.expo.distrib %>%
  filter(pollutant == "no2") %>%
  select(median) %>%
  as.numeric() %>% round()

median.bsl.pm25 <- bsl.expo.distrib %>%
  filter(pollutant == "pm25") %>%
  select(median) %>%
  as.numeric() %>% round()

#variable for whether baseline concentrations are above/below median concentration
dem.bsl <- dem.bsl %>%
  mutate(
    no2_median = ifelse(pollutant == "no2" & exp_avg10_yr >= median.bsl.no2, "above", 
                         ifelse(pollutant == "no2" & exp_avg10_yr < median.bsl.no2,"below", NA)),
    pm25_median = ifelse(pollutant == "pm25" & exp_avg10_yr >= median.bsl.pm25, "above", 
                         ifelse(pollutant == "pm25" & exp_avg10_yr < median.bsl.pm25,"below", NA))
    )

```

Participant baseline characteristics. Percents are calculated based on the number of individuals with available data (missing values removed). Some participants did not have baseline exposure estimates and are thus excluded from above/below median columns. 

```{r}
#Median "NA" column is for participants without exposure predictions for their enrollment year. 

```

 
```{r}
#characteristics for entire cohort 
t1.all <- dem.bsl %>% 
#1 row per participant
  filter(pollutant == "no2") %>%
  unique() %>%
  t1.fn()

kable(t1.all, caption = "Cohort Baseline Characteristics")

#characteristics by pollutant and concentration level
## no2
t1.no2.all <- dem.bsl %>%
  filter(pollutant == "no2") %>%
  #drop ppl w/o  exposure estimates @ their baseline
  drop_na(exp_avg10_yr) %>%
  t1.fn(data = ., column.name = "no2 all")

t1.no2.below <- dem.bsl %>%
  filter(pollutant == "no2",
         no2_median == "below") %>%
  t1.fn(data = ., column.name = "no2 below")

t1.no2.above <- dem.bsl %>%
  filter(pollutant == "no2",
         no2_median == "above") %>%
  t1.fn(data = ., column.name = "no2 above")

t1.no2.na <- dem.bsl %>%
  filter(pollutant == "no2",
         is.na(no2_median) ) %>%
  t1.fn(data = ., column.name = "no2 NA")

t1.no2 <- cbind(t1.no2.all, 
                t1.no2.below, 
                t1.no2.above, 
                t1.no2.na)

## pm2.5
t1.pm25.all <- dem.bsl %>%
  filter(pollutant == "pm25") %>%
  drop_na(exp_avg10_yr) %>%
  t1.fn(data = ., column.name = "pm25 all")

t1.pm25.below <- dem.bsl %>%
  filter(pollutant == "pm25",
         pm25_median == "below") %>%
  t1.fn(data = ., column.name = "pm25 below")

t1.pm25.above <- dem.bsl %>%
  filter(pollutant == "pm25",
         pm25_median == "above") %>%
  t1.fn(data = ., column.name = "pm25 above")

t1.pm25.na <- dem.bsl %>%
  filter(pollutant == "pm25",
         is.na(pm25_median) ) %>%
  t1.fn(data = ., column.name = "pm25 NA")

t1.pm25 <- cbind(t1.pm25.all, 
                t1.pm25.below, 
                t1.pm25.above, 
                t1.pm25.na)

t1.by.pollutant <- cbind(t1.no2, t1.pm25)

kable(t1.by.pollutant, caption = "Cohort Characteristics by 10 yr avg Air Pollution Exposure Estimates Upon Enrollment")

#characteristics by dementia status
t1.dem.case <- dem.bsl %>%
  #1 row per participant
  filter(pollutant == "no2") %>%
  filter(anydementia == 1) %>%
  t1.fn(data = ., column.name = "Dementia")

t1.dem.noncase <- dem.bsl %>%
  #1 row per participant
  filter(pollutant == "no2") %>%
  filter(anydementia == 0) %>%
  t1.fn(data = ., column.name = "No Dementia")

t1.by.anydementia <- cbind(t1.dem.case, t1.dem.noncase)

kable(t1.by.anydementia, caption = "Cohort Characteristics by Dementia Outcome")

```

# Inferential Statistics

```{r}
#3. create survival object w/ TVC dataset (accounts for "delayed entry")

#individuals w/ only baseline observations are dropped/NAs created
s.dem <- with(dem.w, Surv(
  # --> update so that time < time2
  time = age_start_exposure,  
  time2 =age_end_exposure,  
  event = dementia_now
                         )
             )

```

## Primary Analysis

Model 2 (a priori): M1 + gender, education, median household income, race, calendar year 2 ; APOE stratification.

-assuming missing values (e.g., APOE) are MCAR and doing a complete case analysis. This method can be bias if values are not MCAR.

##### --> categorize birth cohort into larger categories - 20 yrs? 
##### --> create table w/ HR results (see analysis plan)

```{r}
m2 <- dem.w %>%
  coxph(s.dem ~ no2_10yr + 
          male + factor(degree) + race_white + factor(income_cat) + factor(birth_cohort) + strata(apoe), 
        data=., 
        robust = T,
        #ties = "exact" #output is weird if use this
        ) 
m2 %>% summary()

```

## Sensitivity Analyses    
### Reduced & Extended Models    
Model 1 (Reduced): Age (time axis), NO2 (time-varying) 

```{r  }

# summary(coxph(s.stanford.tvc~trt, data=stanford.tvc))
m1 <- dem.w %>%
  coxph(s.dem ~ no2_10yr, 
        data=., 
        robust = T,
        ) 

m1 %>% summary()

```

M3 (extended): M2 + smoking + physical activity 

```{r}

m3 <- dem.w %>%
  coxph(s.dem ~ no2_10yr + 
          male + factor(degree) + race_white + factor(income_cat) + factor(birth_cohort) + strata(apoe) + 
          factor(smoke) + exercise_reg, 
        data=., 
        robust = T,
        ) 
m3 %>% summary()

```

Model 4 (Extended & mediation): M3 + hypertension, diabetes, CV summary, heart disease summary, BMI

```{r}
m4 <- dem.w %>%
  coxph(s.dem ~ no2_10yr + 
          male + factor(degree) + race_white + factor(income_cat) + factor(birth_cohort) + strata(apoe) + 
          factor(smoke) + exercise_reg +
          Hypertension + Diabetes + CV_DIS + Heart_Dis + bmi4, 
        data=., 
        robust = T,
        ) 
m4 %>% summary() 

```

Model 5 (APOE interaction): M2 + NO2*APOE

```{r}
m5 <- dem.w %>%
  coxph(s.dem ~ no2_10yr + 
          male + factor(degree) + race_white + factor(income_cat) + factor(birth_cohort) + strata(apoe) +
          no2_10yr*apoe , 
        data=., 
        robust = T,
        ) 
m5 %>% summary()

```

Model 6 (copollutant): M2 + PM2.5 (time-varying)

#### [error b/c so much missing ?income_cat data; & few ppl in each birth category]
#### only "1 event"

```{r}

m6 <- dem.w %>%
  coxph(s.dem ~ no2_10yr +
          male + factor(degree) + race_white  + factor(birth_cohort) + strata(apoe) +
          pm25_10yr, #+ factor(income_cat)
        data=.,
        robust = T,
        )
m6 %>% summary()



# get error above b/c income_cat has so much missing data/ birth cohort categories are small?
# dem.w %>%
#   filter(
#     !is.na(no2_10yr),
#     !is.na(male),
#     !is.na(degree),
#     !is.na(race_white),
#     !is.na(income_cat),
#     !is.na(birth_cohort),
#     !is.na(apoe),
#     !is.na(pm25_10yr),
#   )

```

### Different Exposure Windows 

```{r}


```

### AD outcome 

```{r}

```

### Redefine dementia onset to match ACT study definition

```{r}
##when using ONSET AGE (age_act), need to drop exposure predictions after age_act. filter(age_at_exposure <= age_act) & dementia_now = ifelse(age_at_exposure < age_act, 0, anydementia),

```

### Cox model assumptions and inference validity    
1. Proportional hazards Assumption    
a) Schoenfeld residuals. If the proportional hazards model is correctly specified, there should be no trend in the residuals over time (uncorrelated, with mean zero across event times)

```{r}
## ?? what are the values in each vector/model variable? 
schoenresid = residuals(m2, type = "scaledsch")
## ?? is this failure time? 
time = as.numeric(rownames(schoenresid))

# link no2_10yr & "times" column
cbind(resid=schoenresid[,1], time) %>%
  as.data.frame() %>%
  ggplot(aes(x=time, y=resid)) + 
  geom_point() + 
  geom_smooth() + 
  labs(y = "scaled Schoenfeld residuals",
       x = "failure time (age at dementia incidence)")

```

b. formally test. using cox.zph(). if global p-val is significant, assumption is violated   
-assumption is violated. ?? are median income and exposure year the reason why? 

```{r}
cox.zph(m2)
```

2. linear relationship between continuous covariates and log hazard of the outcome    
### --> use survminer::ggcoxfunctiona()

```{r}
##??
martingalesid = residuals(m2, type = "martingale")
## ? error, diff lengths??
#cbind(resid=martingalesid, no2_10yr=dem.w$no2_10yr) #%>%
  # as.data.frame() %>%
  # ggplot(aes(x=no2_10yr, y=resid)) + 
  # geom_point() + 
  # geom_smooth() + 
  # labs(y = "Martingale residuals"
  #     )

```

### --> $ \triangle \beta $ doesn't work  
3. no outliers or influential observations, $ \triangle \beta $

### --> calculate standardized delta Bs?    
 
```{r}
# B537 L3, slide 125

# ? don't have to fit M2 again w/ subset=ind?
dfbetas <- residuals(m2, type="dfbeta") 

#rename columns
colnames(dfbetas) <- names(coef(m2))

dfbetas %>%  
  as.data.frame() %>%
  #get rows used in model
  rownames_to_column(var = "ind") %>%
  mutate(ind = as.numeric(ind)) %>%
  #make long format
  gather(covariate, value, -ind) %>%
  as.data.frame() %>%
  #plot
  ggplot(aes(x=ind, y=value)) +
  geom_point() +
  facet_wrap(~covariate, scales="free") + 
  labs(x="observation index",
       y= "delta beta")

```



### Restrict Analysis to individuals for whom we have high quality data

```{r}



```

### Adjust for baseline CASI score 

```{r}

```

### Adjust for ACT cohort 

```{r}

```

### IPW for missing APOE

-APOE is more likely to be missing if: 
--predictor variables: ppl born later in time, later intake year, have CV_dis, diabetes, ??low/high education, less exercise, non-whites, ?current smoker   
--outcomes: AD, demenentia, younger during last visit or onset date, later onset year

```{r}
# see B540 L7-8

apoe.l <- fixed.covariates %>%
  mutate(
    apoe_available = !is.na(apoe)
  ) %>% 
  select(-apoe) %>%
  gather("variable", "value", -study_id, -apoe_available) %>%
  mutate(
    group = group_indices(., variable)
  )

#MAR: are missing APOE values assoc w/ other measured variables, including calendar time? 
#NMAR: are missing APOE associated w/ dementia or AD incidence that we don't obseve...? 

apoe.l %>%
  #only look at first few variables
  filter(group %in% c(1:15)) %>%
  ggplot(aes(x=value, fill= apoe_available)) + 
  #geom_histogram(aes(y = stat(density)), position = "dodge") +
  geom_density(alpha=0.3) +  #stat = "density"
  facet_wrap(~variable, scales="free") #+ 
  #scale_y_continuous(labels = percent_format()) 

apoe.l %>%
  #only look at first few variables
  filter(group %in% c(16:max(group))) %>%
  ggplot(aes(x=value, fill= apoe_available)) + 
  #geom_histogram(aes(y = stat(density)), position = "dodge") +
  geom_density(alpha=0.3) +  #stat = "density"
  facet_wrap(~variable, scales="free") #+ 

# ? run models? 
## make wide format for models
apoe.w <- apoe.l %>%
  select(-group) %>%
  spread(variable, value)

```

```{r old model using manually selected predictor variables}
# -predictor variables: ppl born later in time, later intake year, have CV_dis, diabetes, ??low/high education, less exercise, non-whites, ?current smoker   
# --outcomes: AD, demenentia, younger during last visit or onset date, later onset year

# apoe.w %>%
#   glm(apoe_available ~ age_last_visit + birth_cohort + age_intake + intake_yr +
#         CV_DIS + Diabetes + factor(degree) + exercise_reg + race_white + factor(smoke) + 
#         anydementia
#         ,
#   family = "binomial",
#   data = .
#   ) %>%
#   summary()

# are missing APOE values associated w/ AP (above/below median)? 
# ? Is it a confounder...? 

```

```{r}
# IPW with coxph()
###1. use selection model (e.g. lasso for glm) to select predictors of whether or now we have APOE values for people: apoe_available (in apoe.w)
### 2. fit logistic model to determine probability of observing an observation given predictors: glm(apoe_available ~ [predictors])
### 3. use model w/ dataset to get predicted values (probability of being observed), 
### 3b. ? repeate for other variables w/ a lot of missing data (e.g. income?) to get probability of being observed; multiply by apoe probability to get overall probability that both variables were observed
### 4. take the inverse of probabilities to calculate "weights", use these weights in coxph(, weights = c()) 
#### ?? any issues w/ other, non-APOE variables being missing? 



```


```{r}
# optional, could repeate and use weights based on the probability that All of the M2 predictors (not just APOE) are observed (e.g., all_covs_available ~ [predictors])

```




```{r}

```


### Optional

1. finer adjustments for confounders
2. use a continuous adjustment for education. And other variables?

```{r}


```



```{r lincom}
# estimating SE & CIs for linear combinations of coefficients
# m2c <- coxph(surv.addicts ~ dose*prison + strata(clinic), data=addicts)
# #summary(m2c)
# 
# est <- as.numeric(exp(60*coef(m2c)["dose"] + 1*coef(m2c)["prison"] + 100*coef(m2c)["dose:prison"]))
# 
# se <- deltamethod(g = ~exp(60*x1 + x2 + 100*x3),
#                   mean = coef(m2c)[c("dose", "prison", "dose:prison")],
#                   cov= vcov(m2c)[c("dose", "prison", "dose:prison"), 
#                                  c("dose", "prison", "dose:prison")],
#                   ses = T) 
# 
# ci = as.numeric(c(est-se*1.96, est+ se*1.96))

```

