---
title: 'Aim 2: QAQC - Instrument Readings'
author: "Magali Blanco"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
editor_options:
  chunk_output_type: console
---

```{r}
# ---> TO DO:
# ---> ? how to drop very low/high UFP & BC readings?
### --> keep DiscMini

```

```{r, echo=F}
# notes
#BC = 880 nm (IR); 625 nm (Red); 528 nm (Green); 470 nm (Blue); and 375 nm (UV);  

```

# Summary of Script

**All data**   

* keep only BC and PTRAK readings, current MM stops (drop old stops)
* relabeled incorrectly labeled instruments
* split up data into UFP & BC datasets 

**UFP** 

* drop erroneous UFP instrument readings due to wick/other PTRAK issues
	+ < ~300 data points were dropped from 2 periods where PTRAKs appeared to have failed - concentrations were at or near zero and very different than concentrations immediately before/after
	+ these readings did not correlate well to measures from other particle measurement instruments

**BC** 

* perform ONA correction   
* estimate BC from 880 nm (IR) wavelength 

**All data** 

* compare co-located instruments to make sure readings are similar
  + PTRAK - PTRAK   
	+ NanoScan - NanoScan    
	+ PTRAK - NanoScan (< ptrak limit)   
    - used NanoScan particle counts comparable to PTRAK (subtracted counts for particles < 20 nm)   

# Analyses  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=T, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE,
                      fig.height = 5, fig.width = 10
                      )  

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

# load packages 
pacman::p_load(tidyverse, 
               # tables
               knitr, kableExtra,
               # dates
               chron, lubridate,
               # statistics
               Hmisc
               )   

# source global variables & functions
source("0.Global_Fns.R")
source(file.path("A2.0.1_Var&Fns.R"))

images_path <- file.path(images_path0, "1. Stops")
tables_path <- file.path(tables_path0, "1. Stops")

#read in data
mm_full0 <- readRDS(file.path("Data", "Aim 2", "Mobile Monitoring", "all_data_2020-03-08.rda"))

### --> why do these runs have missing variable/value? 2019-05-14_R03, 2019-07-17_R07	

site_ids <- read.csv(file.path("Data", "Aim 2", "Mobile Monitoring", "locations_190715.csv")) %>%
  select(site_id) %>%
  #drop Roosevelt garage stop; drop site w/ only 1 observation - stopped sampling here & added MS0601
  filter(!site_id %in% c("MS0000", "MS0398")) %>%
  unique()

mm_full <- mm_full0 %>%
  # --> why do 2 runs have missing values??
  select(-duration_sec) %>%
  drop_na(value) %>%
  filter(!variable %in% c("ufp_disc_ct_cm3",
                         "ufp_disc_med_size_nm", 
                         "ufp_pt_screen_ct_cm3"
                         )) %>%
  mutate(variable = recode_factor(factor(variable),  
                              "ufp_pt_noscreen_ct_cm3" = "ptrak_pt_cm3")) %>%
  # not sure why there are repeated identical rows for ufp_scan_11_5_ct_cm3
  unique()

#unique(mm_full$variable)

#drop old stops 
mm_full <- left_join(site_ids, mm_full)

#relabel instruments that were incorrectly labeled
mm_full <- mm_full %>%
  mutate(instrument_id = ifelse(instrument_id == "BC_0063", "BC_63",
                                #ifelse(instrument_id == "CO_2", "CO_3",
                                       ifelse(instrument_id == "PMSCAN_1", "PMSCAN_3",
                                              instrument_id))
         )

#Create NanoScan counts (20-420 nm) similar to ptraks: 20-1000 nm)
ns <- mm_full %>%
  #look at variables from NanoScan
  filter(grepl("scan", variable)) %>%
  #not sure why there are repeated identical rows for ufp_scan_11_5_ct_cm3
  unique() %>%
  #make wide format to calculate the difference
  spread(variable, value) %>%
  # dropping ufp_scan_20_5_ct_cm3 counts (17.8-23.7 nm) since PTRAK-NS comparison looks slightly better without these counts.
  mutate(scan_20_420_pt_cm3 = ufp_scan_ct_cm3 - ufp_scan_11_5_ct_cm3 - ufp_scan_15_4_ct_cm3 - ufp_scan_20_5_ct_cm3
         ) %>%
  select(-c(ufp_scan_11_5_ct_cm3, # 10.0-13.3 nm
            ufp_scan_15_4_ct_cm3, # 13.3-17.8 nm
            ufp_scan_20_5_ct_cm3, # 17.8-23.7 nm
            ufp_scan_ct_cm3)) %>%
  #dplyr::rename(scan_10_420_pt_cm3 = ufp_scan_ct_cm3) #%>%
  #make back to long format
  gather("variable", "value", scan_20_420_pt_cm3)

# replace old nanoscan calculations w/ new ones similar to PTRAK
mm_full <- mm_full %>%
  filter(!grepl("scan", variable)) %>%
  rbind(ns) %>%
  arrange(time)   

# create new temporal variables, arrival time and date
mm_full <- mm_full %>%
  dplyr::group_by(runname, site_id) %>%
  #set time to arrival time
  dplyr::mutate(arrival_time = min(time)) %>%
  mutate(date = as.Date(substr(arrival_time, 1,10))) %>%
  ungroup()

```

Instruments Used

```{r}
#instrument used each day
mm_full %>%
  select(instrument_id, time) %>%
  mutate(date = date(time)) %>%
  unique() %>%
  ggplot(aes(x=date, y=instrument_id, colour = instrument_id)) + 
  geom_point(alpha=0.5) + 
  theme(legend.position = "none") + 
  labs(title = "Instruments used",
       x = "Date",
       y = "Instrument ID"
       ) 
 
```

Set up common variables 

```{r, echo=T}

# site note: distribution of pollutants
# summary(mm_full$value[mm_full$variable=="bc_ng_m3"])
# summary(mm_full$value[mm_full$variable=="ptrak_pt_cm3"])


max_bc_limit <- 1e6 #ng/m3
bc_factor <-10 

ptrak_lim <- 5e5 
low_ufp_lvl <- 300

```

# Correct BC readings suing the Optimized Noise-Reduction Algorithm (ONA) 

Distribution of raw BC readings. See that both instruments have negative BC readings.  

```{r}

mm_full %>% 
  filter(grepl("ma200_ir_bc", variable)) %>%
  group_by(instrument_id) %>%
distribution.table(., var.string = "value") %>%
  kable(caption = "Distribution of raw 10-sec Aethelometer readings") %>%
  kable_styling()

```

```{r}
# this does not need to be done separatebly by instrument, results are the same. May be related to different tape positions for differnet instruments, which is already taken into account?
bc.w <- mm_full %>%
  filter(grepl("ma200", variable)) %>%
  spread(variable, value) %>%
  ONA() 

bc <- bc.w %>%
  #only keep ONA-corrected values
  select(site_id:date, 
        bc_ng_m3 = ona_ir,
        #ma200_bc = ma200_ir_bc1
        ) %>%
  gather("variable", "value", contains("bc")) %>%
  # BC_63 has a few missing values during a 2.5-minute period
  drop_na(value)
  
```

Distribution of ONA-corrected BC readings. Still see some negative values, but there are fewer and they are not as negative. The highest values are a bit lower than the raw values.

```{r}

bc %>%  
  group_by(instrument_id) %>%
distribution.table(., var.string = "value") %>%
  kable(caption = "Distribution of ONA-corrected 10-sec Aethalometer readings") %>%
  kable_styling()

```

```{r}
#BC/IR
bc.w %>%
  ggplot(aes(x= ma200_ir_bc1, y = ona_ir)) + 
  geom_point() + 
  geom_abline(intercept = 0, slope = 1) + 
  geom_smooth() + 
  labs(title = "IR (880 nm, black carbon) concentrations before and after ONA correction",
      x = "Aethelometer (ng/m3)",
      y = "ONA-corrected (ng/m3)"
       )
```

replace old aethalometer readings with ONA corrected IR (880 nm) BC readings in full dataset

```{r}
mm_full <- mm_full %>%
  #drop old, raw MA readings
  filter(!grepl("ma200", variable)) %>%
  # attach ONA-corrected IR (880 nm) readings
  rbind(bc) %>%
  mutate(variable = droplevels(variable))

```

# Low instrument readings 

## UFP - PTRAKS 

very low concentrations: Elena (MOVUP Study) excluded UPF < 100 b/c: a) had very few samples that were less than 100 (less then 1%), b) they did not correlate to low measures on other particle measurement instruments, c) these values also resulted in extreme values on some the ratio metrics.

PMPT_93 and PMPT_94 have both had 0 readings. PMPT_94 has had a maximum reading of 500k pt/cm3. 

```{r}

mm_full %>% 
  filter(grepl("ptrak", variable)) %>%
  group_by(instrument_id) %>%
  #rename("Instrument ID" = instrument_id) %>%
distribution.table(., var.string = "value") %>%
  kable(caption = "summary of 1 sec PTRAK distribution (all data)") %>%
  kable_styling()

```

Time series plots of runs with very low instrument readings.

```{r}

low_ptrak <- mm_full %>%
  filter(grepl("ptrak", variable),
         value < low_ufp_lvl) %>%
  select(runname, site_id, instrument_id) %>% 
  unique()

```

Plot of entire day.

```{r, fig.height=8}

y_max <- 5000

mm_full %>%
  mutate(value = ifelse(grepl("bc", variable), value*bc_factor, value)) %>%
  filter(runname %in% low_ptrak$runname) %>%
  time_series_plots(., mytitle = paste0("time series for sites with PTRAK concentrations < ", low_ufp_lvl, " pt/cm3"),
                    hline_value = low_ufp_lvl, 
                    ymax = y_max) +
  labs(subtitle = paste0("only showing concentrations up to", y_max, " pt/cm3"))

```

Zoomed in to sites with low readings.        
* 3-22-2019: low alc noted. has unstable readings most of the stop     
* 7-13-2019 low alc noted. has clear wick issues in the middle and readings seem too stable immediately before and after.  
* 1-4-2020 has low readings at nearby stops MS0176 also has low BC readings. MS0177 has some low BC readings, but not all. readings seem too stable. Sign of wick issues?


```{r, fig.height=8}
df <- data.frame()

#only keep observations for sites with low readings 
for (i in 1:nrow(low_ptrak)) {
  one_site <- mm_full %>%
    filter(runname == low_ptrak$runname[i] &
             site_id == low_ptrak$site_id[i])
  
  df <- rbind(df, one_site)
}

df %>%
  mutate(value = ifelse(grepl("bc", variable), value*bc_factor, value)) %>%
  time_series_plots(., mytitle = paste0("time series for sites with PTRAK concentrations < ", low_ufp_lvl, " pt/cm3"),
                    hline_value = low_ufp_lvl) + 
  facet_wrap(~runname+site_id,
             scales = "free_x") 


```


```{r}
# # ?? need these? DELETE?

# Zoom in on 3-22 & 1/4 times w/ duplicat PTRAK readings. Take out site readings from bad PTRAK since these seem unstable the majority of the 2 minutes and we have other PTRAK readings.

# date1 <- "2019-03-22"
# 
# mm_full %>%
#   mutate(value = ifelse(grepl("bc", variable), value*bc_factor, value)) %>%
#   filter(grepl(date1, runname),
#          time > paste(date1, "17:05"),
#          time < paste(date1, "17:25")) %>%
#   time_series_plots(., mytitle = paste0("time series for runs with PTRAK concentrations < ", low_ufp_lvl, " pt/cm3"),
#                     hline_value = low_ufp_lvl)  

```

```{r}
# # ?? need these? DELETE?

# date2 <- "2020-01-04"
# 
# mm_full %>%
#   mutate(value = ifelse(grepl("bc", variable), value*bc_factor, value)) %>%
#   filter(grepl(date2, runname),
#          time > paste(date2, "06:05"),
#          time < paste(date2, "07:00")) %>%
#   time_series_plots(., mytitle = paste0("time series for runs with PTRAK concentrations < ", low_ufp_lvl, " pt/cm3"),
#                     hline_value = low_ufp_lvl)  

```
 
Total PTRAK readings before dropping any values: 

```{r}

total_ptrak_readings <- mm_full %>%
  filter(grepl("ptrak", variable)) %>%
  nrow()

total_ptrak_readings

```

Dropping sites obsrvations with low UFP readings (< `r low_ufp_lvl` pt/cm3) that may be indicative of:
* low alcohol wick issues   
* instrument error (MOV-UP study used 100 pt/cm3; Kerckhoffs et al. 2017 MM study & Kompmaker et al. 2015 used 500 pt/cm3 & cites other studies that used this criteria)

```{r}
# drop observations for sites with low readings 
for (i in 1:nrow(low_ptrak)) {
  mm_full <- mm_full %>%
    filter(!(runname == low_ptrak$runname[i] &
             site_id == low_ptrak$site_id[i] &
             instrument_id == low_ptrak$instrument_id[i]))
}

```

PTRAK readings dropped, n (%)

```{r}
remaining_ptrak_readings <- mm_full %>%
  filter(grepl("ptrak", variable)) %>%
  nrow()

ptrak_dropped_n <- total_ptrak_readings - remaining_ptrak_readings
ptrak_dropped_pct <- round(ptrak_dropped_n/total_ptrak_readings*100, 2)

paste0(ptrak_dropped_n, " (", ptrak_dropped_pct, "%)")

```

Distribution of remaining PTRAK readings.

```{r}

mm_full %>% 
  filter(grepl("ptrak", variable)) %>%
  group_by(instrument_id) %>%
  distribution.table(., var.string = "value") %>%
  kable(caption = "summary of 1 sec PTRAK distribution after dropping low instrument readings", 
         ) %>%
  kable_styling()
 
```

## BC 

Time series plots of runs with very low instrument readings.

```{r}
low_bc_quant <- 0.001
low_bc_lvl <- 0 #quantile(mm_full$value[mm_full$variable== "bc_ng_m3"], low_bc_quant)[[1]]

low_bc <- mm_full %>%
  filter(grepl("bc", variable),
         value <= low_bc_lvl) %>%
  select(runname, site_id, instrument_id) %>% 
  unique()

```

```{r}
# # SQL query for notes of low BC times
# low_bc_sql <- low_bc %>%
#   mutate(query = paste0("(runname = '", runname, "' AND upper(site_id) = '", site_id, "')")) %>%
#   dplyr::summarize(query = paste(query, collapse = " OR ")) %>%
#   as.character()

```


Plot of entire day.

```{r, fig.height=8}
y_max <- 3000

unique_low_bc_runs <- unique(low_bc$runname)  

bc_day_plots <- mm_full %>%
  mutate(value = ifelse(grepl("bc", variable), value*bc_factor, value)) %>%
  filter(runname %in% unique_low_bc_runs) %>%
  time_series_plots(., mytitle = paste0("time series for runs with BC concentrations < ", low_bc_lvl, " pt/cm3"),
                    hline_value = low_bc_lvl, 
                    ymax = y_max, 
                    facet_runname = F) + 
  labs(subtitle = paste0("only showing concentrations up to ", y_max, " pt/cm3"))


# plot on separate pages since there are many
# need this many pages if plot 9 plots/page
page_no <- ceiling(length(unique_low_bc_runs)/9)

for (i in seq(page_no)) {
  
  myplot <- bc_day_plots +
    ggforce::facet_wrap_paginate(~runname, 
                             nrow = 3, ncol = 3, 
                             page = i, 
                             scales = "free")  
  
  print(myplot)
}

```

zoom in on specific sites 

```{r, fig.height=8}
df <- data.frame()

#only keep observations for sites with low readings 
for (i in 1:nrow(low_bc)) {
  one_site <- mm_full %>%
    filter(runname == low_bc$runname[i] &
             site_id == low_bc$site_id[i])
  
  df <- rbind(df, one_site)
}

bc_site_plots <- df %>%
  mutate(value = ifelse(grepl("bc", variable), value*bc_factor, value)) %>%
  time_series_plots(., plot_bc_factor = bc_factor, 
                    mytitle = paste0("time series for sites with BC concentrations <= ", low_bc_lvl, " pt/cm3"),
                    hline_value = low_bc_lvl, 
                    ymax = y_max, 
                    facet_runname = F) + 
  labs(subtitle = paste0("only showing concentrations up to ", y_max, " pt/cm3"))

# plot on separate pages since there are many
# page_no <- ceiling(length(unique_low_bc_runs)/9)

for (i in seq(page_no)) {
  
  myplot <- bc_site_plots +
    ggforce::facet_wrap_paginate(~runname + site_id, 
                             nrow = 3, ncol = 4, 
                             page = i, 
                             scales = "free")  
  
  print(myplot)
}

```

total BC readings before dropping any values:

```{r}
total_bc_readings <- mm_full %>%
  filter(grepl("bc", variable)) %>%
  nrow()

total_bc_readings

```

The number of zero & negative (<= 0) 10-sec BC readings is very low. 

```{r}
# proportion of negative BC readings
negative_bc_no <- sum(mm_full$value[mm_full$variable == "bc_ng_m3"] <= 0) 
negative_bc_pct <- round(negative_bc_no/total_bc_readings*100, 2)

paste0(negative_bc_no, " (", negative_bc_pct, "%)")

```

### --> ?? drop BC < 0 since these readigns seem to generally be different than neighboring BC readings and diffrent from the patterns of other particle instruments? 

### --> [will email Elena] ?? Drop or keep values < 0 or < threshold (or set to ~ 0/min) since so few and don't know wha cutoff to use for high values? 

**Not doing anything for now since some of the very low values are in series (next to each other), indicating that readings are truly low?? Will averaging very low/high readings result in positive, unbiased readings in the end?**

```{r}
# [??] Set values < 0 equal to ___ the minimum positive BC reading divided by sqrt(2) ___
# 556 notes suggest: LOD/sqrt(2), but manual does not have a LOD for a 10-sec aethalometer reading. 
 
# low_bc_to_drop <- 0
# 
# min_positive_bc <- mm_full %>% 
#   filter(grepl("bc", variable),
#          value >0
#          ) %>%
#   dplyr::summarise(min_positive_bc = min(value)) %>% 
#   as.numeric()
# 
# mm_full <- mm_full %>%
#   #filter(value >0)
#   mutate(value = ifelse(value <= 0, min_positive_bc/sqrt(2), value))

```

Distribution of resulting BC readings 

```{r}
mm_full %>%  
  filter(grepl("bc", variable)) %>%
  group_by(instrument_id) %>%
distribution.table(., var.string = "value") %>%
  kable(caption = "Distribution of final 10-sec Aethalometer readings") %>%
  kable_styling()

```


# High instrument readings

**What others have done**   
*UFP*

### --> ? high conc for PTRAKS = 500 k ? too many high values when using quantile() to manually inspect


*BC*   
* MOVUP study: dropped BC readings > 27k ng/m3 (1% of data)   
* Kompmaker 2015: did minimal data cleaning b/c averaged 1-min readings to 30-mins   
* others have taken other approaches for dealing with noisy measurements and spike concentrations (e.g., Apte 2011 SI details dealing w/ spikes due to instrument jolts)

### --> high conc for BC is quantile?

### --> ? take 1 min rolling avg for PTRAK & plot time series? Can see when concentrations diverge

## High PTRAK readings 

```{r}

high_ptrak_val <- ptrak_lim #*.8 # or quantile?

#bc_factor <- 10

```

Runs with PTRAK readings at or above: `r high_ptrak_val` pt/cm3

```{r}

high_ptrak <- mm_full %>%
  filter(grepl("ptrak", variable),
         #value == ptrak_lim 
         value >= high_ptrak_val
         ) %>%
  select(runname, site_id) %>% 
  unique()

#high_ptrak <- high_ptrak$runname

```
 
Plot of entire day.

```{r}
# plot of entire day
mm_full %>%
  mutate(value = ifelse(variable == "bc_ng_m3", value*bc_factor, value)) %>%
  filter(runname %in% high_ptrak$runname) %>%
  time_series_plots(., plot_bc_factor = bc_factor, 
                    hline_value = ptrak_lim,
                    mytitle = "time series for runs with high PTRAK concentrations")

```

Zoom in on sites of interest. In general, elevated PTRAK readings are associated with delayed elevated NanoScan and BC readings. Thus, keep these high readings? 

### --> ? is the NS time stamp at the end of the reading period? 

```{r}
df <- data.frame()

#only keep observations for sites with high readings 
for (i in 1:nrow(high_ptrak)) {
  one_site <- mm_full %>%
    filter(runname == high_ptrak$runname[i] &
             site_id == high_ptrak$site_id[i])
  
  df <- rbind(df, one_site)
}

df %>%
  mutate(value = ifelse(variable == "bc_ng_m3", value*bc_factor, value)) %>%
  time_series_plots(., plot_bc_factor = bc_factor, 
                    hline_value = ptrak_lim,
                    mytitle = "time series for sites with high PTRAK concentrations") +
  facet_wrap(~runname+site_id, scales = "free") 


```

## High BC readings 

High BC reading quantile & concentration (ng/m3):  

```{r}
high_quant <- 0.99

bc_high_quant <- quantile(mm_full$value[mm_full$variable=="bc_ng_m3"], high_quant)

```

Runs with BC readings above quantile: `r bc_high_quant`

```{r}
high_bc <- mm_full %>%
  filter(grepl("bc", variable),
         value >= quantile(value, high_quant)) %>%
  select(runname, site_id) %>% 
  unique()

```

Plot of entire day.

```{r}
# plot of entire day
mm_full %>%
  mutate(value = ifelse(variable == "bc_ng_m3", value*bc_factor, value)) %>%
  filter(runname %in% high_bc$runname) %>%
  time_series_plots(., plot_bc_factor = bc_factor, 
                    hline_value = ptrak_lim,
                    mytitle = "time series for runs with high BC concentrations")

```

Zoom in on sites of interest.  

```{r}

df <- data.frame()

#only keep observations for sites with high readings 
for (i in 1:nrow(high_bc)) {
  one_site <- mm_full %>%
    filter(runname == high_bc$runname[i] &
             site_id == high_bc$site_id[i])
  
  df <- rbind(df, one_site)
}

df %>%
  mutate(value = ifelse(variable == "bc_ng_m3", value*bc_factor, value)) %>%
  time_series_plots(., plot_bc_factor = bc_factor, 
                    hline_value = ptrak_lim, 
                    mytitle = "time series for sites with high BC concentrations") +
  facet_wrap(~runname+site_id, scales = "free")

```


### --> plot where have divergent estimates. e.g. very high NS readings. near airport? is this where discMini has high readings? 

```{r}

```



# Compare co-located instruments 

## UFP

Compare collocated PTRAK readings. These will later be averaged when duplicate instruments are used. PMPT_94 has collected most of the measurements and been co-located with PMPT_93, PMPT_4 and PMPT_2. PMPT_1 and PMPT_93 collected measurements early on in the study. PMPT_93 was co-located with PMPT_94 and PMPT_4. Can't compare PMPT_1 to other PTRAKs because it was never co-located.

### PTRAK - PTRAK

PTRAK 93 tends to have slightly lower readings than primary PTRAK 94.

```{r}

mm_full %>%
  spread(instrument_id, value) %>%
  colo.plot(data.wide = .,
            x.variable = "PMPT_94",  
            y.variable = "PMPT_93", 
            mytitle = paste0("Comparison of collocated PTRAKs (pt/cm3)"), 
            col.by = "route")
```

PMPT_4 has slightly higher reaadings than primary PMPT_94.

```{r}
mm_full %>%
  spread(instrument_id, value) %>%
  colo.plot(data.wide = .,
            x.variable = "PMPT_94",  
            y.variable = "PMPT_4", 
            mytitle = paste0("Comparison of collocated PTRAKs (pt/cm3)"), 
            col.by = "route")
```

PMPT_2 has slightly lower reaadings than primary PMPT_94.

```{r}
mm_full %>%
  spread(instrument_id, value) %>%
  colo.plot(data.wide = .,
            x.variable = "PMPT_94",  
            y.variable = "PMPT_2", 
            mytitle = paste0("Comparison of collocated PTRAKs (pt/cm3)"), 
            col.by = "route")
```

PMPT_4 has slightly lower reaadings than more commonly used PMPT_93. Instruments have a better fit on Route 2 than 7?

```{r}
mm_full %>%
  spread(instrument_id, value) %>%
  colo.plot(data.wide = .,
            x.variable = "PMPT_93",  
            y.variable = "PMPT_4", 
            mytitle = paste0("Comparison of collocated PTRAKs (pt/cm3)"), 
            col.by = "route")

```


### PTRAK - diSCMini

```{r}

```

### NanoScan - NanoScan

PMPSCAN_3 has slightly lower reaadings than primary PMSCAN_5.

```{r} 
ns_trim <- 0.00  # 0.01

mm_full %>%
  filter(grepl("scan", variable)) %>%
  # time stamps are sometimes a ~ 1 second off.
  mutate(time = floor_date(time, unit = "minute")) %>%
  spread(instrument_id, value) %>%
  colo.plot(
    x.variable = "PMSCAN_5", #x.label = "NanoScan ID 5",
    y.variable = "PMSCAN_3", #y.label = "NanoScan ID 3",
    mytitle = paste0("Comparison of collocated NanoScan instrument readings"), 
    col.by = "route")

```

### PTRAK - NanoScan

Compare NanoScan (24-422 nm) to ptrak (20-1000 nm) since NanoScan tends to be more steady. Assuming few particles > 420 nm, such that NanoScan and PTRAK readings are comparable.

Dropping nanoscan readings above 500k pt/cm3 (PTRAK instrument threshold)

### --> use ceiling_date() instead of floor_date() to align PTRAK & NS readings?   

```{r}
# use wide (takes avg of identical collocatd instruments) since not comparing specific instrument IDs, but instrument Type   
t <- mm_full %>%
  filter((grepl("ptrak", variable) & value < ptrak_lim) |
           grepl("scan", variable)) %>%
  # take avg ptrak reading at the "00" second mark for each instrument, for comparison w/ NanoScan
  mutate(time = floor_date(time, unit = "minute")) %>%  #ceiling_date()
  group_by(time, variable, route, instrument_id) %>%
  dplyr::summarize(value = mean(value)) %>%
  spread(variable, value)

# keep duplicate PTRAK & NS readings
t_ptrak <- t %>%
  select(time, route, ptrak_pt_cm3) %>%
  drop_na()

t_ns <- t %>%
  select(time, scan_20_420_pt_cm3) %>%
  drop_na()

full_join(t_ptrak, t_ns) %>%
  colo.plot(x.variable = "scan_20_420_pt_cm3", x.label = "NanoScan (24-422 nm pt/cm3)",
                      y.variable = "ptrak_pt_cm3", y.label = "PTRAK 94 (20-1000 nm pt/cm3)",
                      mytitle = paste0("Comparison of co-located 1-min avg PTRAK and NanoScan readings"), mysubtitle = paste0("dropped PTRAK readings >= ", ptrak_lim, " pt/cm3"),
            col.by = "route" 
            )

```

## Aethalometer - Aethalometer

First, round reading time to nearest 10-seconds since instrument time stamps may not be for the exact same 10 seconds.

```{r}
 mm_full %>%
  filter(grepl("bc", variable)) %>%
  # 2020-01-07 has issues where the time stamp jumps from being 5 sec into the interval to being 4 sec into the interval. creates merging issues where some points are dropped. round to nearest 5-sec to adjust for this.
  mutate(time_round = round_date(time, unit = "5s")) %>%
  #drop time, otherwise merging has fewer pairs 
  select(-time) %>%
   spread(instrument_id, value) %>%
  colo.plot(x.variable = "BC_63", #x.label = "BC_63 ng/m3",
          y.variable = "BC_66", #y.label = "BC_66 ng/m3",
          mytitle = "Co-located aethalometer readings (ng/m3)", 
          col.by = "route" 
          )
 
```

## --> ? compare BC readings to AQS site readings (2 min & overnight). select a reference instrument that is most similar?

(cite: Minet 2017 & their refs: Deville Cvaellin 2016; Lin 2015)

```{r}

```







Save datasets.

````{r}
# saveRDS(mm_full, file.path("Data", "Aim 2", "Mobile Monitoring", "mm_full_2020-03-08.rda"))

```
