---
title: "Mobile Monitoring Data Report - UFP, PM2.5, BC"
author: "Magali Blanco"
date: ' `r Sys.Date()` '
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}  
knitr::opts_chunk$set(echo = F, cache=T, cache.comments = F, message = F, warning = F, tidy.opts=list(width.cutoff=60), tidy=TRUE, fig.height = 8)  

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
      detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(dplyr, tidyverse, VCA, geepack, knitr)    #chron

#set plot theme
theme_set(theme_linedraw() + theme(legend.position="bottom")) 
 
```


```{r average stop data}
source("A2-3_Var&Fns.R")

#load stop average data 
mm <- readRDS(file.path("Data", "MobileMonitoring", "mm_190917.rda")) %>% 
  #only keep relevant particle values
  filter(variable %in% c("ufp_pt_noscreen_ct_cm3",
                         "ufp_disc_ct_cm3",
                          "ufp_scan_ct_cm3",
                          "bc_ir880nm_ng_m3",
                          "neph_bscat_per_m"
                         )
         )


mm.wide <- readRDS(file.path("Data", "MobileMonitoring", "mm.wide_190917.rda"))  

#load overnight AQS site data 
ptrak <- readRDS(file.path("Data", "MobileMonitoring", "ptrak_190628.rda"))

#start/end dates
#start.date = min(mm$date) #in A2-3 variables.R sourcing
end.date = max(mm$date)

#on what days have instruments been used?
instrument.use <- mm %>% group_by(instrument_id) %>%
  summarize(no_runs = length(unique(runname)))
 
```

Using untrimmed, median instrument values for each stop (there are many stops per location).

Data used in this report are those greater than the `r myquantile_lower` quantile and less than the `r myquantile_upper` quantile.


# Data Completeness 

```{r}
total_site_visits <- nrow(mm.wide)

instrument_readings <- mm.wide %>%
  select(ufp_disc_ct_cm3, 
         ufp_pt_noscreen_ct_cm3, 
         ufp_scan_ct_cm3, 
         bc_ir880nm_ng_m3, 
         neph_bscat_per_m)  

unique_instruments <- dim(instrument_readings)[2]

total_stops <- instrument_readings %>% nrow()

complete_instrument_readings <- instrument_readings %>%
  drop_na() %>% nrow()
   
  
data_completeness <- data.frame(variable = names(instrument_readings),
                 stop_count = NA,
                 percent_complete = NA)

for(i in 1:nrow(data_completeness)) {
  data_completeness$stop_count[i] <- nrow(drop_na(instrument_readings[i]))
  data_completeness$percent_complete[i] <- data_completeness$stop_count[i] / total_site_visits *100  
  }

data_completeness <- data_completeness %>%
  arrange(desc(percent_complete))
  
total_df <- data.frame(variable = c("total stops", paste("all", unique_instruments, "instruments running")),
                       stop_count = c(total_stops, complete_instrument_readings),
                       percent_complete = c(NA, complete_instrument_readings/total_stops*100))

data_completeness <- rbind(total_df, data_completeness)

# data_completeness <- data_completeness %>% 
#   mutate(percent_complete = round(percent_complete, 1)) 
   
kable(data_completeness, digits = 1, caption = "Data Completeness")

```
 
# Data Collection Over Time  

```{r}
#instrument used each day
mm %>%
  #1 record per variable combination - to reduce file size
  select(date, instrument_id, variable) %>% 
  unique() %>%
  ggplot(aes(x=date, y=instrument_id, colour = instrument_id)) + 
  geom_point() + 
  facet_wrap(~variable, scales="free", ncol = 2) + 
  theme(legend.position = "none") + 
  labs(title = "Instruments used over time")

```

```{r}
#stops over time 
mm %>%  
  #1 record per date-site_no to reduce file size
  select(date, site_no, route, aqs_id) %>%   
  unique() %>%
  #all sites
  ggplot(aes(x=date, y=site_no, colour=route)) + 
  geom_point() + 
  #aqs sites 
  geom_point(aes(shape=aqs_id), colour="black") + 
  labs(title = paste("Sites visited over time", subtitle= paste(start.date, "-", end.date))) 
  
```

```{r}
#stops by season/day of week/time of day/hour 
##stops by season
mm %>%  
  select(runname, site_no, season, site_id) %>%   
  unique() %>%
  ggplot(aes(x=site_no, fill=season)) + 
  geom_bar() + 
  labs(title = "Number of visits to each site by season", subtitle= paste(start.date, "-", end.date)) 
  
 
```

```{r}
##stops by time of week
mm %>%  
  select(runname, site_no, time_of_week) %>%   
  unique() %>%
  #all sites
  ggplot(aes(x=site_no, fill=time_of_week)) + 
  geom_bar() + 
  labs(title = "Number of visits to each site by time of week", subtitle= paste(start.date, "-", end.date)) 

```

```{r}
##stops by day of the week
mm %>%  
  select(runname, site_no, day) %>%   
  unique() %>%
  #all sites
  ggplot(aes(x=site_no, fill=day)) + 
  geom_bar() + 
  labs(title = "Number of visits to each site by week day", subtitle= paste(start.date, "-", end.date)) 

```

```{r}
##stops by time of day
mm %>%  
  select(runname, site_no, time_of_day) %>%   
  unique() %>%
  #all sites
  ggplot(aes(x=site_no, fill=time_of_day)) + 
  geom_bar() + 
  labs(title = "Number of visits to each site by time of day", subtitle= paste(start.date, "-", end.date)) 

```

# Data Quality Control  
## Field Collocations  

### PTRAKs

```{r}
#need to compare PMPT_1 to PMPT_93 and PMPT_94, and PMPT_93 to PMPT_94


#the colocations we actually have
colo.plot(primary.instrument = "PMPT_93", 
          secondary.instrument = "PMPT_4",
          r2.digits =3)

colo.plot(primary.instrument = "PMPT_94", 
          secondary.instrument = "PMPT_4",
          r2.digits =3)


```

### --> ? 
Compare nanoscan (20 nm +) to ptrak () since nanoscan is tends to be more steady.
Note that bin counts are for mid points of the bin, so these concentrations

```{r}


```


### DiscMini's

```{r} 
colo.plot(primary.instrument = "PMDISC_3", 
          secondary.instrument = "PMDISC_8")

```

### NanoScans

```{r} 
colo.plot(primary.instrument = "PMSCAN_5", 
          secondary.instrument = "PMSCAN_3")

```


### BC    
Fit is not great, but BC_63 has always/mostly been used.

```{r}
colo.plot(primary.instrument = "BC_63", 
          secondary.instrument = "BC_66")

```

### Neph   
PM25_176 has always/mostly been used.

```{r}
colo.plot(primary.instrument = "PM25_176", 
          secondary.instrument = "PM25_205",
          int.digits=8,
          rmse.digits = 8)

```

## Lab/Garage Collocations

### --> to do

### PTRAKs   

```{r}
# see data in database 
#PTRAK lab collocations:
# 3/20
# 3/22
# 4/9 Lab ptrak
# 4/18 Lab ptrak
# 4/19 (just DiscMini)
# 5/6 Lab ptrak
# 5/8 Lab Ptrak zerp; Discmini colo
# 5/9 lab scan
# 5/10 discmini, nanoscan 
# 5/31 disc
# 7/9 ptrak, scan
# 7/10 disc zero
# 8/9 disc, ptrak



```

## Overnight AQS Site Collocations
Using PST.   

### PTRAK

#### Plots 

Plotting mean hourly concentrations (all data included) at Beacon hill (BH) and 10th & Weller (10W).

```{r}
ptrak %>%
  ggplot(aes(x=hour, y=Conc_pt_cm3)) + 
  geom_boxplot(aes(group=hour, fill=time_of_day)) +
  geom_smooth(se=F, method = "loess") + 
  facet_wrap(~Location) #+DaylightSavings

ptrak %>%
  ggplot(aes(x=day, y=Conc_pt_cm3)) + 
  geom_boxplot() + 
  geom_smooth(aes(x=as.numeric(day)), se=F, method = "loess") + 
  facet_wrap(~Location)

ptrak %>%
  ggplot(aes(x=month, y=Conc_pt_cm3)) + 
  geom_boxplot() + 
  #geom_smooth(aes(x=as.numeric(month)), se=F, method = "loess") + 
  facet_wrap(~Location)


```

#### Models by time of day, time of week and season   
##### LS linear Fit

Weekend is lower than weekday; winter is higher than spring

```{r}
#only spring (vs winter) is significant
lm(Conc_pt_cm3 ~ time_of_day + time_of_week + season, data=ptrak) %>%
  summary() 
 
```

###### --> calc & add to plots adjusted means

```{r}

```


##### ANOVA Variance Component Analyses for random models

```{r}
###### --> ? delete? can't calc variance w/ just 2 categores (e.g., time_of_week) ?

anovaVCA(Conc_pt_cm3 ~ time_of_day + time_of_week + season, 
         Data=as.data.frame(ptrak))  

```

#### Models by hour, day and month   
##### LS linear fit   

noon is higher than midnight; Sat & Sun are lower than Monday; April and June are lower than January 

```{r}
#sign: hour, Sat, Sun, Apr, Jun, Nov
                      #factor(hour_night)
lm(Conc_pt_cm3 ~ factor(hour) + day + month, data=ptrak) %>%
  summary()

```

###### --> calc & add to plots adjusted means

```{r}

```


##### ANOVA Variance Component Analyses for random models

Most of the variation is within groups, some from month, and a minor amount from hour and day.

```{r}
anovaVCA(Conc_pt_cm3 ~ hour + day + month, 
         Data=as.data.frame(ptrak)) 

```

## 2-minute stops a AQS sites

### PTRAK

```{r}
mm %>%
  filter(aqs_location == "Beacon Hill" |
           aqs_location == "10th & Weller") %>%
  ggplot(aes(x=hour, y=value)) + 
  geom_boxplot(aes(group=hour, fill=time_of_day)) +
  geom_smooth(se=F, method = "loess") + 
  facet_wrap(~aqs_location) # + time_of_week
 
mm %>%
  filter(aqs_location == "Beacon Hill" |
           aqs_location == "10th & Weller") %>%
  ggplot(aes(x=day, y=value)) + 
  geom_boxplot() + 
  geom_smooth(aes(x=as.numeric(day)), se=F, method = "loess") + 
  facet_wrap(~aqs_location)

mm %>%
  filter(aqs_location == "Beacon Hill" |
           aqs_location == "10th & Weller") %>%
  ggplot(aes(x=as.numeric(month), y=value)) + 
  geom_boxplot(aes(group=month)) + 
  geom_smooth(aes(), se=F, method = "loess") + 
  facet_wrap(~aqs_location) 


```


# Data Summary  

## Concentrations Over Time

```{r}
#concentrations by time & season 
mm %>%
  ggplot(aes(x=date, y=value)) + 
  geom_boxplot(aes(fill=season, group=month)) + 
  geom_smooth(aes(), se=F, method = "loess") +  
  facet_wrap(~variable, scales="free") +  #, ncol = 2
  labs(title= "Concentrations over time",
       x="month",
       subtitle = paste0(start.date, " - ", end.date)) + 
  scale_y_log10()

```

```{r}
##ptraks
 mm %>%
  filter(variable== "ufp_pt_noscreen_ct_cm3") %>%
  ggplot(aes(x=date, y=value)) + 
  geom_boxplot(aes(fill=season, group=month)) + 
  # error, doesn't print loess line correctly on x axis
  geom_smooth(aes(), se=F, method = "loess") +  
  labs(title= "PTRAK Concentrations over time",
       x="month",
       subtitle = paste0(start.date, " - ", end.date))  + 
  scale_y_log10()   

```


```{r}
#concentrations by week day
mm %>%
  ggplot(aes(x=day, y=value)) + 
  geom_boxplot(aes(fill=time_of_week)) + 
  geom_smooth(aes(x=as.numeric(day)), se=F, method = "loess") +  
  facet_wrap(~variable, scales="free") + 
  labs(title= "Concentrations by week day",
       subtitle = paste0(start.date, " - ", end.date)) + 
  scale_y_log10()
 
```

```{r}
##ptraks
mm %>%
  filter(variable== "ufp_pt_noscreen_ct_cm3") %>%
  ggplot(aes(x=day, y=value)) + 
  geom_boxplot(aes(fill=time_of_week)) + 
  facet_wrap(~variable, scales="free") + #, ncol = 3
  geom_smooth(aes(x=as.numeric(day)), se=F, method = "loess") +  
  labs(title= "PTRAK Concentrations by week day",
       subtitle = paste0(start.date, " - ", end.date))  + 
  scale_y_log10()

```


```{r}
#concentrations by hour and time of day
mm %>%
  ggplot(aes(x=hour, y=value)) + 
  geom_boxplot(aes(fill=time_of_day, group=hour)) + 
  geom_smooth(aes(), se=F, method = "loess") +  
  facet_wrap(~variable, scales="free") + #, ncol = 3
  labs(title= "Concentrations by hour",
       subtitle = paste0(start.date, " - ", end.date))  + 
  scale_y_log10()

 
```

```{r}
##ptraks
mm %>%
  filter(variable== "ufp_pt_noscreen_ct_cm3") %>%
  ggplot(aes(x=hour, y=value)) + 
  geom_boxplot(aes(fill=time_of_day, group=hour)) + 
  #geom_point(alpha=0.1)+
  geom_smooth(aes(x=hour), se=F, method = "loess") +  
  labs(title= "PTRAK Concentrations by hour",
       subtitle = paste0(start.date, " - ", end.date))  + 
  scale_y_log10()


```

## Concentrations by Location


```{r}
mm %>%
  ggplot(aes(x=site_no, y=value)) + 
  geom_point(alpha=0.1, aes(colour=route)) +   
  geom_smooth( se=F, method = "loess") +  
  facet_wrap(~variable, scales="free", ncol = 3) + 
  labs(title= "Concentrations by Site",
       subtitle = paste0(start.date, " - ", end.date)) + 
  scale_y_log10()

```

```{r}
# ptrak
# mm %>%
#   filter(variable == "ufp_pt_noscreen_ct_cm3") %>%
#   ggplot(aes(x=site_no, y=value)) + 
#   geom_point(alpha=0.1, aes(colour=route)) +   
#   geom_smooth( se=F, method = "loess") +  
#   facet_wrap(~variable, scales="free", ncol = 3) + 
#   labs(title= "PTRAK Concentrations by Site",
#        subtitle = paste0(start.date, " - ", end.date))  + 
#  scale_y_log10()

```


## Concentrations by Geographic Features 

### --> need geocovariates  

```{r}
#distance to roadway

#distance to airport

#population density

#elevation

```

## UFP Models by time of day, time of week and season   
### LS linear Fit

AM and noon are higher than early AM; weekend is lower than weekday; winter is higher than spring 

```{r}
   mm %>%
  filter(variable == "ufp_pt_noscreen_ct_cm3") %>%
lm(value ~ time_of_day + time_of_week + season + site_id, data=.) %>%
  summary()

```

#### --> ? use mixed effects model instead (id=site_id)? ?serial correlation structure?

```{r}
# mm %>% 
#   filter(variable == "ufp_pt_noscreen_ct_cm3") %>%
#   #can't have empty levels in geeglm()
#   mutate(season = droplevels(season)) %>%   
#   geeglm(value ~ time_of_day + time_of_week + season,
#        data = ., id = site_id, 
#        corstr = "ar1")

```


### ANOVA Variance Component Analyses for random models

Most of the variation is "noise" within groups; some variation due to site_id

```{r}
###### --> ? delete? can't calc variance w/ just 2 categores (e.g., time_of_week) ?
   mm %>%
  filter(variable == "ufp_pt_noscreen_ct_cm3") %>%
  as.data.frame() %>%
  anovaVCA(value ~ time_of_day + time_of_week + season + site_id, 
         Data=.)  

```

## Models by hour, day and month   
### LS linear fit   

no significant difference between 5 AM and other hours; Wed and Thur are higher than Mon; Sat is lower than Monday; Mar, Apr & Aug higher than Feb 

```{r}
 mm %>%
  filter(variable == "ufp_pt_noscreen_ct_cm3") %>%
lm(value ~ factor(hour) + day + month + site_id, data=.) %>%
  summary()

```

#### --> ? use mixed effects model instead (id=site_id)? ?serial correlation structure?

```{r}

```


### ANOVA Variance Component Analyses for random models

Most of the variation is within groups, a little from site, a minor amount from month & hour

```{r}
mm %>%
  filter(variable == "ufp_pt_noscreen_ct_cm3") %>%
  as.data.frame() %>%
 anovaVCA(value ~ hour + day + month + site_id, 
         Data=.) 

```



## Pollutant Correlations  

UFPs are poorly correlated with PM2.5 & BC

```{r}
mm.wide %>%
  select(ufp_disc_ct_cm3, 
         ufp_pt_noscreen_ct_cm3, 
         ufp_scan_ct_cm3, 
         bc_ir880nm_ng_m3, 
         neph_bscat_per_m) %>%
  plot()  

```

Pearson Correlations

```{r}
mm.wide %>%
  select(ufp_disc_ct_cm3, 
         ufp_pt_noscreen_ct_cm3, 
         ufp_scan_ct_cm3, 
         bc_ir880nm_ng_m3, 
         neph_bscat_per_m) %>%
   cor(use = "complete.obs") %>%
   kable()

```



 
