---
title: "Mobile Monitoring Data Report - UFP, PM2.5, BC"
author: "Magali Blanco"
date: ' `r Sys.Date()` '
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}  
knitr::opts_chunk$set(echo = F, cache=T, cache.comments = F, message = F, warning = F, tidy.opts=list(width.cutoff=60), tidy=TRUE, fig.height = 8)  

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
      detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(dplyr, tidyverse, VCA, geepack, knitr)    #chron

#set plot theme
theme_set(theme_linedraw() + theme(legend.position="bottom")) 
 
```


```{r average stop data}
source("A2.0.1_Var&Fns.R")

#load stop average data 
mm <- readRDS(file.path("Data", "Aim 2", "Mobile Monitoring", "mm_191112.rda")) %>% 
  #only keep relevant particle values
  filter(variable %in% c("ptrak_ct_cm3",
                         "bc_880_ng_m3",
                         #"ufp_disc_ct_cm3",
                          "ufp_scan_10_421_nm_ct_cm3",
                         "scan_20_421_nm_ct_cm3_median"
                          
                         )
         )


mm.wide <- readRDS(file.path("Data", "Aim 2", "Mobile Monitoring", "mm.wide_191112.rda"))  

#load overnight AQS site data 
ptrak <- readRDS(file.path("Data", "Aim 2", "Overnight Collocations", "ptrak.rda"))

#start/end dates
end.date = max(mm$date)

plot_quantile <- 0.99

#on what days have instruments been used?
instrument.use <- mm %>% group_by(instrument_id) %>%
  summarize(no_runs = length(unique(runname)))
 
```

Using untrimmed, median instrument values for each stop (there are many stops per location).

Data used in this report are those greater than the `r myquantile_lower` quantile and less than the `r myquantile_upper` quantile.


# Data Completeness     
##### --> numbers seem high?

```{r}
total_site_visits <- nrow(mm.wide)

instrument_readings <- mm.wide %>%
  select(bc_880_ng_m3_median: scan_20_421_nm_ct_cm3_median)

unique_instruments <- dim(instrument_readings)[2]

total_stops <- instrument_readings %>% nrow()

complete_instrument_readings <- instrument_readings %>%
  drop_na() %>% nrow()
   
  
data_completeness <- data.frame(variable = names(instrument_readings),
                 stop_count = NA,
                 percent_complete = NA)

for(i in 1:nrow(data_completeness)) {
  data_completeness$stop_count[i] <- nrow(drop_na(instrument_readings[i]))
  data_completeness$percent_complete[i] <- data_completeness$stop_count[i] / total_site_visits *100  
  }

data_completeness <- data_completeness %>%
  arrange(desc(percent_complete))
  
total_df <- data.frame(variable = c("total stops", paste("all", unique_instruments, "instruments running")),
                       stop_count = c(total_stops, complete_instrument_readings),
                       percent_complete = c(NA, complete_instrument_readings/total_stops*100))

data_completeness <- rbind(total_df, data_completeness)

kable(data_completeness, digits = 1, caption = "Data Completeness")

```
 
# Data Collection Over Time  

```{r}
#instrument used each day
mm %>%
  filter(
    #grepl(pattern = "ufp|bc", x = variable),
    variable != "scan_20_421_nm_ct_cm3_median"
         ) %>%
  #1 record per variable combination - to reduce file size
  select(date, instrument_id, variable) %>% 
  unique() %>%
  ggplot(aes(x=date, y=instrument_id, colour = instrument_id)) + 
  geom_point() + 
  facet_wrap(~variable, scales="free", ncol = 2) + 
  theme(legend.position = "none") + 
  labs(title = "Instruments used over time")

```

```{r, include=F}
#stops over time 
mm %>%  
  #1 record per date-site_no to reduce file size
  select(date, site_no, route, aqs_id) %>%   
  unique() %>%
  #all sites
  ggplot(aes(x=date, y=site_no, colour=route)) + 
  geom_point() + 
  #aqs sites 
  geom_point(aes(shape=aqs_id), colour="black") + 
  labs(title = paste("Sites visited over time", subtitle= paste(start.date, "-", end.date))) 
  
```

#### ---> replace w/ plot on A2_UK?

```{r}
#stops by season/day of week/time of day/hour 
##stops by season
mm %>%  
  select(runname, site_no, season, site_id) %>%   
  unique() %>%
  ggplot(aes(x=site_no, fill=season)) + 
  geom_bar() + 
  labs(title = "Number of visits to each site by season", subtitle= paste(start.date, "-", end.date)) 
  
 
```

```{r}
##stops by time of week
mm %>%  
  select(runname, site_no, time_of_week) %>%   
  unique() %>%
  #all sites
  ggplot(aes(x=site_no, fill=time_of_week)) + 
  geom_bar() + 
  labs(title = "Number of visits to each site by time of week", subtitle= paste(start.date, "-", end.date)) 

```

```{r}
##stops by day of the week
mm %>%  
  select(runname, site_no, day) %>%   
  unique() %>%
  #all sites
  ggplot(aes(x=site_no, fill=day)) + 
  geom_bar() + 
  labs(title = "Number of visits to each site by week day", subtitle= paste(start.date, "-", end.date)) 

```

```{r}
##stops by time of day
mm %>%  
  select(runname, site_no, time_of_day) %>%   
  unique() %>%
  #all sites
  ggplot(aes(x=site_no, fill=time_of_day)) + 
  geom_bar() + 
  labs(title = "Number of visits to each site by time of day", subtitle= paste(start.date, "-", end.date)) 

```

# Data Quality Control  
## Field Collocations  

### PTRAKs

```{r}
#need to compare PMPT_1 to PMPT_93 and PMPT_94, and PMPT_93 to PMPT_94

#the colocations we actually have

mm %>%
    #select only values from desired instruments
    filter(instrument_id %in% c("PMPT_93", "PMPT_4")) %>%
             #make wide format
  select(-mean_value) %>%           
  spread(instrument_id, median_value) %>%
             #only look at rows where have observations for both instruments
             drop_na("PMPT_93", "PMPT_4") %>%
  filter(PMPT_93 <= quantile(PMPT_93, plot_quantile, na.rm = T),
         PMPT_4 <= quantile(PMPT_4, plot_quantile, na.rm = T)
         ) %>%
  colo.plot.wide.data(data.wide = ., x.variable = "PMPT_93",
                      y.variable = "PMPT_4",
                      mytitle = paste0("excluding top ", (1-plot_quantile)*100, "% of data"),
                      r2.digits = 3  
                      )

mm %>%
    #select only values from desired instruments
    filter(instrument_id %in% c("PMPT_94", "PMPT_4")) %>%
  select(-mean_value) %>%                      
  #make wide format
             spread(instrument_id, median_value) %>%
             #only look at rows where have observations for both instruments
             drop_na("PMPT_94", "PMPT_4") %>%
  filter(PMPT_94 <= quantile(PMPT_94, plot_quantile, na.rm = T),
         PMPT_4 <= quantile(PMPT_4, plot_quantile, na.rm = T)
         ) %>%
  colo.plot.wide.data(data.wide = ., x.variable = "PMPT_94",
                      y.variable = "PMPT_4",
                      mytitle = paste0("excluding top ", (1-plot_quantile)*100, "% of data"),
                      r2.digits = 3  
                      )

```

Compare NanoScan (20-421 nm +) to ptrak (20- 1000 nm) since NanoScan tends to be more steady. Assuming few particles > 421 nm, such that the NanoScan and PTRAK are comparable.

Excluding top 1% of data. 

```{r}

mm.wide %>%
  #take out extreme values that make R2 look very large
  filter(scan_20_421_nm_ct_cm3_median <= quantile(scan_20_421_nm_ct_cm3_median, plot_quantile, na.rm = T),
         ptrak_ct_cm3_median <= quantile(ptrak_ct_cm3_median, plot_quantile, na.rm = T)
         ) %>%
  colo.plot.wide.data(data.wide = ., x.variable = "scan_20_421_nm_ct_cm3_median",
                      y.variable = "ptrak_ct_cm3_median", mytitle = paste0("excluding top ", (1-plot_quantile)*100, "% of data"))

```

 
### NanoScans

```{r} 
colo.plot(primary.instrument = "PMSCAN_5", 
          secondary.instrument = "PMSCAN_3")

```


### BC    
Fit is not great, but BC_63 has always/mostly been used.

```{r}
colo.plot(primary.instrument = "BC_63", 
          secondary.instrument = "BC_66")

```

 
## Lab/Garage Collocations

### --> see Lab collocations in database: Imports table/ lab0, lab C, or Rm 114 

### PTRAKs   

```{r}


```

## Overnight AQS Site Collocations
Using PST.   

### PTRAK

Plotting median hourly concentrations (all data included) at Beacon hill (BH) and 10th & Weller (10W).

#### --> ? color by date (have to create date varaible)? 

```{r}

ptrak %>%
  ggplot(aes(x=month, y=Conc_pt_cm3)) + 
  geom_boxplot() + 
  #geom_smooth(aes(x=as.numeric(month)), se=F, method = "loess") + 
  facet_wrap(~location)

ptrak %>%
  ggplot(aes(x=day, y=Conc_pt_cm3)) + 
  geom_boxplot() + 
  geom_smooth(aes(x=as.numeric(day)), se=F, method = "loess") + 
  facet_wrap(~location)

ptrak %>%
  ggplot(aes(x=hour, y=Conc_pt_cm3)) + 
  geom_boxplot(aes(group=hour, fill=time_of_day)) +
  geom_smooth(se=F, method = "loess") + 
  facet_wrap(~location) #+DaylightSavings

```

```{r}
#only spring (vs winter) is significant
lm(Conc_pt_cm3 ~ factor(hour) + day + month, data=ptrak) %>% summary()

lm(Conc_pt_cm3 ~ time_of_day + time_of_week + season, data=ptrak) %>% summary() 
 
anovaVCA(Conc_pt_cm3 ~ hour + day + month, 
         Data=as.data.frame(ptrak)) 

anovaVCA(Conc_pt_cm3 ~ time_of_day + time_of_week + season, 
         Data=as.data.frame(ptrak))  

```

## 2-minute stops a AQS sites

### PTRAK

```{r}
mm %>%
  filter(aqs_location %in% c("Beacon Hill", "10th & Weller")) %>%
  ggplot(aes(x=hour, y=median_value)) + 
  geom_boxplot(aes(group=hour, fill=time_of_day)) +
  geom_smooth(se=F, method = "loess") + 
  facet_wrap(~aqs_location) # + time_of_week
 
mm %>%
  filter(aqs_location %in% c("Beacon Hill", "10th & Weller")) %>%
  ggplot(aes(x=day, y=median_value)) + 
  geom_boxplot() + 
  geom_smooth(aes(x=as.numeric(day)), se=F, method = "loess") + 
  facet_wrap(~aqs_location)

mm %>%
  filter(aqs_location %in% c("Beacon Hill", "10th & Weller")) %>%
  ggplot(aes(x=as.numeric(month), y=median_value)) + 
  geom_boxplot(aes(group=month)) + 
  geom_smooth(aes(), se=F, method = "loess") + 
  facet_wrap(~aqs_location) 


```


# PTRAK Data Summary  

## Temporal Patterns

```{r}
#month
mm %>%
  filter(variable == "ptrak_ct_cm3") %>%
  ggplot(aes(x=date, y=median_value)) + 
  geom_boxplot(aes(fill=season, group=month)) + 
  geom_smooth(aes(), se=F, method = "loess") +  
  facet_wrap(~variable, scales="free") +  #, ncol = 2
  labs(title= "Concentrations over time",
       x="month",
       subtitle = paste0(start.date, " - ", end.date)) + 
  scale_y_log10()

#day
mm %>%
  filter(variable== "ptrak_ct_cm3") %>%
  ggplot(aes(x=day, y=median_value)) + 
  geom_boxplot(aes(fill=time_of_week)) + 
  geom_smooth(aes(x=as.numeric(day)), se=F, method = "loess") +  
  facet_wrap(~variable, scales="free") + 
  labs(title= "Concentrations by week day",
       subtitle = paste0(start.date, " - ", end.date)) + 
  scale_y_log10()
 
 
#hour
mm %>%
  filter(variable== "ptrak_ct_cm3") %>%
  ggplot(aes(x=hour, y=median_value)) + 
  geom_boxplot(aes(fill=time_of_day, group=hour)) + 
  geom_smooth(aes(), se=F, method = "loess") +  
  facet_wrap(~variable, scales="free") + #, ncol = 3
  labs(title= "Concentrations by hour",
       subtitle = paste0(start.date, " - ", end.date))  + 
  scale_y_log10()

 
```

## Spatial Patterns

### --> use lasso to select geocovariates & then plot UFP vs geocov
### ? bad plots since not usign annual averages & obsrvations at each location taken at slighty different times 

```{r}
mm %>%
 filter(variable== "ptrak_ct_cm3") %>%
  ggplot(aes(x=site_no, y=median_value)) + 
  geom_point(alpha=0.2, aes(colour=route)) +   
  geom_smooth( se=F, method = "loess") +  
  facet_wrap(~variable, scales="free", ncol = 3) + 
  labs(title= "Concentrations by Site",
       subtitle = paste0(start.date, " - ", end.date)) + 
  scale_y_log10()



#### --> use lasso instead to select predictive geocovariates 





mm.w.geo <- mm.wide %>% 
  # geocovariates
  select(arrival_time:season,
         ptrak_ct_cm3_median,
         m_to_a1:m_to_a3,
        #airport, largest airport
        m_to_airp, #m_to_l_airp,
        m_to_rr,
        #dist to and length of
        m_to_truck, #contains("tl_s"),
        #m_to_bus, contains("bus_"),
        elev_elevation,
        #contains("lu_industrial_"),
        #contains("ndvi_q50_"),
        pop_s01000, #contains("pop_s")
       )

mm.w.geo$m_to_a1_or_a2 <- apply(mm.w.geo[c("m_to_a1", "m_to_a2")], 1, function(x) min(x))

mm.geo <- gather(mm.w.geo, geocovariate, value, m_to_a1:m_to_a1_or_a2)



geo_quantile <- 0.99

mm.geo %>%
  filter(ptrak_ct_cm3_median <= quantile(ptrak_ct_cm3_median, geo_quantile, na.rm = T)) %>%
  ggplot(aes(x=value, y=ptrak_ct_cm3_median)) + 
  geom_point(alpha=0.3) + 
  geom_smooth( se=F, method = "loess") +  
  facet_wrap(~geocovariate, scales = "free_x") + 
  labs(title = paste0("excludes top ", (1-geo_quantile)*100, "% of observations") )
  
 


```

## Models 
 
variability: error > location > month > day > hour    
variability: error > location > time_of_week > season > time_of_day

```{r}
mm %>%
  filter(variable == "ptrak_ct_cm3") %>%
lm(median_value ~ factor(hour) + factor(day, ordered = F) + factor(month, ordered = F) + site_id, data=.) %>%
  summary()

mm %>%
  filter(variable == "ptrak_ct_cm3") %>%
lm(median_value ~ time_of_day + time_of_week + season + site_id, data=.) %>%
  summary()

mm %>%
  filter(variable == "ptrak_ct_cm3") %>%
  as.data.frame() %>%
 anovaVCA(median_value ~ hour + day + month + site_id, 
         Data=.) 

mm %>%
  filter(variable == "ptrak_ct_cm3") %>%
  as.data.frame() %>%
  anovaVCA(median_value ~ time_of_day + time_of_week + season + site_id, 
         Data=.)  


```


## Pollutant Correlations (R)  

UFPs are poorly correlated with PM2.5 & BC

```{r}
mm.wide %>%
  select(pt = ptrak_ct_cm3_median, 
         ns_20_421 = scan_20_421_nm_ct_cm3_median, 
         bc = bc_880_ng_m3_median) %>%
  cor(method = "pearson",
      use ="complete.obs") %>%
  corrplot::corrplot(method="color", 
                     addCoef.col = T, 
                     title = "Pearson Correlation (R)"
                     )


```

 

 
