---
title: "Aim 2"
author: "Magali Blanco"
date: ' `r Sys.Date()` '
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache=T, cache.comments = F, message = F, warning = F, tidy.opts=list(width.cutoff=60), tidy=TRUE, fig.height = 8)  

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
      detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(dplyr, tidyverse, knitr, Himsc, EnvStats, lubridate, glmnet, ggpubr)   
#Himsc: describe(); EnvStats: summaryFull(); ggpubr: ggarrange()

#set plot theme
theme_set(theme_linedraw() + theme(legend.position="bottom")) 

source("A2.0.1_Var&Fns.R")

```

```{r}
#load stop average data 
mm <- readRDS(file.path("Data", "Aim 2", "Mobile Monitoring", "mm_191112.rda"))
mm.wide <- readRDS(file.path("Data", "Aim 2", "Mobile Monitoring", "mm.wide_191112.rda"))

mm.w.ptrak <- mm.wide %>% 
  select(
    site_id: site_lat, site_no, aqs_site:season,
    ufp_pt_noscreen_ct_cm3
  ) %>%
  #remove NAs
  drop_na(ufp_pt_noscreen_ct_cm3)

```

```{r simple avg for ACT symposium} 
# #estimate site-specific median pollutant concentrations up until now
# median.conc <- mm %>% 
#   #filter(instrument_id =="PMDISC_8") %>% 
#   group_by(site_id, site_lat, site_long, instrument_id) %>% 
#   rename(lat = site_lat, long = site_long) %>%
#   summarize(
#     median_conc = median(value)
#   )
# 
# ##estimates for specific pollutants from primary instruments
# ufp.conc <- median.conc %>%
#   filter(instrument_id == "PMDISC_8") %>%
#   rename(median_conc_ct_m3 = median_conc )
# 
# no2.conc <- median.conc %>%
#   filter(instrument_id == "NO2_2") %>%
#   rename(median_conc_ppb = median_conc )
# 
# # write.csv(ufp.conc, "ufp.conc.csv", row.names = F)
# # write.csv(no2.conc, "no2.conc.csv", row.names = F)

```

### ? use wide format (uses avg when multiple readings exist from duplicate instruments)
## ?? how to select all columns After a certain one? 

# Available MM observations 
  
Lot of locations w/o observations during each season-time_of_week-time_of_day combination
#### --> knit error: "passing fn as a global data...misspelled the 'data' argument in ggplot()

```{r}

#counts at each location by quarter, DOW, hour bins - see how much missin data there are
#df w/ times that each stop Should have sampled
loc <- unique(mm.w.ptrak$site_id)
seasons <- unique(mm.w.ptrak$season)
dow <- unique(mm.w.ptrak$time_of_week)
hour <- unique(mm.w.ptrak$time_of_day)

loc.vec <- rep(loc, each = length(seasons)*length(dow)*length(hour))
seasons.vec <- rep(seasons, each = length(dow)*length(hour), times=length(loc))
dow.vec <- rep(dow, each = length(hour), times=length(loc)*length(seasons))
hour.vec <- rep(hour, times=length(loc)*length(seasons)*length(dow))
 
df <- data.frame(
  site_id = loc.vec,
  season = seasons.vec,
  time_of_week = dow.vec,
  time_of_day = hour.vec
)

#add unique stop variables
df <- mm.w.ptrak %>%
  select(site_id:site_no, aqs_site) %>%
  #drop_na() %>%
  unique()  %>%
  left_join(df)

#df has more rows now b/c some locations were sampled multiple times during same season-week-hour combo
avail_data <- left_join(df, mm.w.ptrak) %>%
  mutate(
    ufp_available = as.numeric(!is.na(ufp_pt_noscreen_ct_cm3))
  )

samples_per_stop <- avail_data %>% 
  group_by(route, site_id, season, time_of_week, time_of_day) %>%
  summarize(
    N = sum(ufp_available)
  )

samples_per_stop_season <- samples_per_stop %>%
    #total stops per season
    group_by(route, site_id, season) %>%
  summarize(
    N = sum(N)
  ) 

#season
samples_per_stop_season %>% 
  ggplot(aes(x=N, fill = (N==0))) + 
  geom_bar() + 
  labs(title= paste("Histograms of No. Samples per", paste0(names(samples_per_stop)[1], collapse = "-"), ""),
       x = paste0("No. Samples from stop locations (", length(unique(avail_data$site_id)), ")")
       ) + 
  facet_grid(~season) 


#season-week
samples_per_stop_season_wk <- samples_per_stop %>%
    #total stops per season
    group_by(route, site_id, season, time_of_week) %>%
  summarize(
    N = sum(N)
  ) 


# t2 <- samples_per_stop_season_wk %>%  filter(
#   season == "fall",  
#   N == 0
#   ) %>% 
#   arrange(time_of_week, route) 

#t2 %>%  write.csv("fall missing stops.csv", row.names = F)


samples_per_stop_season_wk %>% 
  ggplot(aes(x=N, fill = (N==0))) + 
  geom_bar() + 
  labs(title= paste("Histograms of No. Samples per", paste0(names(samples_per_stop)[1], collapse = "-"), ""),
       x = paste0("No. Samples from stop locations (", length(unique(mm.w.ptrak$site_id)), ")")) + 
  facet_grid(time_of_week~season) 
    
  
```

# Trim data   

```{r}
untrimmed.plot <- mm.w.ptrak %>% 
  ggplot(aes(y=ufp_pt_noscreen_ct_cm3)) + 
  geom_boxplot() + #scale_y_log10() +
  labs(title = "untrimmed data")

#trip bottom and top 5% observations 
trimmed.plot <- mm.w.ptrak %>%
  filter(ufp_pt_noscreen_ct_cm3 <= quantile(ufp_pt_noscreen_ct_cm3, (1-trim_quantile), na.rm = T),
         ufp_pt_noscreen_ct_cm3 >= quantile(ufp_pt_noscreen_ct_cm3, (trim_quantile), na.rm = T)
         ) %>%
  ggplot(aes(y=ufp_pt_noscreen_ct_cm3)) + 
  geom_boxplot() +
  labs(title = paste0("trimmed top and bottom ", trim_quantile*100, "% of data " ))
  
ggarrange(untrimmed.plot, trimmed.plot ) 

```

### --> use windosirzed mean instead? 

```{r}
mm.w.ptrak <- mm.w.ptrak %>%
  #trim high and low values
  filter(ufp_pt_noscreen_ct_cm3 <= quantile(ufp_pt_noscreen_ct_cm3, (1-trim_quantile), na.rm = T),
         ufp_pt_noscreen_ct_cm3 >= quantile(ufp_pt_noscreen_ct_cm3, (trim_quantile), na.rm = T)
         )

```


```{r example location w/ missing obs, eval=F}
## ERROR: changed data frame??? avail_data or mm.w.ptrak...?

#example location 
MC0002 <- avail_data %>%
  select(
    site_id,
    date:ufp_available
  ) %>%
  filter(site_id == "MC0002") %>%
  arrange(season, time_of_week, time_of_day)

#early_am and evening will be weighted more heavily b/c we never have sampling @ night
#


#test models
MC0002 %>%
  lm(ufp_pt_noscreen_ct_cm3 ~ time_of_day + time_of_week + season, data = .) %>%
  summary()

```

```{r lme code exp}
#mixed effecdts model
## see B540 L4 # ~57 

#can set covariance structure using nlme::lme()
#fit0 <-lme( fev1 ~ age0 + ageL + relevel(sex, ref="male")*ageL + relevel(f508, ref="none")*ageL,
#method = "ML", #best for inferences at pop/?indiv lvl 
#random = reStruct( ~ 1 + ageL | id, pdClass="pdSymm", REML=F), #random int & slope
#correlation = corAR1( form = ~ 1 | id ), ## within group corr structure
#data = cf.long )

#make predictions
#pred.fit1 <- predict(fit1)

# library(nlme)
# me1 <- mm.w.ptrak %>% 
#   #model won't run w/ missing Y's
#   drop_na(ufp_pt_noscreen_ct_cm3) %>%
#   mutate(
#     time_of_day = factor(time_of_day, ordered = F),
#     time_of_week = factor(time_of_week, ordered = F)
#     ) %>%
#   lme(ufp_pt_noscreen_ct_cm3 ~ time_of_day + time_of_week + season,
#       method = "ML", 
#       # --> add rand slope for other vars?      
#       random =  reStruct( ~ 1 + time_of_day | site_id,  #random int & slope for each site_id
#                                pdClass="pdSymm", REML=F),
#       # --> add? 
#       #correlation = corAR1( form = ~ 1 | site_id), ## within group corr structure
#             data = .) 
# 
# me1 %>% summary()
# 
# mm.w.ptrak$pred.me1 <- predict(me1, newdata = mm.w.ptrak)

#plot(residuals(me1) )
```

# 2. Calculate Weighted Means   
? Assume temporal pattern is same across locations. Strong assumption, but can’t relax this very easily  
  
a) calculate weighted annual avg (~Feb - Oct) means using different methods to determine how sensitve estimates are to using these approaches: 
* weighted (various weights) vs unweighted
* binning hour differently 

b) how may these estimates be impacted over space? are some sites more likely to be sensitive to using different estimation methods? 
* LUR: site means [?using various methods] ~ geocovariates  (e.g. dist to rd/airport/rr, elevation, pop density )    

? c) ? do LUR by hour to see how diff geocovarites impact diff hourly estimates? 

```{r}
#drop winter values for now (drops Feb) since little data here & fewer stops meet min requirements
mm.w.ptrak <- mm.w.ptrak %>%
  filter(season != "winter")

# hist of unique times_of_day locations have sampled at 
mm.w.ptrak %>%
  ggplot(aes(x=hour)) +
  geom_bar(aes( ))

#when are sites least likely to sample (inclue these gaps in other hours)
# mm.w.ptrak %>%
#   ggplot(aes(x=hour, y=site_id)) +
#   geom_point(alpha=0.05)
# 
# table(mm.w.ptrak$hour)

# New Hour Binning
mm.w.ptrak <- mm.w.ptrak %>%
  mutate(
     #? add tod w/ more categories?
    
    tod5 = factor(ifelse(hour %in% seq(3,8), "3_8",
                                       ifelse(hour %in% seq(9,11), "9_11", 
                                              ifelse(hour %in% seq(12,15), "12_15",
                                                     ifelse(hour %in% seq(16,20), "16_20", "21_2")))),
                         levels= c("3_8", "9_11", "12_15", "16_20", "21_2")),
  
    tod3 = factor(ifelse(hour %in% seq(0,8), "0_8",
                                       ifelse(hour %in% seq(9,15), "9_15", "16_23")),
                         levels= c("0_8", "9_15", "16_23")),
    tod2 = factor(ifelse(hour %in% seq(0,11), "0_11", "12_23"),
                         levels= c("0_11", "12_23"))
    )

```

? as bins increase, sample size will also increase. keep same locations for consistency??

```{r}
#calc diff summary measures
mm.w.ptrak <- mm.w.ptrak %>%
  group_by(site_id) %>%
  #unweighted mean
  mutate(mean_uw = round(mean(ufp_pt_noscreen_ct_cm3)),
         #unweighted median
         median_uw = round(median(ufp_pt_noscreen_ct_cm3)),
         )


#means by seaason-wk-tod5 (5 tods)
   #there are no sites w/ sampling during all 30/40 unique times (5 tod x 2 tow x 3 or 4 seasons)
s_tow2_tod5 <- mm.w.ptrak %>%
  group_by(site_id, season, time_of_week, tod5) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  ungroup() %>%
  unite("swh", season:tod5, remove = F) %>%
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(swh)) 
rm(s_tow2_tod5)

#means by seaason-wk-tod3 (3 tods)
   #there are no sites w/ sampling during all 18/24 unique times (3 tod x 2 tow x 3/4 seasons)
s_tow2_tod3 <- mm.w.ptrak %>%
  group_by(site_id, season, time_of_week, tod3) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  ungroup() %>%
  unite("swh", season:tod3, remove = F) %>%
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(swh)) 
rm(s_tow2_tod3)

#means by seaason-wk-tod2 (2 tods)
   #8 sites w/ sampling during all 12 unique times (2 tod x 2 tow x 3 seasons)
   #none when looking at all 4 seasons
s_tow2_tod2 <- mm.w.ptrak %>%
  group_by(site_id, season, time_of_week, tod2) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  ungroup() %>%
  unite("swh", season:tod2, remove = F) %>%
  #calc number of unique sampling times 
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(swh)) %>%
  #only incude locations with every season, TOW, TOD
  filter(unique_times ==12)

#tod2_df %>% select(site_id) %>% n_distinct()

## using tod2, caclulate weighted mean for 8 sites w/ obs during all time combinations
s_tow2_tod2 <- s_tow2_tod2 %>%
  mutate(
    season.wt = 1/length(unique(season)),
    wk.wt = ifelse(time_of_week == "weekday", 5/7, 2/7),
    tod.wt = ifelse(tod2 == "0_11", 12/24, 12/24),
    wt = season.wt * wk.wt * tod.wt,
    wt_times_mean = wt*mean) %>% 

# #check that weights for all sites = 1. #looks good
# s_tow2_tod2 %>%
#   group_by(site_id) %>%
#   summarize(n = sum(wt))
  
  #calculate each site's weighted mean
  group_by(site_id) %>%
  dplyr::summarize(mean_wt = sum(wt_times_mean))
  

#means by season-wk7 (2 tods)
   #no sites w/ sampling during all 42/56 unique times (2 tod x 7 tow x 3/4 seasons)
s_tow7_tod2 <- mm.w.ptrak %>%
  group_by(site_id, season, day, tod2) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  ungroup() %>%
  unite("swh", season:tod2, remove = F) %>%
  #calc number of unique sampling times 
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(swh))  
  #only incude locations with every season, TOW 
rm(s_tow7_tod2)

#no tod
   # no site w/ 21/28 unique times (3/4 seasons x 7 days)
s_tow7 <- mm.w.ptrak %>%
  group_by(site_id, season, day) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  ungroup() %>%
  unite("swh", season:day, remove = F) %>%
  #calc number of unique sampling times 
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(swh))
rm(s_tow7)

   # 219 sites w/ sampling during all 6 unique times (2 tow x 3 seasons)
   # 65 sites w/ sampling 8 unique times (2 two x 4 seasons)
s_tow2 <- mm.w.ptrak %>%
  group_by(site_id, season, time_of_week) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  ungroup() %>%
  unite("swh", season:time_of_week, remove = F) %>%
  #calc number of unique sampling times 
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(swh)) %>%
  #only incude locations with every season, TOW, TOD
  filter(unique_times ==6) %>%
#no_tod_df %>% select(site_id) %>% n_distinct()
#  caclulate weighted mean for sites w/ obs during all time combinations
  mutate(
    season.wt = 1/length(unique(season)),
    wk.wt = ifelse(time_of_week == "weekday", 5/7, 2/7),
    wt = season.wt * wk.wt,
    wt_times_mean = wt*mean) %>%
  #calculate each site's weighted mean
  group_by(site_id) %>%
  dplyr::summarize(mean_wt = sum(wt_times_mean))


# month tow2
#    no sites w/ 18 unique times (9 mo x 2 tow)
m_tow2 <- mm.w.ptrak %>%
  group_by(site_id, month, time_of_week) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  ungroup() %>%
  unite("swh", month:time_of_week, remove = F) %>%
  #calc number of unique sampling times 
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(swh)) #%>%
rm(m_tow2)


#season
   # 305 sites w/ sampling during all 3 seasons
s <- mm.w.ptrak %>%
  group_by(site_id, season) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  #calc number of unique sampling times 
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(season)) %>%
  #only incude locations with every season, TOW, TOD
  filter(unique_times ==3) %>%
#no_tod_df %>% select(site_id) %>% n_distinct()
#  caclulate weighted mean for sites w/ obs during all time combinations
  mutate(
    season.wt = 1/3,
    wt_times_mean = season.wt*mean) %>%
  #calculate each site's weighted mean
  group_by(site_id) %>%
  dplyr::summarize(mean_wt = sum(wt_times_mean))

#month-weighted avg
   # 309 sites. some sites have more months sampled than others
m.diff_no_months <- mm.w.ptrak %>%
  group_by(site_id, month) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  #calc number of unique sampling times 
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(month)) %>% 
#  caclulate weighted mean for sites w/ obs during all time combinations
  mutate(
    wt_times_mean = mean /unique_times) %>%
  #calculate each site's weighted mean
  group_by(site_id) %>%
  dplyr::summarize(mean_wt = sum(wt_times_mean))
 
#month-weighted avg
   # 101 sites w/ all months sampled 

#locations w/ samples Mar - Nov
months.sampled <- month.abb[3:11]

m.same_months <- mm.w.ptrak %>%
  group_by(site_id, month) %>%
  dplyr::summarize(mean = mean(ufp_pt_noscreen_ct_cm3)) %>%
  filter(month %in% months.sampled) %>%
  #calc number of unique sampling times 
  group_by(site_id) %>%
  mutate(unique_times = n_distinct(month)) %>% 
  filter(unique_times == length(months.sampled)) %>%
#  caclulate weighted mean for sites w/ obs during all time combinations
  mutate(
    wt_times_mean = mean /length(months.sampled)) %>%
  #calculate each site's weighted mean
  group_by(site_id) %>%
  dplyr::summarize(mean_wt = sum(wt_times_mean))


```

compare mean estimates for 8 sites using different methods

### ---> ? do we want "noisier" estimates? taking site-specific characterstics into account?

```{r}
#join all estimates
annual <- mm.w.ptrak %>%
  select(site_id, mean_uw, median_uw) %>%
  unique() %>%
  left_join(s_tow2_tod2) %>%
  rename(mean_sea_tow2_tod2 = mean_wt) %>%
  left_join(s_tow2) %>%
  rename(mean_sea_tow2 = mean_wt) %>%
  left_join(s) %>%
  rename(mean_sea = mean_wt) %>%
  # don't include these b/c these are fundamentally different? e.g., some sites have few months
  # left_join(m.diff_no_months) %>%
  # rename(mean_m.diff_no_months = mean_wt) %>%
  left_join(m.same_months) %>%
  rename(mean_mo = mean_wt)  

annual.l <- annual %>%
  gather(key = "method", value = "ufp", -site_id) %>%
  drop_na()

#plot of number of locations that met criteria for estimating means 
annual.l %>%
  ggplot(aes(x=method)) + 
  geom_bar() + 
  labs(title = "no. sites with mean/median estimates using various methods")

#estimates for 8 locations with all estimation methods

ufp_by_method(dt = annual)

```


### --> make plots of diff btw 'gold standard' and other methods 


### --> add sample size to plots 
### --> standardize SD? 

```{r}
#estimates for locations w/ 5+ estimation methods (not considering tod)
annual %>%
  select(-mean_sea_tow2_tod2) %>%
  drop_na() %>%
  gather(key = "method", value = "ufp", -site_id) %>%
  ggplot(aes(x=site_id, y= ufp, col=method)) + 
  geom_boxplot(aes(group=site_id), show.legend = F) +
  geom_point(aes(shape=method)) + 
  #geom_line() +
  labs(title = paste0("Site mean/median for ", months.sampled[1], " - ", months.sampled[length(months.sampled)], " (Spring - Winter)")  )

annual %>%
  select(-mean_sea_tow2_tod2) %>%
  drop_na() %>%
  gather(key = "method", value = "ufp", -site_id) %>%
  ggplot(aes(x=method, y= ufp, col=site_id)) + 
  geom_boxplot(aes(group=method), show.legend = F) +
  geom_point(alpha=0.3, show.legend = F) + 
  labs(title = paste0("Site mean/median for ", months.sampled[1], " - ", months.sampled[length(months.sampled)], " (Spring - Winter)"))

# DELETE? can kind of see this in previosu plot
# annual %>%
#   select(-mean_sea_tow2_tod2) %>%
#   drop_na() %>%
#   gather(key = "method", value = "ufp", -site_id) %>%
#   group_by(method) %>%
#   summarize(method_sd = sd(ufp)) %>%
#   ggplot(aes(x=method, y= method_sd)) + 
#   geom_point(aes()) + 
#   labs(title = paste0("SD of mean/median estimation methods for several location; ", months.sampled[1], " - ", months.sampled[length(months.sampled)], " (Spring - Winter)"))

```



includes most locations

```{r}
#estimates for locations w/ 4 estimation methods (not considering tow or tod)




```


## --> compare ?select months/season w/ 3-5 dow?

```{r}

```




# ? Compare trimmed means at AQS sites  
if have 6 months of data, compare to the 6-mo estimate at AQS sites
