---
title: 'Aim 2: Discussion'
author: "Magali Blanco"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---


**TO DO**   


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=F, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE,
                      dpi=300
                      #fig.height = 4
                      )  
set.seed(1)


pacman::p_load(knitr, kableExtra,
               tidyverse, dplyr,
               ggrepel #avoid overlapping labels when plotting
               )  

source("0.Global_FNs.R")

# #wraps a long string
# wrapper <- function(x, ...) 
# {
#   paste(strwrap(x, ...), collapse = "\n")
# }

```

# Load data

```{r}
# model performance
## ufp
ufp_r2 <- readxl::read_xlsx(file.path("..", "References", "Mean UFP & BC comparison.xlsx"), sheet = 3) %>%
  #only include studies that have been fully entered into excel or qualify
  filter(Include_in_analysis == TRUE) %>%
  select(-Include_in_analysis) %>%
  mutate(
    Mean_Model_R2 = as.numeric(Mean_Model_R2),
    #shorten names
    Location = gsub("Amsterdam", "AMS", Location),
    Location = gsub("Rotterdam", "RTM", Location)
  ) %>%
  arrange(Sampling_type, Study) %>%
  #remove white spaces from start & end of strings
  mutate_if(is.character, ~str_trim(.))

## bc
bc_r2 <- readxl::read_xlsx(file.path("..", "References", "Mean UFP & BC comparison.xlsx"), sheet = 4) %>%
   #only include studies that have been fully entered into excel or qualify
  filter(Include_in_analysis == TRUE) %>%
  select(-Include_in_analysis) %>%
  mutate(
    #only keep few digits - for some reason some values have many decimal #s
    Model_R2 = ifelse(nchar(Model_R2) > 9, substr(Model_R2, 1,4), Model_R2),
    
    Mean_Model_R2 = as.numeric(Mean_Model_R2),
    #shorten names
    Location = gsub("Amsterdam", "AMS", Location),
    Location = gsub("Rotterdam", "RTM", Location)
  ) %>%
  arrange(Sampling_type, Study)%>%
  #remove white spaces from start & end of strings
  mutate_if(is.character, ~str_trim(.))

```

```{r}
common_r2_names <-str_subset(string = names(bc_r2),pattern =  paste0(names(ufp_r2), collapse = "|"), )   

```


```{r}
#concentrations
## ufp
ufp0 <- readxl::read_xlsx(file.path("..", "References", "Mean UFP & BC comparison.xlsx")) %>%
  #only include studies that have been fully entered into excel or qualify
  filter(Include_in_analysis == TRUE) %>%
  select(-Include_in_analysis) %>%
  #use median if mean is missing
  mutate(Mean = ifelse(is.na(Mean), yes = Median, no = Mean),
         Location = gsub("Amsterdam", "AMS", Location),
         Location = gsub("Utrecht", "UT", Location),
         Location = gsub("Maastricht", "MMa", Location),
         Location = gsub("Rotterdam", "RTM", Location),

         This_study = grepl("Blanco", Study),
         
         Year = as.numeric(str_extract(Study, "[0-9]{4}"))
         
         ) %>%
  #remove white spaces from start & end of strings
  mutate_if(is.character, ~str_trim(.))

ours <- ufp0 %>%
  filter(grepl("Blanco", Study))

# ufp <- ufp0 %>%
#   #drop this study
#   filter(!grepl("Blanco", Study)) 


## bc
bc0 <- readxl::read_xlsx(file.path("..", "References", "Mean UFP & BC comparison.xlsx"), sheet=2) %>%
  #only include studies that have been fully entered into excel or qualify
  filter(Include_in_analysis == TRUE) %>%
  select(-Include_in_analysis) %>%
  mutate(
         Location = gsub("Amsterdam", "AMS", Location),
         Location = gsub("Utrecht", "UT", Location),
         Location = gsub("Maastricht", "MMa", Location),
         Location = gsub("Rotterdam", "RTM", Location),
         
         This_study = grepl("Blanco", Study)
         
         ) %>%
  rename(Mean = Mean_Median) %>%
  #remove white spaces from start & end of strings
  mutate_if(is.character, ~str_trim(.))

bc_ours <- bc0 %>%
  filter(grepl("Blanco", Study))

```

```{r}
common_conc_names <-str_subset(string = names(bc0),pattern =  paste0(names(ufp0), collapse = "|"), )   

```


```{r, eval=F}
# TEST combine R2 and mean tables

## error: get a lot of repeats 

ufp_1 <- ufp0 %>%
  select(Study, Location, Title,
         Sampling_type,
         Mean, Instrument, Instrument_range)

ufp_r2_1 <- ufp_r2 %>%
  select(Study, Location, Title,
         Sampling_type,
    Sampling_approach:site_mean_only)

test <- ufp_1 %>% full_join(ufp_r2_1)



```


```{r}
#common variables
text_size <- 3

```

# Model R2

## Plots


-Note: plotting the primary analysis R2 or the average R2 if multiple models were run without a clear indication of which was the primary model

- Our study model performance is the avg R2 for the primary analysis, using 10-fold CV and OOS predictions.

-"multi-site avg modelled" means the study (spatial model w/ a multi-visit requirement):    
a) had multiple visits per site (>1)   
b) modeled site averages    
- "unclear" means that either the site visit number or other modeling information was unavailable or unclera from the text


**UFP**

### --> put legend on bottom or widen pic? 

```{r, fig.width=8}
ufp_r2 %>%
  select(common_r2_names) %>%
  rbind(bc_r2[common_r2_names]) %>% 
  mutate(Pollutant = c(rep("UFP", nrow(ufp_r2)), rep("BC", nrow(bc_r2)) ),
         Pollutant = factor(Pollutant, levels = c("UFP", "BC")),

    # are they calculating a site average from >1 visit & using this in their analysis (e.g., vs using individual site visits in a regression)
    multi_visit_mean = visits_per_site > 1 & site_mean_only == "yes",
    multi_visit_mean = ifelse(is.na(multi_visit_mean), "unclear", multi_visit_mean),

    this_study = grepl("Blanco", Study),
     
    
  ) %>% 
  group_by(Pollutant) %>%
  mutate(this_study_r2 = Mean_Model_R2[grepl("Blanco", Study)]
         
         ) %>% 
  
  
  ggplot(data=., aes(x=Sampling_type, y=Mean_Model_R2,
             label = Location,
             # shape=multi_visit_mean,
             # col = this_study,
             )) + 
  
      geom_line(alpha=0.4) +
     stat_summary(fun.y = "median", geom = "point", size = 3, 
                  alpha=0.4
                  ) +
      
      geom_point(aes(shape=multi_visit_mean,
             col = this_study,),
             size=2
             ) +
  
  # #city names 
  geom_text_repel(aes(col = this_study,), size = text_size,  
                  #fontface = "plain",  
                    )    +
  #this study
  geom_hline(#yintercept = .$this_study_r2, #.$Mean_Model_R2[.$this_study],
    aes(yintercept = this_study_r2),
             linetype = "dashed",
             alpha=0.5
             ) +

  # # Add a second axis and specify its features
  # scale_y_continuous(sec.axis = sec_axis(~./ufp_r2$Mean_Model_R2[grepl("Blanco", ufp_r2$Study)],
  #                                        name="Model R2 Relative to This Study",
  #                                        breaks = seq(0,1.2,0.2)
  #                                        )) +
      
      
  #      #Add a second axis and specify its features
  # scale_y_continuous(sec.axis = sec_axis(~./r2$this_study_r2 ),
  #                                        name="Model R2 Relative to This Study",
  #                                        breaks = seq(0,1.2,0.2)
  #                                        ) +
      
      facet_wrap(~Pollutant ) +
      
      theme(plot.caption = element_text(size=8,
                                    hjust = 0),
            legend.position = "bottom"
            ) +
  #avoid overlapping x axis ticks
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +

  labs(title = "Model R2 estimates from other studies",
       y = "Model R2",
       x = "Sampling Type",
       caption = paste0("AMS = Amsterdam; UT = Utrecht \nGray dot is the median; vertical line is the range"),
       col = "This Study",
       shape = "Multi-site \nAvg Modelled"
       ) 
  
  

```

```{r, eval=F}
# UFP
#next 2 plots are already shown above

#All studies 

ufp_r2 %>%
  { 
  ggplot(data=., aes(x=Sampling_type, y=Mean_Model_R2,
             label = Location
             )) + 
  
  geom_boxplot(col=alpha("red", 0.4)) +
  
  # #city names 
  geom_text_repel(aes(col = grepl("Blanco", .$Study)),
                  size = text_size,  
                  fontface = "plain",  
                   )    +
  #this study
  geom_hline(yintercept = .$Mean_Model_R2[grepl("Blanco", .$Study)],
             linetype = "dashed",
             alpha=0.5
             ) +
  geom_point(aes(col = grepl("Blanco", .$Study)),
             ) + 
    # Add a second axis and specify its features
  scale_y_continuous(sec.axis = sec_axis(~./ufp_r2$Mean_Model_R2[grepl("Blanco", ufp_r2$Study)], 
                                         name="Model R2 Relative to This Study", 
                                         breaks = seq(0,1.2,0.2)
                                         )
                     ) +
  #theme_classic(base_size = 15) + 
  theme( 
        legend.position = "none",
        plot.caption = element_text(size=8,
                                    hjust = 0),
        ) + 
  labs(title = "Model R2 estimates from other LUR UFP studies",
       y = "Model R2",
       x = "Sampling Type",
       caption = "AMS = Amsterdam; UT = Utrecht"
       )  
  }

```

```{r, eval=F}
#only showing Short-term stationary similar to ours: site visits > 1 & analysis uses site means (vs indidivual site readings)


ufp_r2 %>%
  #drop ST mobile studies w/o a clear # of visits per road segment
  #filter(!is.na(visits_per_site)) %>%
  mutate(
    # are they calculating a site average from >1 visit & using this in their analysis (e.g., vs using individual site visits in a regression)
    multi_visit_mean = visits_per_site > 1 & site_mean_only == "yes"
  ) %>%
  filter(multi_visit_mean == TRUE) %>%
  { 
  
  ggplot(data=., aes(x=Sampling_type, y=Mean_Model_R2,
             label = Location
             )) + 
  
  #geom_boxplot(alpha=0.1) +
  geom_boxplot(col=alpha("red", 0.4)) +
  
  # #city names 
  geom_text_repel(aes(col = grepl("Blanco", .$Study)),
                  size = text_size,  
                  fontface = "plain",  
                   )    +
  #this study
  geom_hline(yintercept = .$Mean_Model_R2[grepl("Blanco", .$Study)],
             linetype = "dashed",
             alpha=0.5
             ) +
  
  geom_point(aes(col = grepl("Blanco", .$Study)),
             ) + 
  
      # facet_wrap(~multi_visit_mean, #scales = "free_x",
      #            ncol = 1,
      #            labeller = as_labeller( c("TRUE" = "analysis uses site mean from multiple visits",
      #                                      "FALSE" = "analysis uses individual site visits"
      #                                      )
      #            )
      #            ) +
  # Add a second axis and specify its features
  scale_y_continuous(sec.axis = sec_axis(~./ufp_r2$Mean_Model_R2[grepl("Blanco", ufp_r2$Study)], 
                                         name="Model R2 Relative to This Study", 
                                         breaks = seq(0,1.2,0.2)
                                         )
                     ) +
      theme(
        legend.position = "none",
        plot.caption = element_text(size=8,
                                    hjust = 0),
        ) + 
  labs(title = "Model R2 estimates from other LUR UFP studies\nthat model overall site average from mulitple site visits",
       y = "Model R2",
       x = "Sampling Type",
       caption = paste0("AMS = Amsterdam; UT = Utrecht"#,
                       # "\nNA plot: number of visits per road segment is unclear"
                        )
       )  
  }

```



```{r, eval=F}
# **BC**
# All studies

bc_r2 %>%  
  ggplot(data=., aes(x=Sampling_type, y=Mean_Model_R2,
                     label = Location
             )) + 
  
  #geom_boxplot(alpha=0.1) +
  geom_boxplot(colour=alpha("red", 0.4)) +
  
  # #city names 
  geom_text_repel(aes(col = grepl("Blanco", bc_r2$Study)),
                  size = text_size,  
                  fontface = "plain",  
                   ) +
  #this study
  geom_hline(yintercept = bc_r2$Mean_Model_R2[grepl("Blanco", bc_r2$Study)],
             linetype = "dashed",
             alpha=0.5
             ) +
  
  geom_point(aes(col = grepl("Blanco", bc_r2$Study)),
             #alpha=0.8
             ) + 
  
  # Add a second axis and specify its features
  scale_y_continuous(sec.axis = sec_axis(~./bc_r2$Mean_Model_R2[grepl("Blanco", ufp_r2$Study)], 
                                         name="Model R2 Relative to This Study", 
                                         breaks = seq(0,1.2,0.2)
                                         )
                     ) +
   
  theme(
    legend.position = "none",
    plot.caption = element_text(size=8, hjust = 0),
        ) + 
  labs(title = "Model R2 estimates from other LUR BC studies",
       y = "Model R2",
       x = "Sampling Type",
       caption = "AMS = Amsterdam; RTM = Rotterdam"
  )
  
   
  
```

```{r, eval=F}
#only showing short-term stationary similar to ours: site visits > 1 & analysis uses site means (vs indidivual site readings)

#fig.height=8, fig.width=6

bc_r2 %>%
  #drop ST mobile studies w/o a clear # of visits per road segment
  #filter(!is.na(visits_per_site)) %>%
  mutate(
    # are they calculating a site average from >1 visit & using this in their analysis (e.g., vs using individual site visits in a regression)
    multi_visit_mean = visits_per_site > 1 & site_mean_only == "yes"
  ) %>%
    filter(multi_visit_mean == TRUE) %>%

  { 
  
  ggplot(data=., aes(x=Sampling_type, y=Mean_Model_R2,
             label = Location
             )) + 
  
  #geom_boxplot(alpha=0.1) +
  geom_boxplot(col=alpha("red", 0.4)) +
  
  # #city names 
  geom_text_repel(aes(col = grepl("Blanco", .$Study)),
                  size = text_size,  
                  fontface = "plain",  
                   )    +
  #this study
  geom_hline(yintercept = .$Mean_Model_R2[grepl("Blanco", .$Study)],
             linetype = "dashed",
             alpha=0.5
             ) +
  
  geom_point(aes(col = grepl("Blanco", .$Study)),
             ) + 
  
      # facet_wrap(~multi_visit_mean, #scales = "free_x", 
      #            ncol = 1,
      #            labeller = as_labeller( c("TRUE" = "analysis uses site mean from multiple visits",
      #                                      "FALSE" = "analysis uses individual site visits"
      #                                      )
      #            )
      #            ) +
  # Add a second axis and specify its features
  scale_y_continuous(sec.axis = sec_axis(~./ufp_r2$Mean_Model_R2[grepl("Blanco", ufp_r2$Study)], 
                                         name="Model R2 Relative to This Study", 
                                         breaks = seq(0,1.2,0.2)
                                         )
                     ) +
      theme(
        legend.position = "none",
        plot.caption = element_text(size=8,
                                    hjust = 0),
        ) + 
  labs(title = "Model R2 estimates from other LUR BC studies\nthat model overall site average from multiple visits",
       y = "Model R2",
       x = "Sampling Type",
       caption = paste0("AMS = Amsterdam; UT = Utrecht"#,
                        #"\nNote: A few studies are excluded due to an uncler number of visits per site or road segment"
                        )
       )  
  }


```


### sampling approaches in other studies (all)

**UFP**

```{r}
ufp_r2 %>%
  mutate(
    multi_visit_mean = visits_per_site > 1 & site_mean_only == "yes",
    multi_visit_mean = ifelse(is.na(multi_visit_mean), "unclear", multi_visit_mean),
  ) %>%  
  gather("variable", "Value", visits_per_site, visit_duration_min, total_site_duration_min, sites_per_model) %>%
  mutate(This_study = ifelse(grepl("Blanco", Study), "This Study", "Other Studies"),
         
         variable = recode_factor(factor(variable),
                                  "sites_per_model" = "sites",
                                  "visit_duration_min" = "visit duration (min)",
                                  "total_site_duration_min" = "total site duration (min)",
                 "visits_per_site"  = "visits per sites"               
         ),
         variable = factor(variable, levels = c("sites", "visits per sites", "visit duration (min)","total site duration (min)")
                           ),
         ) %>%
  
  ggplot(data=., aes(x=Sampling_type, y=Value)) +
      
      stat_summary(fun.y = "median", geom = "point", size = 3,  alpha=0.4) +

  geom_point(aes(col = This_study, shape=multi_visit_mean)) +
     
      #avoid overlapping x axis ticks
      scale_x_discrete(guide = guide_axis(n.dodge=2))+
      scale_y_log10() +
      
      facet_wrap(~variable, scales = "free") +
  
  theme(legend.position = "bottom") +
      
  
  
      labs(title = "Sampling approaches across other UFP studies",
           x = "Sampling type",
           col = "",
           shape = "Multi-site \nAvg Modelled",
           caption = "gray circle is the median"
           )

```

**BC**

```{r}
bc_r2 %>%
  mutate(
    multi_visit_mean = visits_per_site > 1 & site_mean_only == "yes",
    multi_visit_mean = ifelse(is.na(multi_visit_mean), "unclear", multi_visit_mean),
  ) %>%  
  gather("variable", "Value", visits_per_site, visit_duration_min, total_site_duration_min, sites_per_model) %>%
  mutate(This_study = ifelse(grepl("Blanco", Study), "This Study", "Other Studies"),
         
         variable = recode_factor(factor(variable),
                                  "sites_per_model" = "sites",
                                  "visit_duration_min" = "visit duration (min)",
                                  "total_site_duration_min" = "total site duration (min)",
                 "visits_per_site"  = "visits per sites"               
         ),
         variable = factor(variable, levels = c("sites", "visits per sites", "visit duration (min)","total site duration (min)")
                           )
         
         ) %>%
  
  ggplot(data=., aes(x=Sampling_type, y=Value)) +
      
      stat_summary(fun.y = "median", geom = "point", size = 3,alpha=0.4) +

  geom_point(aes(col = This_study, shape=multi_visit_mean)) +
     
      #avoid overlapping x axis ticks
      scale_x_discrete(guide = guide_axis(n.dodge=2))+
      scale_y_log10() +
      
      facet_wrap(~variable, scales = "free") +
  
  theme(legend.position = "bottom") +

      
      labs(title = "Sampling approaches across other BC studies",
           x = "Sampling type",
           col = "",
           shape = "Multi-site \nAvg Modelled",
           caption = "gray circle is the median"
           
           )
```


```{r, eval=F}
 
# t <- ufp_r2 %>%
#   gather("variable", "Value", visits_per_site, visit_duration_min, total_site_duration_min, sites_per_model) %>%
#   mutate(This_study = grepl("Blanco", Study),
#          variable = recode_factor(factor(variable),
#                                   "sites_per_model" = "sites",
#                                   "visit_duration_min" = "visit duration (min)",
#                                   "total_site_duration_min" = "total site duration (min)",
#                  "visits_per_site"  = "visits per sites"               
#                                   
#          )
#          
#          ) %>% # View()
#   filter(This_study==TRUE) %>%
#   group_by(variable) %>%
#   summarize(h_line = Value)
  
 
  
ufp_r2 %>%
  gather("variable", "Value", visits_per_site, visit_duration_min, total_site_duration_min, sites_per_model) %>%
  mutate(This_study = ifelse(grepl("Blanco", Study), "This Study", "Other Studies"),
         
         variable = recode_factor(factor(variable),
                                  "sites_per_model" = "sites",
                                  "visit_duration_min" = "visit duration (min)",
                                  "total_site_duration_min" = "total site duration (min)",
                 "visits_per_site"  = "visits per sites"               
         )
         
         ) %>% 
  {
  ggplot(data=., aes(x=Sampling_type, y=Value,
              label = Location
              )) + 
  
  geom_boxplot(col=alpha("red", 0.4)) +

  # # #city names
  # geom_text_repel(aes(col = This_study),
  #                 size = text_size,
  #                 fontface = "plain",
  #                 show.legend = F
  #                  )    +
  
      #this study
  # geom_hline(data=t, aes(yintercept = h_line),
  #            linetype = "dashed",
  #            alpha=0.5
  #            ) +

  geom_point(aes(col = This_study,
                 ),
             alpha=0.5,
             #alpha == ifelse(.$This_study == "This Study", 1, 0.5)
              
             ) +
      geom_point(data=subset(., This_study=="This Study"),
                 aes(col = This_study,
                     )
                 ) +
      
      #avoid overlapping x axis ticks
      scale_x_discrete(guide = guide_axis(n.dodge=2))+
      scale_y_log10() +
      
      facet_wrap(~variable, scales = "free",
                 ) + 
      
      labs(title = "Sampling approaches across other LUR UFP studies",
           x = "Sampling type",
           col = ""
           )
      
  }
  

```



```{r, eval=F}
# **BC** 
# 
# sampling approach in other LUR BC studies (all)


 
# t <- bc_r2 %>%
#   gather("variable", "Value", visits_per_site, visit_duration_min, total_site_duration_min, sites_per_model) %>%
#   mutate(This_study = ifelse(grepl("Blanco", Study), "This Study", "Other Studies"),
#          variable = recode_factor(factor(variable),
#                                   "sites_per_model" = "sites",
#                                   "visit_duration_min" = "visit duration (min)",
#                                   "total_site_duration_min" = "total site duration (min)",
#                  "visits_per_site"  = "visits per sites"               
#                                   
#          )
#          
#          ) %>% # View()
#   filter(This_study==TRUE) %>%
#   group_by(variable) %>%
#   summarize(h_line = Value)
  
 
  
  bc_r2 %>%
  gather("variable", "Value", visits_per_site, visit_duration_min, total_site_duration_min, sites_per_model) %>%
  mutate(This_study = ifelse(grepl("Blanco", Study), "This Study", "Other Studies"),
         variable = recode_factor(factor(variable),
                                  "sites_per_model" = "sites",
                                  "visit_duration_min" = "visit duration (min)",
                                  "total_site_duration_min" = "total site duration (min)",
                 "visits_per_site"  = "visits per sites"               
         )
         ) %>% 
  {
  ggplot(data=., aes(x=Sampling_type, y=Value,
              label = Location
              )) + 
  
  geom_boxplot(col=alpha("red", 0.4)) +

  # # #city names
  # geom_text_repel(aes(col = This_study),
  #                 size = text_size,
  #                 fontface = "plain",
  #                 show.legend = F
  #                  )    +
  
      #this study
  # geom_hline(data=t, aes(yintercept = h_line),
  #            linetype = "dashed",
  #            alpha=0.5
  #            ) +

  geom_point(aes(col = This_study,
                 ),
             alpha=0.5,
             #alpha= ifelse(grepl("Blanco", .$Study), 1, 0.5)

             ) +
      geom_point(data=subset(., This_study=="This Study"), 
                 aes(col = This_study,
                     )
                 ) +
      
      #avoid overlapping x axis ticks
      scale_x_discrete(guide = guide_axis(n.dodge=2))+
      scale_y_log10() +
      
      facet_wrap(~variable, scales = "free") + 
      
      labs(title = "Sampling approaches across other LUR BC studies",
           x = "Sampling type",
           col = ""
           )
  }
  

```



## Tables

### --> update Sampling_type if change study order

**UFP**

```{r}
ufp_r2 %>%
  select(
    ## update if change study order!
    #Sampling_type,
    
    Study, Location, # = Study_location,
    Sampling = Sampling_approach, 
    
    Analysis_uses_site_mean = site_mean_only, Model_R2) %>%
  rename_if( grepl("_", names(.)), ~gsub("_", " ", .) ) %>%
  kable(caption = "Model R2 estimates from other LUR UFP studies", 
        digits = 2) %>%
  
  ### --> check that these are correct in the end
  pack_rows("Fixed site; Continuous", 1, 5) %>%
  pack_rows("Short-term Mobile", 6, 14) %>%
  pack_rows("Short-term Stationary", 15, 23) %>%
  add_footnote(., label = c("Location: AMS = Amsterdam; MMa = Maastricht; RTM = Rotterdam; UT = Utrecht"),
                notation = "none"
                ) %>%
  kable_styling()
  
```


**BC**

```{r}
bc_r2 %>%
 select(
   
   ## update if change study order!
   #Sampling_type,
   
   Study, Location,# = Study_location, 
   Sampling = Sampling_approach, Analysis_uses_site_mean = site_mean_only, Model_R2) %>%  
  rename_if( grepl("_", names(.)), ~gsub("_", " ", .) ) %>%
  kable(caption = "Model R2 estimates from other LUR BC studies", 
        digits = 2
        ) %>%
  pack_rows("Fixed site; Continuous", 1, 9) %>%
  pack_rows("Short-term mobile", 10, 14) %>%
  pack_rows("Short-term stationary", 15, nrow(bc_r2)) %>%
  add_footnote(., label = c("Location: AMS = Amsterdam; MMa = Maastricht; RTM = Rotterdam; UT = Utrecht"),
               notation = "none"
               ) %>%
  kable_styling()

```



# Concentrations in other studies

## Plots

both pollutants 

```{r, fig.width=8}
set.seed(2)

ufp0 %>%
  select(common_conc_names) %>%
  rbind(bc0[common_conc_names]) %>%
  mutate(Pollutant = c(rep("UFP (pt/cm3)", nrow(ufp0) ), rep("BC (ng/m3)", nrow(bc0) )),
         Pollutant = factor(Pollutant, levels = c("UFP (pt/cm3)", "BC (ng/m3)"))
         ) %>%  
  
  ggplot(aes(y = Mean,
           x=Sampling_type, #x=1,
           )) + 
  
   geom_line( ) +
     stat_summary(fun.y = "median", geom = "point", size = 3,) +
  
  geom_point(aes(col = This_study),
             #alpha=0.6
             ) + 
  # #city names 
  geom_text_repel(size = text_size-0.5, 
                  fontface = "plain",  
                  #make labels more spaced out
                  #box.padding = 0.5, 
                  aes(label = Location,
                      col = This_study
                      )
                   ) +
  
  # # Add a second axis and specify its features
  # scale_y_continuous(sec.axis = sec_axis(~./ours$Mean, 
  #                                        name="UFP Conc Relative to This Study", 
  #                                        breaks = seq(1,9, 2)
  #                                        )
  #                    ) +
  # #this study
  # geom_hline(yintercept = ours$Mean,
  #            linetype = "dashed",
  #            alpha=0.5
  #            ) +
  facet_wrap(~Pollutant, scales="free") +
  
  labs(x = "Sampling Type",
       y = "Mean Concentration",
       title = "Mean (or median) Conc reported by other studies",
       caption = "AMS = Amsterdam; MMa = Maastricht; RTM = Rotterdam; UT = Utrecht \nBlack line and dot are the range and median",
       col = "This Study"
       )


```






```{r, eval=F}
# **UFP**

set.seed(2)

ufp0 %>%
ggplot(aes(y = Mean,
           x=Sampling_type, #x=1,
           )) + 
  geom_boxplot(col=alpha("black", 0.4)) + 
  geom_point(aes(col = This_study,
                 ),
             alpha=0.6
             ) + 
  # #city names 
  geom_text_repel(size = text_size-0.5, 
                  fontface = "plain",  
                  #make labels more spaced out
                  #box.padding = 0.5, 
                  aes(label = Location,
                      col = This_study
                      )
                   ) +
  
  # Add a second axis and specify its features
  scale_y_continuous(sec.axis = sec_axis(~./ours$Mean, 
                                         name="UFP Conc Relative to This Study", 
                                         breaks = seq(1,9, 2)
                                         )
                     ) +
  #this study
  geom_hline(yintercept = ours$Mean,
             linetype = "dashed",
             alpha=0.5
             ) +
  
  theme(
  # axis.text.x = element_blank(),
  # axis.ticks = element_blank(),
  
  legend.position = "none"
  ) +
  
  labs(x = "Sampling Type",
       y = "Mean UFP Conc (pt/cm3)",
       title = "Mean UFP Conc reported in other LUR UFP studies",
       caption = "AMS = Amsterdam; MMa = Maastricht; RTM = Rotterdam; UT = Utrecht"
       )
  

```

```{r, eval=F}
#**BC**

set.seed(2)

bc0 %>%
ggplot(aes(y = Mean,
           x=Sampling_type,
           )) + 
  geom_boxplot(col=alpha("black", 0.4)) + 
  geom_point(aes(col = This_study,
                 ),
             alpha=0.6
             ) + 
  # #city names 
  geom_text_repel(size = text_size-0.5, 
                  fontface = "plain",  
                  #make labels more spaced out
                  #box.padding = 0.5, 
                  aes(label = Location,
                      col = This_study
                      )
                   ) +
  
  # Add a second axis and specify its features
  scale_y_continuous(sec.axis = sec_axis(~./bc_ours$Mean, 
                                         name="BC Conc Relative to This Study", 
                                         breaks = c(1, seq(5,20,5))
                                         )
                     ) +
  #this study
  geom_hline(yintercept = bc_ours$Mean,
             linetype = "dashed",
             alpha=0.5
             ) +
  
  theme(
  # axis.text.x = element_blank(),
  # axis.ticks = element_blank(),
  
  legend.position = "none"
  ) +
  
  labs(x = "Sampling Type",
       y = "Mean BC Conc (pt/cm3)",
       title = "Mean BC Conc reported in other LUR UFP studies",
       caption = "AMS = Amsterdam; MMa = Maastricht; RTM = Rotterdam; UT = Utrecht"
       )
  

```


## Tables

**UFP**

```{r}
ufp0 %>%
  group_by(Study) %>%
   select(Study, Location,  
    Mean_UFP_Conc = Mean,
    Instrument, Instrument_range
  ) %>%
    
   rename_if( grepl("_", names(.)), ~gsub("_", " ", .) ) %>%
  kable(caption = "Average UFP Concentrations (pt/cm3) reported in other LUR UFP studies", 
        digits = 2) %>%
  kable_styling()
  
```

**BC** 

```{r}
bc0 %>%
  #group_by(Study) %>%
   select(Study, Location,  
    Mean_BC_Conc = Mean,
  ) %>%
   rename_if( grepl("_", names(.)), ~gsub("_", " ", .) ) %>%
  kable(caption = "Average BC Concentrations (ng/m3) reported in other LUR UFP studies", 
        digits = 2) %>%
  kable_styling()
```


# UFP instrument ranges

- instrument ranges are not necessarily related to observed higher/lower concentrations? 

```{r, fig.height=6}
ufp0 %>%
  mutate(Study = paste0(Location, "; ", 
                        Study)
         ) %>%
   ggplot(aes(y= Study, #Location, 
             xmin = instrument_low_range_nm, xmax=instrument_high_range_nm,
             col = Mean,
             )) + 
  geom_linerange(size=1.5
    ) +
  scale_color_gradient(low = "yellow", high = "red") +
  
  scale_x_log10(breaks = c(4, 10, 20, 300, 1000, 3000)) +
  
  facet_grid(rows = vars(Sampling_type), 
             scales="free_y", #space="free_y"
             ) +
  
  labs(x = "Instrument Particle Size Range (nm)",
       col = "Avg PNC\n(pt/cm3)",
       title = "UFP size ranges collected in various UFP studies"
       )
  

```


# Other

### --> ? delete these plots?

**UFP**

```{r}

### ???delete?

ufp_r2 %>%
  #?
  #filter(!grepl("Fixed", Sampling_type)) %>%
  
  gather("var", "value", sites_per_model, visits_per_site, total_site_duration_min) %>%
  
  ggplot(aes(x=value, y=Mean_Model_R2,
             col = Sampling_type
             )) + 
  geom_smooth(se=F, method="lm") +  
    geom_point(alpha=0.5) + 

  facet_grid(~var, #~var, 
             scales="free_x")
  
  
```

THe more sites, the fewer visits likely to occur

```{r}
# ? delete? 

ufp_r2 %>%
  ggplot(aes(x=sites_per_model, y=visits_per_site, col=Mean_Model_R2)) + 
  geom_point() +
    geom_smooth(se=F, method="lm") +  

  facet_wrap(~Sampling_type, scales="free_x"
             ) + 
  labs(
    x = "Sites",
    y = "Repeat visits per site",
    title = "Site visits vs number of sites in different studies\nUFP"
  )

```

fixed and short-term mobile with longer site sampling durations generally have higher R2. This is not true for short-term stationary

```{r}
# ? delete? 

ufp_r2 %>%
  ggplot(aes(#x=sites_per_model, 
             x=total_site_duration_min, y=Mean_Model_R2,
             col = sites_per_model
             )) + 
  geom_point() +
    geom_smooth(se=F, method="lm") +  

  facet_wrap(~Sampling_type, scales="free") + 
  labs(
    y = "R2",
    x = "Total site duration (min)",
    title = "Total site sampling duration vs number of sites in different studies\nUFP"
  )


```

 

UFP levels by publication year

```{r}
ufp0 %>%
   ggplot(aes(y= Mean, #Location, 
              x=Year
              
             )) + 
  geom_point() + 
  labs(y = "Mean UFP (pt/cm3)",
       title = "UFP concentrations reported, by study publication year"
       )

```


