---
title: "Location Jittering"
author: "Magali Blanco"
date: ' `r Sys.Date()` '
output: html_document
editor_options: 
  chunk_output_type: console
---

# About this script 

Purpose:  
* script jitteres locations, typically for mapping Cohort locations

Notes:   
* jittering results in some locations being moved to water locations. This may only be a problem if generating a highly zoomed in map. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,  
                      cache=F, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE
                      )  

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(kableExtra,
               tidyverse
               
               )

source("0.Global_Fns.R")

set.seed(1)

```

# Calculating the maximum jittered distance between original and jittered lcoations 

Since jittering occurs in the x and y directions, the final distance between an original and a jittered location can be calculated using the Pythagorean Theorem:

$$a^2 + b^2 = c^2 $$

where:   
* $a$ and $b$ are the amount of jittering that occurs in the lambert_x and lambert_y coordinates, and $c$ is the difference between the original and the jittered location

We will jitter x and y the same amount such that $a=b$ and thus:

$$a^2 + a^2 = c^2 $$  
$$2a^2 = c^2 $$

If we want to set a maximum jittered distance ($c$), we can solve for $a$ to determine how much jittering should occur in both directions:

$$a = \sqrt{\frac{c^2}{2} } $$


```{r}
# fn calculates jittering "amount" given a maximum distanace that any point should be jittered

max_point_distance <- function(c){
  a =sqrt(c^2/2)
  return(a)
  
}

```


```{r}
max_dist <- 1000

```

Thus, if we want the maximum jittered distance (c) = `r max_dist` m, lambert_x and y should be jittered by the following amount (a): 

```{r}
max_point_distance(max_dist)

```


# Jitter Cohort locations in the ST area

```{r}
cohort0 <- read.csv(file.path("Data", "ACT", "cohort_st_pm25_preds.csv"))[-1]

cohort <- cohort0 %>%
  mutate(
    jitter_lamb_x = jitter(lambert_x, 
                      # min/max amount of jittering that shoud be done (in CRS units - m)
                      amount = max_point_distance(max_dist)
                      ),
    jitter_lamb_y = jitter(lambert_y, 
                      # min/max amount of jittering that shoud be done (in CRS units - m)
                      amount = max_point_distance(max_dist)
                      ),
    
    
  )

```

```{r}
# # difference between jitter & original coordinates
# cohort %>%
#   mutate(
#     diff_x = jitter_lamb_x - lambert_x
#   ) %>%
#   distribution.table(var.string = "diff_x")
 

#difference in original and jittered location (c)
cohort %>%
  mutate(
    c = sqrt((jitter_lamb_x-lambert_x)^2 + (jitter_lamb_y-lambert_y)^2)
  ) %>%
  distribution.table(var.string = "c") %>%
  kable(caption = "Amount of jittering (m) that occurred for the cohort locations.") %>%
  kable_styling()

 
cohort %>%
  mutate(
    c = sqrt((jitter_lamb_x-lambert_x)^2 + (jitter_lamb_y-lambert_y)^2)
  ) %>%
  ggplot(aes(x=c)) +
  geom_histogram() + 
  labs(x= "Total Point Location Jittering (m)",
      title = "Amount of jittering (m) that occurred for the cohort locations."  
       )


```


### --> Future work: see tutorial for ensuring that jittering is within a polygon? note: ST land area polygon may have geometry issues? 

https://stackoverflow.com/questions/27622483/how-can-ggplot2-keep-jittered-locations-within-a-map-boundary-such-as-a-us-stat 

 


# Save datasets

```{r}
# cohort %>%
#   # only keep jittered locations & PM2.5 predictions
#   #select(location_id, jitter_lamb_x, jitter_lamb_y, conc_lt) %>%
#   write.csv(., file = file.path("Data", "ACT", paste0("cohort_st_pm25_preds_jittered_",max_dist,  ".csv")),
#             row.names = F)



```

