---
title: "Grids"
author: "Magali Blanco"
date: "2/23/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

**Notes**

* lambert_x and lambert_y coordinates generated for covariates use CRS USA_Contiguous_Lambert_Conformal_Conic   
* for coordinates in meters, can use CRS 32148 (WA North)   
* for coordinates in degrees (lat/long), can use CRS 4326   

* grids w/ fine street-level contrast usually have a 50 m resolution

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,  
                      cache=F, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE
                      )  

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(tidyverse,
               sf,
               ggmap #::get_stamenmap()
               )

```

```{r}
#common variables
project_crs <- 4326  #lat/long
m_crs <- 32148

```

# Functions to create grids

```{r}
# fn returns grid given a point grid range & an estimated number of desired points
 
# min_x = 310000
# min_y = -15000
# max_x = 430000
# max_y = 235000
# no_pts = 10000
# id_suffix = "st"

make_grid <- function(
  # grid area
  min_x, min_y, max_x, max_y,
  # estimated number of desired points
  no_pts,
  id_suffix = "id"
  ) {
  
  
  #area of entire grid area
  grid_area <- (max_y-min_y)*(max_x-min_x)
  
  #area of each small box? - can't remember why mathematically this works eventhough the # of points != the number of boxes created inside the entire grid
  grid_area <- grid_area/no_pts
  
  #distance between each point (length of side of each small box)
  resolution <- sqrt(grid_area)
    
  y <- seq(min_y, max_y, resolution)
  x <- seq(min_x, max_x, resolution)
  
  new_grid <- expand.grid(x = x, y = y) %>%
    mutate(native_id = paste0(id_suffix, "_", 
                             #make all numbers the same length
                             str_pad(seq(1, length.out = nrow(.)),
                                     width = nchar(nrow(.)), 
                                     pad="0")
                             )
           ) %>%
    select(native_id, everything())
  
  return(list(
    grid = new_grid,
    resolution = resolution
    )
  )
  
}

```


```{r}
# fn returns coordinates for a different transformation. it convertes the dataset into a spatial object, calculates coordinates for a diff refernce system, converts these to a df, and attaches these to the original coordinates

# tutorial: https://ryanpeek.org/2017-08-03-converting-XY-data-with-sf-package/ 

add_crs <- function(
  dt,
  original_crs, original_coords,
  new_crs, new_coord_names = c("long", "lat") 
  ) {

  library(sf)
  
  #convert flat file to spatial file, give it the original CRS
  dt_sp <- st_as_sf(dt, coords = original_coords, crs = original_crs)
  
  #convert to different CRS, and save the coordinates
  new_coords <- st_transform(dt_sp, crs = new_crs) %>%
    st_coordinates() %>%
    as.data.frame() %>%
    rename(new_x = X, 
           new_y = Y) %>%
    mutate(new_crs = new_crs)  
  
  #rename new columns
  names(new_coords) <- c(new_coord_names, paste0(c(new_coord_names, "crs"), collapse = "_"))
  
  # add new coords to original dt
  dt2 <- cbind(dt, new_coords)
  
  
  return(dt2)
  
  
  }


```


# New MM Study Grid w/ even spacing

**don't update w/ fns b/c already sent this to Amanda?**

```{r}
# make new grid
no_pts <- 5000
  
#based on CRS EPSG: 4326
min_long <- -122.439228999999997 
max_long <- -121.879884000000004
min_lat <- 47.166826999999998
max_lat <- 48.060476000000001

grid_area <- (max_long-min_long)*(max_lat-min_lat)/no_pts
pt_dist <- sqrt(grid_area)
  
long <- seq(min_long, max_long, pt_dist)
lat <- seq(min_lat, max_lat, pt_dist)

new_grid <- expand.grid(longitude = long,
                     latitude = lat) %>%
  mutate(native_id = paste("grid_mm_", 
                           #make all numbers the same length
                           str_pad(seq(1, length.out = nrow(.)),
                                   width = nchar(nrow(.)), 
                                   pad="0")
                           )
         ) %>%
  select(native_id, everything())


```


# Grid for ST study area

Using fns

### --> modify fn to get resolution

```{r}
st_area <- read_sf(file.path("..",  "GIS", "Shapefiles", "Study area", "Spatiotemporal",  "st_area.shp")) 

buff <- 3000
#add 2k m buffer (for smoothing later)
st_area_buffer <- st_area %>%
  #convert to meters
  st_transform(m_crs) %>%
  st_buffer(buff) #%>%
  #convert back to original CRS
  #st_transform(project_crs) 

#plot(st_area)

pts <- 19000

st_grid <- make_grid(min_x = 310000, min_y = -15000, max_x = 430000,max_y = 235000,
                  no_pts = pts, 
                  id_suffix = "st"
                  )$grid %>%
  mutate(x_y_crs = "wa_north_m_32148") %>%
  
  # add lat/long coordinates
  add_crs(dt = ., 
                original_crs = m_crs, original_coords = c("x", "y"), 
                # calculate lat/long using a WA reference system
                new_crs = project_crs, new_coord_names = c("long", "lat") 
                ) 

 
#only keep points in ST area
st_grid_small <- st_grid %>%
  st_as_sf(., coords=c("x","y"), crs= m_crs,
           #coords=c("long","lat"), crs= project_crs
           remove=F
           ) %>%
  filter(as.logical(st_intersects(., st_area_buffer))) %>%  
  as.data.frame() %>% select(-geometry)

```

plot 

### --> why is this plotted in lat/long??

```{r}


ggplot() + 
  geom_point(data=st_grid_small, 
             aes(x=x, y=y),
             #inherit.aes = F,
          alpha=0.3
          ) +
geom_sf(data=st_area, inherit.aes = F,
        alpha=0.3
        )

```

plot to make sure it looks ok: Full & zoomed in version to make sure points are evenly spaced out

```{r}
# bbox <- st_bbox(st_area)
# names(bbox) <- c("left", "bottom", "right", "top")
# background_map <- suppressMessages(get_stamenmap(bbox =  bbox,
#                                                  maptype = "terrain",
#                                                  zoom = 8
#                                                  )) 
# 
# 
#  
# # check that smaller st pts are within ST area 
# ggmap(background_map) + 
#   geom_sf(data=st_area, inherit.aes = F) +
#   geom_sf(data=st_grid_small, inherit.aes = F,
#           alpha=0.5
#           )  

```

```{r}

st_grid_small %>%
  ggplot(aes(x=x, y=y)) + 
  geom_point(alpha=0.2) +
  #  ensures that one unit on the x-axis is the same length as one unit on the y-axis
  coord_fixed(ratio = 1) 

# zoomed in - #look at a few points to make sure they look even
st_grid %>%
  filter(x < quantile(x, 0.05),
         y < quantile(y, 0.03),
         ) %>%
  ggplot(aes(x=x, y=y)) + 
  geom_point() +
  #  ensures that one unit on the x-axis is the same length as one unit on the y-axis
  coord_fixed(ratio = 1) 



```


# Grid for Seattle

Using fns

```{r}

seattle_grid <- make_grid(min_x = 380812, min_y = 44440, 
                          max_x = 402503,max_y = 86964,
                  no_pts = 5000, 
                  id_suffix = "seattle"
                  )$grid %>%
  mutate(x_y_crs = "wa_north_m_32148") %>%
  
  # add lat/long coordinates
  add_crs(dt = ., 
                original_crs = 32148, original_coords = c("x", "y"), 
                new_crs = 4326, new_coord_names = c("long", "lat") 
                ) 

 
```

plot to make sure it looks ok: Full & zoomed in version to make sure points are evenly spaced out

```{r}

seattle_grid %>%
  ggplot(aes(x=x, y=y)) + 
  geom_point(alpha=0.2) +
  #  ensures that one unit on the x-axis is the same length as one unit on the y-axis
  coord_fixed(ratio = 1) 

# # zoomed in - #look at a few points to make sure they look even
# seattle_grid %>%
#   filter(x < quantile(x, 0.05),
#          y < quantile(y, 0.03),
#          ) %>%
#   ggplot(aes(x=x, y=y)) + 
#   geom_point() +
#   #  ensures that one unit on the x-axis is the same length as one unit on the y-axis
#   coord_fixed(ratio = 1) 



```


# Grid for Downtown Seattle


```{r}
# pts are ~ 140 m apart (pt_dist)
dt_grid <- make_grid(min_x = 383218, min_y = 61864, 
                          max_x = 391900, max_y = 72925,
                  no_pts = 5000, 
                  id_suffix = "dt"
                  )$grid %>%
  mutate(x_y_crs = "wa_north_m_32148") %>%
  
  # add lat/long coordinates
  add_crs(dt = ., 
                original_crs = 32148, original_coords = c("x", "y"), 
                new_crs = 4326, new_coord_names = c("long", "lat") 
                ) 

 
```

plot to make sure it looks ok. 

```{r}

dt_grid %>%
  ggplot(aes(x=x, y=y)) + 
  geom_point(alpha=0.2) +
  #  ensures that one unit on the x-axis is the same length as one unit on the y-axis
  coord_fixed(ratio = 1) 


```


# Save grids

```{r}
# grid for MM study area
#write.csv(new_grid, file = file.path("Output", "new_grid.csv"), row.names = F)

# grids for PM2.5 maps
# write.csv(st_grid, file = file.path("Output", "st_grid.csv"), row.names = F)
# write.csv(st_grid_small, file = file.path("Output", "st_grid_small.csv"), row.names = F)
# write.csv(seattle_grid, file = file.path("Output", "seattle_grid.csv"), row.names = F)
# write.csv(dt_grid, file = file.path("Output", "downtown_grid.csv"), row.names = F)

```

 